### 2022-08-02 02-structure.Rmd
structure

# Overview {-}

<!-- This page is useful because by calling the indicator rmd files as child files we bypass the need to sequentially numbering them. An overview page is anywhays useful.  -->

## Test 1 {-}
This indicator is about ....

## Test 2 {-}
And this indicator is about...

```{r, child= c('indicators/indicatorTest.Rmd')}
```

```{r, child= c('indicators/indicatorTest2.Rmd')}
```