### 2022-08-02 index.Rmd

--- 
title: "Indicators for Ecosystem Condition in Norway"
author: "Anders L. Kolstad"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
url: https://ninanor.github.io/ecosystemCondition/index.html
# cover-image: path to the social sharing image like images/cover.jpg
description: |
  This is a site for documenting the analyses related to the development and design of ecosystem condition indicators.
biblio-style: apalike
csl: chicago-fullnote-bibliography.csl
---

# Introduction {-}

On this site you may view the nitty gritty details for how cetain indicators for ecosystem condtion has been calculated.

To do:

Embed shiny app?

Link to author guidelines?