### 2022-08-02 indicators/indicatorTest.Rmd

# This is an indicator {-#test1}

```{r}
plot(rnorm(10,10,10))
```