### 2022-08-05 02-overview.Rmd
proj6

# Overview of indicators {-}

<!-- This page is useful because by calling the indicator rmd files as child files we bypass the need to sequentially numbering them. An overview page is anywhays useful.  -->

## Areal av isbreer / Glacier area {-}
This indicator reflects the relative loss of glacial area between tha last climatic normal persion and today.

```{r, echo=F}
source("indicators/indicatorMetaData/Areal av isbreer.R")
knitr::kable(t(out), 
            format = "html",
            caption = paste("Fakatark for tilstandindikatoren ", 
                        out$Indikator, col.names = NULL))
```

## Indicator X {-}
And this indicator is about...

```{r, child= c('indicators/indicatorTest.Rmd')}
```

```{r, child= c('indicators/indicatorTest2.Rmd')}
```

```{r, child= c('indicators/breareal.Rmd')}
```