### 2022-08-08 02-overview.Rmd
rm com. rev. ch.

# Overview of indicators {-}

*_Lacking content_*
The idea of this page is to give a quick overview of the available indicators, and to possibly present the Norwegian fact sheets for each. For now this page is just an example what it could look like.

## Areal av isbreer / Glacier area {-}
This indicator reflects the relative loss of glacial area between tha last climatic normal persion and today.

```{r, echo=F}
source("indicators/indicatorMetaData/Areal av isbreer.R")
knitr::kable(t(out), 
            format = "html",
            caption = paste("Fakatark for tilstandindikatoren ", 
                        out$Indikator, col.names = NULL))
```

## Indicator X {-}
And this indicator is about...