### 2022-08-18 naturtype.Rmd
scaling and nat.type

# (PART\*) GENERAL TOPICS {-}

# On the application of the naturetype dataset {#naturtype}

<br />

```{r, include=FALSE}
library(knitr)
library(sf)
library(tmap)
library(ggplot2)
library(units)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

Author: Anders L. Kolstad

Date: `r Sys.Date()`

Here I will investigate a specific data set, [Naturtyper - Miljødirektoratets instruks](https://kartkatalog.geonorge.no/metadata/naturtyper-miljoedirektoratets-instruks/eb48dd19-03da-41e1-afd9-7ebc3079265c) and try to evaluate its usability for designing indicators for ecosystem condition. This involves assessing both the spatial and temporal representativity, as well as conseptual alignment with the [Norwegian system for ecosystem condition assessments](https://www.regjeringen.no/no/dokumenter/fagsystem-for-fastsetting-av-god-okologisk-tilstand/id2558481/).

The precision in field mapping will not be assess in itself. We assume some, or even considerable, sampling error, but this is inherent to all field data.

## Import data and general summary statistics

* Total mapped area
* Temporal trend
* Histogram of conditions (shiny?)

```{r}
# local path
path <- "R:/GeoSpatialData/Habitats_biotopes/Norway_Miljodirektoratet_Naturtyper_nin/Original/Natur_Naturtyper_NiN_norge_med_svalbard_25833.gdb"
dat <- sf::st_read(dsn = path)
names(dat)
```
Import outline of Norway
```{r}
nor <- sf::read_sf("data/outlineOfNorway_EPSG25833.shp")
```

The dataset has 95k polygons, each with 36 variables:
```{r}
dim(dat)
```
It's therefore takes a little minute to render a plot, but this is the code to do it:

```{r, eval = T}
tmap_mode("view")
tm_shape(dat) + 
  tm_polygons(col="tilstand")+
  tm_shape(nor)+
  tm_polygons(alpha = 0,border.col = "black")
```

### Area
Calculating the area

```{r}
dat$area <- sf::st_area(dat)
summary(dat$area)
```

The smallest polygons are 20 m2, and the biggset is 8.5 km2

Sum of mapped area divided by Norwegian mainland area
```{r}
sum(dat$area)/sf::st_area(nor)
```

About 0.5% has been mapped.
It is therefore essential that these 0.5% are representative (of the remaining 99.5%).

## Overview of nature types
* What are the dominant ecosystem groups recorded. 
```{r}
ggplot(dat, aes(x = hovedøkosystem))+
  geom_bar()+
  coord_flip()
```

This project is concerned only with wetlands (våtmark), semi-natural areas (semi-naturlig mark) and naturally open areas below the forest line (naturlig åpne områder i lavlandet/under skoggrensa).
```{r}
dat$hovedøkosystem[dat$hovedøkosystem == "Naturlig åpne områder i lavlandet"] <- "Naturlig åpne områder under skoggrensa"
```

```{r}
unique(dat$hovedøkosystem)
utvlagt <- c("Semi-naturlig mark",
             "Våtmark",
             "Naturlig åpne områder under skoggrensa")
dat2 <- dat[dat$hovedøkosystem %in% utvlagt,]
```

```{r}
temp <- tapply(dat2$area, dat2$hovedøkosystem, sum)
temp <- data.frame(eco = row.names(temp),
                   area = temp$temp)
ggplot(temp, aes(x = eco, y=area))+
  geom_bar(stat="identity")+
  coord_flip()
```

The same bias towards sami-natural areas are seen when looking at the area mapped, as when looking the number of polygons.

* What are the dominant nature types recorded.

```{r}

temp <- dat2[dat2$hovedøkosystem=="Våtmark",]
ggplot(temp, 
       aes(x = naturtype,
           y = area))+
  geom_bar(stat = "sum")+
  coord_flip()

temp <- tapply(temp$area, temp$naturtype, sum)
temp <- data.frame(nat = row.names(temp),
                   area = temp)

temp <- temp[order(temp$area),]
ggplot(temp[order(temp$area),], 
       aes(x = nat,
           y = area))+
  geom_bar(stat = "identity")+
  coord_flip()
```

* Can we group these into main ecosystem type categories?
* Are there any regional differences

## Environmental space

* Extract environmental data from the polygons and compare against reference data points.
* Uni-variate (e.g. _boreal hei_ plotted against elevation, oceanity, distance from roads, etc)

## Represntaivity in mapping units
Only a subset of nature types are mapped. How well do the mapped units represent, or cover, the range of nature types that is included for each main ecosystem group?

## Conclution 1 
Conclude about which main ecosystem types that can reliably assess using the available mapping units?

What about suitability for defining reference values?

## Suitability 
In the subset of mapping units that can be said to represent an ecosystem group, which NiN or MdirPRAM variables should we use?

## Concrete examples
### Andel lynghei med minst to faser i kystlynghei

### Areal uten skadet og/eller død røsslyng i kystlynghei

### Aktuell bruksintensitet