### 2022-11-15 DRAFT_rosslyng.Rmd
add røsslyng indicator

---
title: "DEPRECATED: Areal uten død eller skadet røsslyng i kystlynghei"
output:
  html_document
---

# DEPRECATED: Areal uten død eller skadet røsslyng i kystlynghei

<br />

_Author and date:_
Anders L. Kolstad

```{r}
Sys.Date()
```

<br />

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(knitr)
library(sf)
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Fill in which ecosystem the indicator belongs to, as well as the ecosystem characteristic it should be linked to. It's OK to use some Norwegian here -->
```{r, echo=F}
Ecosystem <- "Semi-naturlig mark" # e.g. "Skog og fjell"
Egenskap  <- "Primærproduksjon" # e.g. "Primærproduksjon"
ECT       <- "Structual state characteristic" # e.g. "Structural state characteristic"
Contact   <- "Anders L. Kolstad" # e.g. "Anders Kolstad"
```

```{r, echo=F}
metaData <- data.frame(Ecosystem,
                       "Økologisk egenskap" = Egenskap,
                       "ECT class" = ECT)
knitr::kable(metaData)
```

<!-- Don't remove these three html lines -->
<br />
<br />
<hr />

<!-- Document you work below. Try not to change  the headers too much. Data can be stored on NINA server. Since the book is rendered on the R Server this works fine, but note that directory paths are different on the server compared to you local machine. If it is not too big you may store under /data/ on this repository -->

## Introduction
Here I quickly go through the ANO data set for the years 2019-2021 to investigate the frequeny of poins that fall within coastal heathands (kystlynghei). This will decide wether we can use this data to inform an indikcator about the proportion of dead Calluna vulgaris (røsslyng).

## Analyses

#### ANO data
Data is downloaded from [here](https://kartkatalog.miljodirektoratet.no/Dataset/Details/2054).

Imnport and list terms: 
```{r}
ano <- sf::st_read("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Naturovervaking_eksport.gdb",
                   layer = "ANO_SurveyPoint")
#st_crs(ano)
names(ano)

```

Table of the number of sampling points per year:
```{r}
table(ano$aar)
```

```{r, fig.cap="The distribution of ANO points that fall within different main ecosystems."}
par(mar=c(15,5,0,0))
barplot(table(ano$hovedoekosystem_punkt), las=2)
```

Sub-setting the data to only consider seni natural areas.
```{r}
snat <- ano[ano$hovedoekosystem_punkt == "semi_naturlig_mark",]
snat2 <- snat[!is.na(snat$GlobalID),]
```
230 points are semi-natural.

```{r}
table(snat$hovedtype_1m2)
```
Only 49 1m2 samples are kystlynghei.

```{r}
table(ano$hovedtype_250m2)
```
Only 26 of the 250m2 circles are kystlynghei.

## Conclusion
Because of the random allocation of sampling points, not enough points fall in coastal heathlands to allow us to create a trustworth indicator for this nature type. With a few more years of data we might end up with about 100 points, which is borderline what we need in order to calculate condition values at a regional level with some level of precition.