### 2023-01-20 DRAFT_homogeneous_impact_areas.Rmd
bug: Cache large operation

---
title: "-"
output:
  html_document
editor_options: 
  markdown: 
    wrap: sentence
---

# Homogeneous Impact Areas

<br />

*Author and date:*

Anders Kolstad

```{r, echo=FALSE}
Sys.Date()
```

<br />

<!-- Load all you dependencies here -->

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(tmap)
library(dplyr)
library(stars)
library(ggplot2)
library(units)
library(ggpubr)
library(DT)
#library(osmplotr)
```

## Introduction

This chapter documents the creation of a wall-to-wall map of homogeneous impact areas in Norway.
The map is produces by binning values in the infrastructure index into four discrete categories.
It is likely to be a good predictor for some indicators, such as *slitasje* and the presence of *alien species*.

## About the underlying data

The infrastructure index is explained [here](https://brage.nina.no/nina-xmlui/handle/11250/2983607).
It is a wall-to-wall raster over Norway with 100 m resolution.
Each pixel is given a value along a continuous gradient from 0 to around 15.4, representing the frequency or surrounding cells within 500 m with human infrastructure (houses, roads, ect).

### Representativity in time and space

The infrastructure index is calculated with data that have slightly different dates, but can be said to represent the current status, the period 2010-2020.
The data covers all of mainland Norway.

### References

Bakkestuen, V., Dervo, B.K., Bærum, K.M. og Erikstad, L. 2022. 
Prediksjonsmodellering av naturtyper i ferskvann.
NINA Rapport 2079.
Norsk institutt for naturforskning.

## Analyses

### Import data

The path must be conditional:

```{r}
dir <- substr(getwd(), 1,2)
```

```{r}
path <- ifelse(dir == "C:", 
        "R:/GeoSpatialData/Utility_governmentalServices/Norway_Infrastructure_Index/Original/Infrastrukturindeks_UTM33/infra_tiff.tif",
        "/data/R/GeoSpatialData/Utility_governmentalServices/Norway_Infrastructure_Index/Original/Infrastrukturindeks_UTM33/infra_tiff.tif")
```

Import a stars proxy (no data imported yet)

```{r}
infra <- stars::read_stars(path)
```

Print the coordinate reference system:
```{r}
st_crs(infra)$input
```

### Trondheim example

It's easier to see what's happening if we zoom in a bit.
Lets get a boundary box around Trondheim.

```{r}
myBB <- st_bbox(c(xmin=260520.12, xmax = 278587.56,
                ymin = 7032142.5, ymax = 7045245.27),
                crs = st_crs(infra))
```

Cropping the raster to the bbox

```{r}
infra_trd <- sf::st_crop(infra, myBB)
```

#### Get OSM highways

Lets download some base maps to help validate and contextualize the infrastructure index.

Transform to lat long due to osm requirements

```{r, warning=F, eval=F}
infra_trd_ll <- sf::st_transform(infra_trd, 4326)
```

Get the boundary box of the cropped raster

```{r, eval=F}
myBB_ll <- sf::st_bbox(infra_trd_ll)
```

Download highways using the above bbox.

```{r, warning=FALSE, message=F, eval=F}
hw <- 
  osmplotr::extract_osm_objects(
    bbox = myBB_ll,
    key = "highway",
    sf = T)
```

Transforming highway data back into utm, although not strictly necessary.

```{r, eval=F}
hw_utm <- sf::st_transform(hw, sf::st_crs(infra_trd)) 
```

This object contains too many roads (about 30k).
I'll take out the unnamed roads.

```{r, eval=F}
hw_utm <- hw_utm[!is.na(hw_utm$name),]
```

```{r, eval=F, echo=F}
# Caching the sf object
saveRDS(hw_utm, "data/cache/highways_trondheim.rds")
```

```{r, echo=F}
#Load object
hw_utm <- readRDS("data/cache/highways_trondheim.rds")
```

#### Discretize

Here I define a simplified categorical typology for the infrastructure index using four classes.

```{r}
names(infra_trd) <- "infrastructureIndex"
infra_trd_reclassed <-  infra_trd %>%
  mutate(infrastructureIndex = case_when(
    infrastructureIndex < 1 ~ 0,
    infrastructureIndex < 6 ~ 1,
    infrastructureIndex < 12 ~ 2,
    infrastructureIndex >= 12 ~ 3
  ))
```

Lets plot these two maps side by side

```{r}
map_trd_reclassed <- tm_shape(infra_trd_reclassed)+
  tm_raster(title="Infrastructure Index",
            #palette = "viridis",
            style="cat")+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="white")
```

```{r}
map_trd <- tm_shape(infra_trd)+
  tm_raster(title="Infrastructure Index",
            style="cont",
            palette = "viridis")+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="white")
```

```{r, fig.cap="Infrastructure indeks over Trondheim, comparing the continous scale with the ordinal four-step scale. Major roads are in white."}
tmap_arrange(map_trd,
             map_trd_reclassed,
             ncol=1)
```

I tweaked the thresholds for the bins so that the categories match my knowledge about the land use intensity in Trondheim, for example that most of the forest next to Trondheim (to the left in the map) was in the second lowest class.
This looks pretty good to me.
It depicts a gradient in human presence from high within the buildt zone, to *no to very low* in the forests and mountains to the west.
Note that there is still considerable human activity also there in the form of outdoor recreation and even forestry.

#### Aggregate

The resolution in this map is more than we need, and the size of the data implies that the next step, the vectorization, would take too long.
We therefore aggregate, ar reduce the resolution.

```{r}
dim <- st_dimensions(infra)
paste("Resolution is", dim$x$delta, "by", dim$x$delta, "meters")
```

```{r, warning=F, message=F}
infra_trd_reclassed_agg <- st_warp(infra_trd_reclassed, cellsize = c(1000, 1000), 
                                   crs = st_crs(infra_trd_reclassed), 
                                   use_gdal = TRUE,
                                   method = "average")
```

```{r}
dim <- st_dimensions(infra_trd_reclassed_agg)
paste("Resolution is", dim$x$delta, "by", dim$x$delta, "meters")
```

Each cell is now the average of the aggregated cells, and hence the value is continuous again.
Let's make it discrete.

```{r}
names(infra_trd_reclassed_agg) <- "infrastructureIndex"
infra_trd_reclassed_agg <-  infra_trd_reclassed_agg %>%
  mutate(infrastructureIndex = case_when(
    infrastructureIndex < 1 ~ 0,
    infrastructureIndex < 6 ~ 1,
    infrastructureIndex < 12 ~ 2,
    infrastructureIndex >= 12 ~ 3
  ))
```

```{r, fig.cap="Infrastrcture indeks over Trondheim on a four-step discrete scale, aggregated to 1x1 km resolution using the mean function."}
tm_shape(infra_trd_reclassed_agg)+
  tm_raster(title="Infrastructure Index",
            style="cat")+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="white")
```

This resolution is more that good enough for our purpose here.
I see that the map extends a bit into the fjord. We want to cut away these areas. We can use a map of the outline of Norway to do this.

```{r}
outline <- st_read("data/outlineOfNorway_EPSG25833.shp")
```

The CRS needs to be identical 
```{r}
infra_trd_reclassed_agg <- st_transform(infra_trd_reclassed_agg, 25833)
```

```{r}
infra_trd_reclassed_agg_terrestrial <- st_crop(infra_trd_reclassed_agg, outline)
```

```{r, fig.cap="Infrastrcture indeks over Trondheim on a four-step discrete scale, aggregated to 1x1 km resolution using the mean function."}
tm_shape(infra_trd_reclassed_agg_terrestrial)+
  tm_raster(title="Infrastructure Index",
            style="cat")+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="white")+
  tm_shape(outline)+
  tm_borders(lwd =2,
             col = "black")
```

I have treated the raster cells as points when cropping, so that cells where the center or the cell is outside he terrestrial delineation are removed. I think this makes the most sense for unbiased area statistics, but some coastal indicator data could also be excluded because of this. The errors would be smaller with a finer resolution raster, but computation time is a problem.

### Aggregate the entire map to 1x1 km

```{r, eval=F}
# runtime about 30 sec
infra_agg <- st_warp(infra, cellsize = c(1000, 1000), 
                                   crs = st_crs(infra), 
                                   use_gdal = TRUE,
                                   method = "average")
# The CRS needs to be identical for st_crop
infra_agg <- st_transform(infra_agg, 25833)
```

### Cut out marine areas
I see that the map extends a bit into the fjord. We want to cut away these areas. We can use a map of the outline of Norway to do this.

```{r, eval=F}
infra_agg_terrestrial <- st_crop(infra_agg, outline)
saveRDS(infra_agg_terrestrial, "P:/41201785_okologisk_tilstand_2022_2023/data/cache/infra_agg_terrestrial.rds")
```

```{r, include=F}
# silently load cached file
infra_agg_terrestrial <- readRDS("P:/41201785_okologisk_tilstand_2022_2023/data/cache/infra_agg.rds")
```

### Discretize the entire map

```{r}
names(infra_agg_terrestrial) <- "infrastructureIndex"

infra_agg_discrete <- infra_agg_terrestrial %>%
  mutate(infrastructureIndex = case_when(
    infrastructureIndex < 1 ~ 0,
    infrastructureIndex < 6 ~ 1,
    infrastructureIndex < 12 ~ 2,
    infrastructureIndex >= 12 ~ 3
  ))
```

### Vectorize

This step might seem rather stupid.
We want to vectorize a rather large raster.
This makes it a quite big data object.
The reason is that there is no really good way to burn polygon data on to raster grid cells after the disuse of the `{raster}` package.
It was not straight forward then either.
But calculating intersections between polygons is very fast and easy.

```{r, warning=F}
infra_agg_discrete_vect <- eaTools::ea_homogeneous_area(infra_agg_discrete, groups = infrastructureIndex)
```

Let's plot this now.
First, lets crop it to reduce computation time.

```{r}
myBB <- st_bbox(c(xmin=260520.12, xmax = 278587.56,
                ymin = 7032142.5, ymax = 7045245.27),
                crs = st_crs(infra_agg_discrete_vect))

infra_agg_discrete_vect_trd <- st_crop(infra_agg_discrete_vect, myBB)
```

```{r test-2, fig.cap="A vectorized version of the Infrastructure indeks over Trondheim on a four-step discrete scale."}
tm_shape(infra_agg_discrete_vect_trd)+
  tm_polygons(col="infrastructureIndex",
              style = "cat")+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="white")
```

As can be seen in the figure above, `st_warp` merges grid cells that have the same value, and the vectorized raster doesn't not end up being that big in the end.

## Chech national distribution

```{r}
regions <- sf::read_sf("data/regions.shp", options = "ENCODING=UTF8")
unique(regions$region)
```

```{r}
st_crs(infra_agg_discrete_vect) == st_crs(regions)
```

Since the two layers are completely overlapping, we can get the intersections
```{r, eval=F}
infra_stats <- eaTools::ea_homogeneous_area(infra_agg_discrete_vect,
                             regions,
                             keep1 = "infrastructureIndex",
                             keep2 = "region")
saveRDS(infra_stats, "P:/41201785_okologisk_tilstand_2022_2023/data/cache/infra_stats.rds")
```

```{r, echo=F}
# silently read cached file
infra_stats <- readRDS("P:/41201785_okologisk_tilstand_2022_2023/data/cache/infra_stats.rds")
```

Let's calculate the areas of these polygons and compare the HIF in the five regions. 
```{r HIF-region, fig.cap="Stacked barplot showing the distribution of human impact factor across five regions in Norway."}
infra_stats$area <- sf::st_area(infra_stats)

temp <- as.data.frame(infra_stats) %>%
  group_by(region, infrastructureIndex) %>%
  summarise(area = mean(area))

ggarrange(
  ggplot(temp, aes(x = region, y = area, fill = factor(infrastructureIndex)))+
    geom_bar(position = "stack", stat = "identity")+
    guides(fill = "none")+
    coord_flip()+
    xlab("")
  ,
  ggplot(temp, aes(x = region, y = area, fill = factor(infrastructureIndex)))+
    geom_bar(position = "fill", stat = "identity")+
    guides(fill = guide_legend("HIF"))+
    coord_flip()+
    xlab("")
)
```

This distribution looks reasonable. The relative proportions are similar in the five regions, but _Østlandet_ and _Sørlandet_ have more infrastructure in general.

Here are the numbers behind the figure above.
```{r}
temp$area <- round(temp$area,0)
DT::datatable(temp)
```

## Export
I will write this data to the NINA P server.

```{r, eval=F}
saveRDS(infra_agg_discrete_vect, "P:/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks/homogeneous_impact_areas.rds")
```

```{r, eval=F}
HIA_returned <- readRDS("P:/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks/homogeneous_impact_areas.rds")
```