### 2023-02-17 DRAFT_master_grid.Rmd
create master grid

---
title: ""
output:
  html_document
---

# Master grid

<br />

*Author and date:* Anders L. Kolstad

<br />

<!-- Load all you dependencies here -->

```{r setup, include=FALSE}
library(knitr)
library(sf)
library(stars)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This chapter describes the making of a master raster grid for all of mainland Norway. This grid is used to align all ecosystem and indicator maps so that it becomes possible to do easy resamping, masking and aggregating.

## About the underlying data

I will use the [statistical grid from Norway (5x5km)](https://kartkatalog.geonorge.no/metadata/statistisk-rutenett-5000m/32ac0653-d95c-446c-8558-bf9b79f4934e) and resample this to a higher resolutio of 50 x 50 meters.

## Analyses

```{r}
# Set up conditional file paths
dir <- substr(getwd(), 1,2)

pData <- ifelse(dir == "C:", 
               "P:/41201785_okologisk_tilstand_2022_2023/data/",
               "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/")
```

Import data
```{r}
#st_layers(paste0(pData, "Basisdata_0000_Norge_25833_StatistiskRutenett5km_FGDB.gdb"))
masterGrid_50m <- sf::read_sf(paste0(pData, "Basisdata_0000_Norge_25833_StatistiskRutenett5km_FGDB.gdb")) 
```

```{r, eval=F}
tmap_mode("view")
tm_shape(grid_5km$SHAPE)+
  tm_borders()
```

![](images/masterGrid_screenShot.PNG){width="800"}

Confirming CRS is EPSG25833:
```{r}
st_crs(grid_5km)
```

### Make grid

Use the bbox and split into 50 x 50 meter cells

```{r}
(masterGrid_50m <- st_as_stars(st_bbox(grid_5km), dx = 50, dy = 50))
```

This file is 6GB.

Number of cells, in millions, is:
```{r}
(nrow(masterGrid_50m)*ncol(masterGrid_50m))/10^6
```

So, quite a lot. For the area accounts we might consider using 10x10m.

### Eksport file (final product)

```{r}
stars::write_stars(masterGrid_50m, paste0(pData, "masterGrid_50m.tiff"))
```

### Test import
```{r}
temp <- stars::read_stars(paste0(pData, "masterGrid_50m.tiff"))
st_dimensions(temp)
```