### 2023-03-09 DRAFT_functional plant indicators_wetland.Rmd
data handling wetland Rmd continued

---
title: "Template for indicator documentation"
output:
  html_document
---

<!-- This is a template for how to document the indicator analyses.

Make a copy of this file and store it in the root folder (same as where this template is stored) folder. 
Name te file DRAFT_myIndicatorName.Rmd.

Because this is a stand-alone Rmd file inside a bookdown project, in order to knit the file for preview purposes,
you need to type rmarkdown::render("template.Rmd") in the console, 
as the the usual Ctrl + Alt + k will not work.

Make sure also to not change the order, or modify the headers, unless you really need to. This is because it
easier to read the web page as a whole of the indicators are presented using the same layout.

When the analyses is complete, Anders Kolstad, or someone else, will perform a rapid code review, making sure 
the analysis is repeatable by someone else. Be therefore very careful to annotate 
and comment on every step in the analysis. When completed, Anders will make a copy
of the working file and put it in the same project root folder, but removing the 'DRAFT_' part of the name. 
After modifying it slightly (mainly removing the top YAML header and updating the rmd_files argument in '_bookdown.yml') 
the output will become available on the website (https://ninanor.github.io/ecosystemCondition/). -->

<!-- Examples:

For an example, see 'indicators/breareal-workingFile.Rmd' 
and the subsequent 'breareal.Rmd' 
which again, when rendered, produces this:
https://ninanor.github.io/ecosystemCondition/areal-av-isbreer.html

Notice the degree of annotation.

-->

<!-- Replace 'indicator name' with yor own indicator name  -->
# Functional plant indicators (Moisture, Light, pH, Nitrogen)

<br />

_Author and date:_
Joachim Töpper

```{r}
Sys.Date()
```

<br />

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(sf)
library(tidyr)
library(dplyr)
library(plyr)
library(stringr)
library(tidyverse)
library(readxl)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Fill in which ecosystem the indicator belongs to, as well as the ecosystem characteristic it should be linked to. It's OK to use some Norwegian here -->
```{r, echo=F}
Ecosystem <- "våtmark"
Egenskap  <- c("Primærproduksjon","Abiotiske forhold")
ECT       <- "" # e.g. "Structural state characteristic"
Contact   <- "Joachim Töpper"
```

```{r, echo=F}
metaData <- data.frame(Ecosystem,
                       "Økologisk egenskap" = Egenskap,
                       "ECT class" = ECT)
knitr::kable(metaData)
```

<!-- Don't remove these three html lines -->
<br />
<br />
<hr />

<!-- Document you work below. Try not to change  the headers too much. Data can be stored on NINA server. Since the book is rendered on the R Server this works fine, but note that directory paths are different on the server compared to you local machine. If it is not too big you may store under /data/ on this repository -->

## Introduction
<!-- Text here -->

## About the underlying data
<!-- Text here -->

### Representativity in time and space
<!-- Text here -->

### Original units
<!-- Text here -->

### Temporal coverage
<!-- Text here -->

### Aditional comments about the dataset
<!-- Text here -->

## Ecosystem characteristic
### Norwegian standard
<!-- Text here -->

### <!-- Text here -->
<!-- Text here. Note that there are two parallell systems for defining ecosystem properties - one Norwegian and one international (SEEA EA). See man/SEEA EA examples.pdf for help -->

## Collinearities with other indicators
<!-- Text here -->

## Reference state and values
### Reference state
<!-- Text here -->

### Reference values, thresholds for defining _good ecological condition_, minimum and/or maximum values
<!-- Text here -->

## Uncertainties
<!-- Text here -->

## References
<!-- Text here -->

## Analyses
### Data sets

ANO data

```{r}
st_layers(dsn = "P:/41201785_okologisk_tilstand_2022_2023/data/Naturovervaking_eksport.gdb")
ANO.sp <- st_read("P:/41201785_okologisk_tilstand_2022_2023/data/Naturovervaking_eksport.gdb",
                   layer="ANO_Art")
ANO.geo <- st_read("P:/41201785_okologisk_tilstand_2022_2023/data/Naturovervaking_eksport.gdb",
                  layer="ANO_SurveyPoint")
head(ANO.sp)
head(ANO.geo)
```

Plant indicators from Tyler et al. 2021
```{r}
## Tyler indicator data
ind.dat <- read.table("P:/41201785_okologisk_tilstand_2022_2023/data/functional plant indicators/Tyler et al_Swedish plant indicators.txt",
                        sep = '\t', header=T, quote = '')
head(ind.datr)

```

Generalized species lists NiN
```{r}
load("P:/41201785_okologisk_tilstand_2022_2023/data/functional plant indicators//reference from NiN/Eco_State.RData")
str(Eco_State)
```

#### Plant indicator data - data handling
```{r}

names(ind.dat)
names(ind.dat)[1] <- 'species'
ind.dat$species <- as.factor(ind.dat$species)
summary(ind.dat$species)
ind.dat <- ind.dat[!is.na(ind.dat$species),] # removing species-NAs
ind.dat[,'species.orig'] <- ind.dat[,'species'] # make a backup of the original species variable
ind.dat[,'species'] <- word(ind.dat[,'species'], 1,2) # trimming away sub-species & co, and descriptor info

# the trimming above leaves some duplicates that need to be taken care of
ind.dat[duplicated(ind.dat[,'species']),"species"] ind.dat.dup <- ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat[ind.dat$species %in% ind.dat.dup,c("Light","Moisture","Soil_reaction_pH","Nitrogen","species.orig","species")]
ind.dat <- ind.dat %>% filter( !(species.orig %in% list("Ammophila arenaria x Calamagrostis epigejos",
                                                            "Anemone nemorosa x ranunculoides",
                                                            "Armeria maritima ssp. elongata",
                                                            "Asplenium trichomanes ssp. quadrivalens",
                                                            "Calystegia sepium ssp. spectabilis",
                                                            "Campanula glomerata 'Superba'",
                                                            "Dactylorhiza maculata ssp. fuchsii",
                                                            "Erigeron acris ssp. droebachensis",
                                                            "Erigeron acris ssp. politus",
                                                            "Erysimum cheiranthoides L. ssp. alatum",
                                                            "Euphrasia nemorosa x stricta var. brevipila",
                                                            "Galium mollugo x verum",
                                                            "Geum rivale x urbanum",
                                                            "Hylotelephium telephium (ssp. maximum)",
                                                            "Juncus alpinoarticulatus ssp. rariflorus",
                                                            "Lamiastrum galeobdolon ssp. argentatum",
                                                            "Lathyrus latifolius ssp. heterophyllus",
                                                            "Medicago sativa ssp. falcata",
                                                            "Medicago sativa ssp. x varia",
                                                            "Monotropa hypopitys ssp. hypophegea",
                                                            "Ononis spinosa ssp. hircina",
                                                            "Ononis spinosa ssp. procurrens",
                                                            "Pilosella aurantiaca ssp. decolorans",
                                                            "Pilosella aurantiaca ssp. dimorpha",
                                                            "Pilosella cymosa ssp. gotlandica",
                                                            "Pilosella cymosa ssp. praealta",
                                                            "Pilosella officinarum ssp. peleteranum",
                                                            "Poa x jemtlandica (Almq.) K. Richt.",
                                                            "Poa x herjedalica Harry Sm.",
                                                            "Ranunculus peltatus ssp. baudotii",
                                                            "Sagittaria natans x sagittifolia",
                                                            "Salix repens ssp. rosmarinifolia",
                                                            "Stellaria nemorum L. ssp. montana",
                                                            "Trichophorum cespitosum ssp. germanicum")
) )

# checking duplicates again
ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat.dup <- ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat[ind.dat$species %in% ind.dat.dup,c("Light","Moisture","Soil_reaction_pH","Nitrogen","species.orig","species")]
# getting rid of sect. for Hieracium
ind.dat <- ind.dat %>% mutate(species=gsub("sect. ","",species.orig))
ind.dat[,'species'] <- word(ind.dat[,'species'], 1,2)

ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat.dup <- ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat[ind.dat$species %in% ind.dat.dup,c("Light","Moisture","Soil_reaction_pH","Nitrogen","species.orig","species")]
# only hybrids left -> get rid of these
ind.dat <- ind.dat[!duplicated(ind.dat[,'species']),]
ind.dat[duplicated(ind.dat[,'species']),"species"]

ind.dat$species <- as.factor(ind.dat$species)
summary(ind.dat$species)
# no duplicates left

head(ind.dat)

```

#### ANO monitoring data - data handling
```{r}
head(ANO.sp)
head(ANO.geo)

## fix NiN information
ANO.geo$hovedtype_rute <- substr(ANO.geo$kartleggingsenhet_1m2,1,3) # take the 3 first characters
ANO.geo$hovedtype_rute <- gsub("-", "", ANO.geo$hovedtype_rute) # remove hyphen
unique(as.factor(ANO.geo$hovedtype_rute))

## fix NiN-variables
colnames(ANO.geo)
colnames(ANO.geo)[42:47] <- c("groeftingsintensitet",
                              "bruksintensitet",
                              "beitetrykk",
                              "slatteintensitet",
                              "tungekjoretoy",
                              "slitasje")
head(ANO.geo)

# remove variable code in the data
ANO.geo$groeftingsintensitet <- gsub("7GR-GI_", "", ANO.geo$groeftingsintensitet) 
unique(ANO.geo$groeftingsintensitet)
ANO.geo$groeftingsintensitet <- gsub("X", "NA", ANO.geo$groeftingsintensitet)
unique(ANO.geo$groeftingsintensitet)
ANO.geo$groeftingsintensitet <- as.numeric(ANO.geo$groeftingsintensitet)
unique(ANO.geo$groeftingsintensitet)

ANO.geo$bruksintensitet <- gsub("7JB-BA_", "", ANO.geo$bruksintensitet) 
unique(ANO.geo$bruksintensitet)
ANO.geo$bruksintensitet <- gsub("X", "NA", ANO.geo$bruksintensitet)
unique(ANO.geo$bruksintensitet)
ANO.geo$bruksintensitet <- as.numeric(ANO.geo$bruksintensitet)
unique(ANO.geo$bruksintensitet)

ANO.geo$beitetrykk <- gsub("7JB-BT_", "", ANO.geo$beitetrykk) 
unique(ANO.geo$beitetrykk)
ANO.geo$beitetrykk <- gsub("X", "NA", ANO.geo$beitetrykk)
unique(ANO.geo$beitetrykk)
ANO.geo$beitetrykk <- as.numeric(ANO.geo$beitetrykk)
unique(ANO.geo$beitetrykk)

ANO.geo$slatteintensitet <- gsub("7JB-SI_", "", ANO.geo$slatteintensitet) 
unique(ANO.geo$slatteintensitet)
ANO.geo$slatteintensitet <- gsub("X", "NA", ANO.geo$slatteintensitet)
unique(ANO.geo$slatteintensitet)
ANO.geo$slatteintensitet <- as.numeric(ANO.geo$slatteintensitet)
unique(ANO.geo$slatteintensitet)

ANO.geo$tungekjoretoy <- gsub("7TK_", "", ANO.geo$tungekjoretoy) 
unique(ANO.geo$tungekjoretoy)
ANO.geo$tungekjoretoy <- gsub("X", "NA", ANO.geo$tungekjoretoy)
unique(ANO.geo$tungekjoretoy)
ANO.geo$tungekjoretoy <- as.numeric(ANO.geo$tungekjoretoy)
unique(ANO.geo$tungekjoretoy)

ANO.geo$slitasje <- gsub("7SE_", "", ANO.geo$slitasje) 
unique(ANO.geo$slitasje)
ANO.geo$slitasje <- gsub("X", "NA", ANO.geo$slitasje)
unique(ANO.geo$slitasje)
ANO.geo$slitasje <- as.numeric(ANO.geo$slitasje)
unique(ANO.geo$slitasje)

## check that every point is present only once
length(levels(as.factor(ANO.geo$ano_flate_id)))
length(levels(as.factor(ANO.geo$ano_punkt_id)))
summary(as.factor(ANO.geo$ano_punkt_id))
# there's many double presences, probably some wrong registrations of point numbers,
# but also double registrations (e.g. ANO0159_55)
# CHECK THIS when preparing ecosystem-datasets for scaling

# fix species names
ANO.sp$Species <- ANO.sp$art_navn
unique(as.factor(ANO.sp$Species))
ANO.sp[,'Species'] <- word(ANO.sp[,'Species'], 1,2) # lose subspecies
ANO.sp$Species <- str_to_title(ANO.sp$Species) # make first letter capital
ANO.sp$Species <- gsub("( .*)","\\L\\1",ANO.sp$Species,perl=TRUE) # make capital letters after hyphen to lowercase
ANO.sp$Species <- gsub("( .*)","\\L\\1",ANO.sp$Species,perl=TRUE) # make capital letters after space to lowercase

## merge species data with indicators
ANO.sp.ind <- merge(x=ANO.sp[,c("Species", "art_dekning", "ParentGlobalID")], 
                y= ind.dat[,c("species", "Light", "Moisture", "Soil_reaction_pH", "Nitrogen")],
                by.x="Species", by.y="species", all.x=T)
summary(ANO.sp.ind)

## checking which species didn't find a match
unique(ANO.sp.ind[is.na(ANO.sp.ind$Moisture),'Species'])

# fix species name issues
ind.dat <- ind.dat %>% 
  mutate(species=str_replace(species,"Aconitum lycoctonum", "Aconitum septentrionale")) %>% 
  mutate(species=str_replace(species,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  mutate(species=str_replace(species,"Carex myosuroides", "Kobresia myosuroides")) %>%
  mutate(species=str_replace(species,"Clinopodium acinos", "Acinos arvensis")) %>%
  mutate(species=str_replace(species,"Artemisia rupestris", "Artemisia norvegica")) %>%
  mutate(species=str_replace(species,"Cherleria biflora", "Minuartia biflora"))

ANO.sp <- ANO.sp %>% 
  mutate(Species=str_replace(Species,"Arctous alpinus", "Arctous alpina")) %>%
  mutate(Species=str_replace(Species,"Betula tortuosa", "Betula pubescens")) %>%
  mutate(Species=str_replace(Species,"Blysmopsis rufa", "Blysmus rufus")) %>%
  mutate(Species=str_replace(Species,"Cardamine nymanii", "Cardamine pratensis")) %>%
  mutate(Species=str_replace(Species,"Carex adelostoma", "Carex buxbaumii")) %>%
  mutate(Species=str_replace(Species,"Carex leersii", "Carex echinata")) %>%
  mutate(Species=str_replace(Species,"Carex paupercula", "Carex magellanica")) %>%
  mutate(Species=str_replace(Species,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  mutate(Species=str_replace(Species,"Carex viridula", "Carex flava")) %>%
  mutate(Species=str_replace(Species,"Chamaepericlymenum suecicum", "Cornus suecia")) %>%
  mutate(Species=str_replace(Species,"Cicerbita alpina", "Lactuca alpina")) %>%
  mutate(Species=str_replace(Species,"Empetrum hermaphroditum", "Empetrum nigrum")) %>%
  mutate(Species=str_replace(Species,"Festuca prolifera", "Festuca rubra")) %>%
  mutate(Species=str_replace(Species,"Galium album", "Galium mollugo")) %>%
  mutate(Species=str_replace(Species,"Galium elongatum", "Galium palustre")) %>%
  mutate(Species=str_replace(Species,"Helictotrichon pratense", "Avenula pratensis")) %>%
  mutate(Species=str_replace(Species,"Helictotrichon pubescens", "Avenula pubescens")) %>%
  mutate(Species=str_replace(Species,"Hieracium alpina", "Hieracium Alpina")) %>%
  mutate(Species=str_replace(Species,"Hieracium alpinum", "Hieracium Alpina")) %>%
  mutate(Species=str_replace(Species,"Hieracium hieracium", "Hieracium Hieracium")) %>%
  mutate(Species=str_replace(Species,"Hieracium hieracioides", "Hieracium umbellatum")) %>%
  mutate(Species=str_replace(Species,"Hieracium murorum", "Hieracium Vulgata")) %>%
  mutate(Species=str_replace(Species,"Hieracium oreadea", "Hieracium Oreadea")) %>%
  mutate(Species=str_replace(Species,"Hieracium prenanthoidea", "Hieracium Prenanthoidea")) %>%
  mutate(Species=str_replace(Species,"Hieracium vulgata", "Hieracium Vulgata")) %>%
  mutate(Species=str_replace(Species,"Hieracium pilosella", "Pilosella officinarum")) %>%
  mutate(Species=str_replace(Species,"Hieracium vulgatum", "Hieracium umbellatum")) %>%
  mutate(Species=str_replace(Species,"Hierochloã« alpina", "Hierochloë alpina")) %>%
  mutate(Species=str_replace(Species,"Hierochloã« hirta", "Hierochloë hirta")) %>%
  mutate(Species=str_replace(Species,"Hierochloã« odorata", "Hierochloë odorata")) %>%
  mutate(Species=str_replace(Species,"Listera cordata", "Neottia cordata")) %>%
  mutate(Species=str_replace(Species,"Leontodon autumnalis", "Scorzoneroides autumnalis")) %>%
  mutate(Species=str_replace(Species,"Loiseleuria procumbens", "Kalmia procumbens")) %>%
  mutate(Species=str_replace(Species,"Mycelis muralis", "Lactuca muralis")) %>%
  mutate(Species=str_replace(Species,"Omalotheca supina", "Gnaphalium supinum")) %>%
  mutate(Species=str_replace(Species,"Omalotheca norvegica", "Gnaphalium norvegicum")) %>%
  mutate(Species=str_replace(Species,"Omalotheca sylvatica", "Gnaphalium sylvaticum")) %>%
  mutate(Species=str_replace(Species,"Oreopteris limbosperma", "Thelypteris limbosperma")) %>%
  mutate(Species=str_replace(Species,"Oxycoccus microcarpus", "Vaccinium microcarpum")) %>%
  mutate(Species=str_replace(Species,"Oxycoccus palustris", "Vaccinium oxycoccos")) %>%
  mutate(Species=str_replace(Species,"Phalaris minor", "Phalaris arundinacea")) %>%
  mutate(Species=str_replace(Species,"Pinus unicinata", "Pinus mugo")) %>%
  mutate(Species=str_replace(Species,"Poa alpigena", "Poa pratensis")) %>%
  mutate(Species=str_replace(Species,"Poa angustifolia", "Poa pratensis")) %>%
  mutate(Species=str_replace(Species,"Pyrola grandiflora", "Pyrola rotundifolia")) %>%
  mutate(Species=str_replace(Species,"Rumex alpestris", "Rumex acetosa")) %>%
  mutate(Species=str_replace(Species,"Syringa emodi", "Syringa vulgaris")) %>%
  mutate(Species=str_replace(Species,"Taraxacum crocea", "Taraxacum officinale")) %>%
  mutate(Species=str_replace(Species,"Taraxacum croceum", "Taraxacum officinale")) %>%
  mutate(Species=str_replace(Species,"Trientalis europaea", "Lysimachia europaea")) %>%
  mutate(Species=str_replace(Species,"Trifolium pallidum", "Trifolium pratense"))

## merge species data with indicators
ANO.sp.ind <- merge(x=ANO.sp[,c("Species", "art_dekning", "ParentGlobalID")], 
                    y= ind.dat[,c("species", "Light", "Moisture", "Soil_reaction_pH", "Nitrogen")],
                    by.x="Species", by.y="species", all.x=T)
summary(ANO.sp.ind)
# checking which species didn't find a match
unique(ANO.sp.ind[is.na(ANO.sp.ind$Moisture),'Species'])
# don't find synonyms for these in the ind lists

## trimming away the points without information on NiN, species or cover
ANO.sp.ind <- ANO.sp.ind[!is.na(ANO.sp.ind$Species),]
ANO.sp.ind <- ANO.sp.ind[!is.na(ANO.sp.ind$art_dekning),]

summary(ANO.sp.ind)
head(ANO.sp.ind)
rm(ANO.sp)

```

#### NiN reference data - data handling

```{r}

## generalized species lists from NiN
str(Eco_State)

# species
Eco_State$Concept_Data$Species$Species_List$species
# environments
t(Eco_State$Concept_Data$Env$Env_Data)
# abundances
t(Eco_State$Concept_Data$Species$Species_Data)

## transposing abundance data
NiN.sp <- t(Eco_State$Concept_Data$Species$Species_Data)
NiN.sp <- as.data.frame(NiN.sp)
NiN.sp$sp <- as.factor(as.vector(Eco_State$Concept_Data$Species$Species_List$species))
# only genus and species name
NiN.sp$sp <- word(NiN.sp$sp, 1,2)
NiN.sp$spgr <- as.factor(as.vector(Eco_State$Concept_Data$Species$Species_List$art.code))
# if relevant, trimming to desired species groups (for forests e.g. removing trees)
#NiN.sp <- NiN.sp[NiN.sp$spgr!="a1a",]

## environment data
NiN.env <- Eco_State$Concept_Data$Env$Env_Data

## merging with indicator values
NiN.sp.ind <- merge(NiN.sp,ind.dat, by.x="sp", by.y="species", all.x=T)
summary(NiN.sp.ind)

NiN.sp.ind[NiN.sp.ind==999] <- NA

## checking which species didn't find a match
unique(NiN.sp.ind[is.na(NiN.sp.ind$Moisture) & NiN.sp.ind$spgr %in% list("a1a","a1b","a1c"),'sp'])

## fix species name issues
ind.dat <- ind.dat %>% 
  mutate(species=str_replace(species,"Aconitum lycoctonum", "Aconitum septentrionale")) %>% 
  mutate(species=str_replace(species,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  mutate(species=str_replace(species,"Carex myosuroides", "Kobresia myosuroides")) %>%
  mutate(species=str_replace(species,"Clinopodium acinos", "Acinos arvensis")) %>%
  mutate(species=str_replace(species,"Artemisia rupestris", "Artemisia norvegica")) %>%
  mutate(species=str_replace(species,"Cherleria biflora", "Minuartia biflora"))

NiN.sp <- NiN.sp %>% 
  mutate(sp=str_replace(sp,"Aconitum lycoctonum", "Aconitum septentrionale")) %>% 
  mutate(sp=str_replace(sp,"Anagallis minima", "Lysimachia minima")) %>% 
  mutate(sp=str_replace(sp,"Arctous alpinus", "Arctous alpina")) %>%
  mutate(sp=str_replace(sp,"Betula tortuosa", "Betula pubescens")) %>%
  mutate(sp=str_replace(sp,"Blysmopsis rufa", "Blysmus rufus")) %>%
  mutate(sp=str_replace(sp,"Cardamine nymanii", "Cardamine pratensis")) %>%
  mutate(sp=str_replace(sp,"Carex adelostoma", "Carex buxbaumii")) %>%
  mutate(sp=str_replace(sp,"Carex leersii", "Carex echinata")) %>%
  mutate(sp=str_replace(sp,"Carex paupercula", "Carex magellanica")) %>%
  mutate(sp=str_replace(sp,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  mutate(sp=str_replace(sp,"Carex _vacillans", "Carex vacillans")) %>%
  mutate(sp=str_replace(sp,"Carex viridula", "Carex flava")) %>%
  mutate(sp=str_replace(sp,"Chamaepericlymenum suecicum", "Cornus suecia")) %>%
  mutate(sp=str_replace(sp,"Cornus suecia", "Cornus suecica")) %>%
  mutate(sp=str_replace(sp,"Cicerbita alpina", "Lactuca alpina")) %>%
  mutate(sp=str_replace(sp,"Dactylorhiza sphagnicola", "Dactylorhiza majalis")) %>%
  mutate(sp=str_replace(sp,"Diphasiastrum complanatum", "Lycopodium complanatum")) %>%
  mutate(sp=str_replace(sp,"Elymus alaskanus", "Elymus kronokensis")) %>%
  mutate(sp=str_replace(sp,"Empetrum hermaphroditum", "Empetrum nigrum")) %>%
  mutate(sp=str_replace(sp,"Erigeron eriocephalus", "Erigeron uniflorus")) %>%
  mutate(sp=str_replace(sp,"Festuca prolifera", "Festuca rubra")) %>%
  mutate(sp=str_replace(sp,"Galium album", "Galium mollugo")) %>%
  mutate(sp=str_replace(sp,"Galium elongatum", "Galium palustre")) %>%
  mutate(sp=str_replace(sp,"Glaux maritima", "Lysimachia maritima")) %>%
  mutate(sp=str_replace(sp,"Helictotrichon pratense", "Avenula pratensis")) %>%
  mutate(sp=str_replace(sp,"Helictotrichon pubescens", "Avenula pubescens")) %>%
  mutate(sp=str_replace(sp,"Hieracium alpina", "Hieracium Alpina")) %>%
  mutate(sp=str_replace(sp,"Hieracium alpinum", "Hieracium Alpina")) %>%
  mutate(sp=str_replace(sp,"Hieracium aurantiacum", "Pilosella aurantiaca")) %>%
  mutate(sp=str_replace(sp,"Hieracium hieracium", "Hieracium Hieracium")) %>%
  mutate(sp=str_replace(sp,"Hieracium hieracioides", "Hieracium umbellatum")) %>%
  mutate(sp=str_replace(sp,"Hieracium lactucella", "Pilosella lactucella")) %>%
  mutate(sp=str_replace(sp,"Hieracium murorum", "Hieracium Vulgata")) %>%
  mutate(sp=str_replace(sp,"Hieracium oreadea", "Hieracium Oreadea")) %>%
  mutate(sp=str_replace(sp,"Hieracium prenanthoidea", "Hieracium Prenanthoidea")) %>%
  mutate(sp=str_replace(sp,"Hieracium vulgata", "Hieracium Vulgata")) %>%
  mutate(sp=str_replace(sp,"Hieracium pilosella", "Pilosella officinarum")) %>%
  mutate(sp=str_replace(sp,"Hieracium vulgatum", "Hieracium umbellatum")) %>%
  mutate(sp=str_replace(sp,"Hierochloã« alpina", "Hierochloë alpina")) %>%
  mutate(sp=str_replace(sp,"Hierochloã« hirta", "Hierochloë hirta")) %>%
  mutate(sp=str_replace(sp,"Hierochloã« odorata", "Hierochloë odorata")) %>%
  mutate(sp=str_replace(sp,"Huperzia appressa", "Huperzia selago")) %>%
  mutate(sp=str_replace(sp,"Hylotelephium maximum", "Hylotelephium telephium")) %>%
  mutate(sp=str_replace(sp,"Lappula myosotis", "Lappula squarrosa")) %>%
  mutate(sp=str_replace(sp,"Lepidotheca suaveolens", "Matricaria discoidea")) %>%
  mutate(sp=str_replace(sp,"Listera cordata", "Neottia cordata")) %>%
  mutate(sp=str_replace(sp,"Leontodon autumnalis", "Scorzoneroides autumnalis")) %>%
  mutate(sp=str_replace(sp,"Loiseleuria procumbens", "Kalmia procumbens")) %>%
  mutate(sp=str_replace(sp,"Logfia arvensis", "Filago arvensis")) %>%
  mutate(sp=str_replace(sp,"Mentha _verticillata", "Mentha verticillata")) %>%
  mutate(sp=str_replace(sp,"Minuartia rubella", "Sabulina rubella")) %>%
  mutate(sp=str_replace(sp,"Minuartia stricta", "Sabulina stricta")) %>%
  mutate(sp=str_replace(sp,"Mycelis muralis", "Lactuca muralis")) %>%
  mutate(sp=str_replace(sp,"Omalotheca supina", "Gnaphalium supinum")) %>%
  mutate(sp=str_replace(sp,"Omalotheca norvegica", "Gnaphalium norvegicum")) %>%
  mutate(sp=str_replace(sp,"Omalotheca sylvatica", "Gnaphalium sylvaticum")) %>%
  mutate(sp=str_replace(sp,"Ononis arvensis", "Ononis spinosa")) %>%
  mutate(sp=str_replace(sp,"Oreopteris limbosperma", "Thelypteris limbosperma")) %>%
  mutate(sp=str_replace(sp,"Oxycoccus microcarpus", "Vaccinium microcarpum")) %>%
  mutate(sp=str_replace(sp,"Oxycoccus palustris", "Vaccinium oxycoccos")) %>%
  mutate(sp=str_replace(sp,"Phalaris minor", "Phalaris arundinacea")) %>%
  mutate(sp=str_replace(sp,"Phalaroides arundinacea", "Phalaris arundinacea")) %>%
  mutate(sp=str_replace(sp,"Pinus unicinata", "Pinus mugo")) %>%
  mutate(sp=str_replace(sp,"Platanthera montana", "Platanthera chlorantha")) %>%
  mutate(sp=str_replace(sp,"Poa alpigena", "Poa pratensis")) %>%
  mutate(sp=str_replace(sp,"Poa angustifolia", "Poa pratensis")) %>%
  mutate(sp=str_replace(sp,"Poa laxa", "Poa flexuosa")) %>%
  mutate(sp=str_replace(sp,"Poa _herjedalica", "Poa herjedalica")) %>%
  mutate(sp=str_replace(sp,"Poa _jemtlandica", "Poa jemtlandica")) %>%
  mutate(sp=str_replace(sp,"Poa lindebergii", "Poa arctica")) %>%
  mutate(sp=str_replace(sp,"Pyrola grandiflora", "Pyrola rotundifolia")) %>%
  mutate(sp=str_replace(sp,"Rhamnus catharticus", "Rhamnus cathartica")) %>%
  mutate(sp=str_replace(sp,"Rumex alpestris", "Rumex acetosa")) %>%
  mutate(sp=str_replace(sp,"Salix _fragilis", "Salix fragilis")) %>%
  mutate(sp=str_replace(sp,"Saxifraga _opdalensis", "Saxifraga opdalensis")) %>%
  mutate(sp=str_replace(sp,"Spergularia salina", "Spergularia marina")) %>%
  mutate(sp=str_replace(sp,"Syringa emodi", "Syringa vulgaris")) %>%
  mutate(sp=str_replace(sp,"Taraxacum crocea", "Taraxacum officinale")) %>%
  mutate(sp=str_replace(sp,"Taraxacum croceum", "Taraxacum officinale")) %>%
  mutate(sp=str_replace(sp,"Taraxacum erythrospermum", "Taraxacum officinale")) %>%
  mutate(sp=str_replace(sp,"Taraxacum hamatum", "Taraxacum officinale")) %>%
  mutate(sp=str_replace(sp,"Trientalis europaea", "Lysimachia europaea")) %>%
  mutate(sp=str_replace(sp,"Trifolium pallidum", "Trifolium pratense")) %>%
  mutate(sp=str_replace(sp,"Vicia orobus", "Vicia cassubica"))

## merge species data with indicators
NiN.sp.ind <- merge(NiN.sp,ind.dat, by.x="sp", by.y="species", all.x=T)
summary(NiN.sp.ind)

NiN.sp.ind[NiN.sp.ind==999] <- NA

# checking which species didn't find a match
unique(NiN.sp.ind[is.na(NiN.sp.ind$Moisture) & NiN.sp.ind$spgr %in% list("a1a","a1b","a1c"),'sp'])
# ok now

#### CONTINUE HERE ####

## matching with NiN ecosystem types
# NB! beware of rogue spaces in the 'Nature_type' & 'Sub_Type' variables, e.g. "Spring_Forest "
NiN.wetland <- NiN.sp.ind[,c("sp",paste(NiN.env[NiN.env$Nature_Type=="Mire","ID"]),colnames(ind.dat)[19:23])]   # Light, Moisture, Soil_reaction_pH, Nitrogen
NiN.wetland[1,]
names(NiN.wetland)

#NiN.wetland <- NiN.wetland[,c(1:41,46:48,59:69)]
cbind(colnames(NiN.wetland),
      c("",
        'V3-C1a','V3-C1b','V3-C1c','V3-C1d','V3-C1e',
        'V1-C1a','V1-C1b','V1-C1c','V1-C1d','V1-C1e',
        'V1-C2a','V1-C2b','V1-C2c','V1-C2d',
        'V1-C3a','V1-C3b','V1-C3c','V1-C3d',
        'V1-C4a','V1-C4b','V1-C4c','V1-C4d',
        'V1-C4e','V1-C4f','V1-C4g','V1-C4h',
        'V3-C2','V1-C5',
        'V1-C6a','V1-C6b',
        'V1-C7a','V1-C7b',
        'V1-C8a','V1-C8b',
        'V2-C1a','V2-C1b',
        'V2-C2a','V2-C2b',
        'V2-C3a','V2-C3b',
        "V4-C4","V4-C5",
        "V4-C2","V4-C3",
        'V8-C1','V8-C2','V8-C3',
        rep("",10),
        rep("",5)
      )
)

NiN.wetland <- NiN.wetland[,c(1:43,46:48,59:63)]
colnames(NiN.wetland)[2:46] <- c('V3-C1a','V3-C1b','V3-C1c','V3-C1d','V3-C1e',
                              'V1-C1a','V1-C1b','V1-C1c','V1-C1d','V1-C1e',
                              'V1-C2a','V1-C2b','V1-C2c','V1-C2d',
                              'V1-C3a','V1-C3b','V1-C3c','V1-C3d',
                              'V1-C4a','V1-C4b','V1-C4c','V1-C4d',
                              'V1-C4e','V1-C4f','V1-C4g','V1-C4h',
                              'V3-C2','V1-C5',
                              'V1-C6a','V1-C6b',
                              'V1-C7a','V1-C7b',
                              'V1-C8a','V1-C8b',
                              'V2-C1a','V2-C1b',
                              'V2-C2a','V2-C2b',
                              'V2-C3a','V2-C3b',
                              'V4-C2','V4-C3',
                              'V8-C1','V8-C2','V8-C3'
)
head(NiN.wetland)

# translating the abundance classes into %-cover
coverscale <- data.frame(orig=0:6,
                         cov=c(0,1/32,1/8,3/8,0.6,4/5,1)
)

NiN.wetland.cov <- NiN.wetland
colnames(NiN.wetland.cov)
for (i in 2:46) {
  NiN.wetland.cov[,i] <- coverscale[,2][ match(NiN.wetland[,i], 0:6 ) ]
}

summary(NiN.wetland)
summary(NiN.wetland.cov)

```

#### Regions
<!-- In case you need to map the indicator value to regions, you can do that here. Remove this chapter if not relevant. -->
Importing a shape file with the regional delineation. 
```{r}
#reg <- sf::st_read("data/regions.shp")
#st_crs(reg)
```

### Scaled indicator values
<!-- Text and analyses here -->

### Uncertainty
<!-- Text here -->

## Prepare export
<!-- Text here -->

### Eksport file (final product)
<!-- Export final file. Ideally a georeferenced shape or raster wit indicators values (raw and normalised), reference values and errors. -->