### 2023-03-20 DRAFT_alien_species.Rmd
start to validate

---
title: "Alien species"
output:
  html_document
---
# Alien species {#alien-species}

*Author and date:*

Anders L. Kolstad

March 2023

```{r setup, include=FALSE}
library(knitr)
library(sf)
library(tmap)
library(ggpubr)
library(DT)
library(units)
library(eaTools)
#library(tidyverse)
library(tibble)
library(ggplot2)
library(tidyr)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)

# Set up conditional file paths
dir <- substr(getwd(), 1,2)

path <- ifelse(dir == "C:", 
               "R:/GeoSpatialData/Habitats_biotopes/Norway_Miljodirektoratet_Naturtyper_nin/Original/Naturtyper_nin_0000_norge_25833_FILEGDB/Naturtyper_nin_0000_norge_25833_FILEGDB.gdb",
               "/data/R/GeoSpatialData/Habitats_biotopes/Norway_Miljodirektoratet_Naturtyper_nin/Original/Naturtyper_nin_0000_norge_25833_FILEGDB/Naturtyper_nin_0000_norge_25833_FILEGDB.gdb")

pData <- ifelse(dir == "C:", 
               "P:/41201785_okologisk_tilstand_2022_2023/data/",
               "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/")

```

```{r, echo=F}
Ecosystem <- "Våtmark, Naturlig åpne områder under skoggrensa, Semi-naturlig mark" 
Egenskap  <- "Primærproduksjon (evt. Abiotiske fohold)"
ECT       <- "Physical state characteristics" 
Contact   <- "anders.kolstad@nina.no"
```

```{r, echo=F}
metaData <- data.frame(Ecosystem,
                       "Økologisk egenskap" = Egenskap,
                       "ECT class" = ECT)
knitr::kable(metaData)
```

<hr/>

## Introduction
This indicator represent the the ecological on-site effect from  alien species. Whilst the introduction of alien species is best considered a pressure variable, the physical abundance or relative abundance of alien species on a given location is seen here as a manifestation of this negative anthropogenic impact, and is therefore a good condition indicator.

A related variable has been used previously for [alpine](https://ninanor.github.io/IBECA/faktaark.html#Frav%C3%A6r_av_fremmede_arter) and forest ecosystems in Norway. These indicators were based on ANO, a national monitoring program, where alien species cover was estimated to the closest whole %. The upper and lower reference levels that was used back then was 0 or 100% of the species composition made up of alien species, respectively. The threshold value between good and reduced condition was defined as 5%. In this approach, the indicator represented the relative area with an absence of alien species. It is in a way the flip-side of _alien species abundance_. The latter was said to be a pressure variable, whilst the former reacts as direct and linear consequence of changes to this variable, and was therefore accepted as a condition indicator. We do not make this distinction here.

This time we are using the [nature type mapping dataset](#naturtype) which records the _effect_ of alien species,, rather than the % cover. The recording scale is called R7 (Fig. \@ref(fig:R7)) and is an ordinal scale which has a non-linear relationship with %cover recorded in ANO (Fig. \@ref(fig:fremmedart-prosentVSinnslag)). In order to combine these two data sets, we must convert one of these scales to match the other.

Here I base the indicator on the R7 scale, since this is, presumably, the bigger data set. I therefore convert %cover into R7 classes based on what I can interpret from Fig. \@ref(fig:fremmedart-prosentVSinnslag).

Upper reference value is then 1, and lower reference value is 7, on the R7 scale.
For the threshold value I will set this to value 4 on the R7 scale. 
This is based on interpretation by the author and based of the definition of this value in Fig. \@ref(fig:R7). 
Step 4 on the R7 scale is the step that, concordantly, best matches the 5% threshold in the old approach, as seen from Fig. \@ref(fig:fremmedart-prosentVSinnslag).

```{r R7, fig.cap="The R7 scale used when recording the effect of alien species in the nature type monitoring program"}
knitr::include_graphics("images/R7 scale.PNG")
```

```{r fremmedart-prosentVSinnslag, fig.cap="Comparing the R7 scale (x axis) which is used in the nature type monitoring program with %cover (y-axis), which is how alien species is recorded in ANO. I am assuming the x axis is shifted downwards compared to the original R7 scale which range from 1 to 7, so that a value of 0 here exualrs to an R7 value of 1. Cropped from Evju et al (2022)."}
knitr::include_graphics("images/fremmedart_prosentVSinnslag.PNG")
```

Since the nature type data is not sampled in a systematic or random way, we must take extra care not to over-extrapolate in space. We delineate _homogeneous impact areas_ (HIA) based on four classes of increasing infrastructure, and we say that the field data is representative inside the HIA and region. In [a related indicator](#slitasje-og-kjørespor) we used the accounting areas as regions (5 accounting areas in Norway). For this indicator, we expect more gradients also inside accounting areas, and therefore we define areas as municipalities. This means that there will be more missing data in our final indicator maps, since not all municipalities will have sufficient data points to estimate indicator values.

We then calculate an area weighted mean (and error) indicator value for each HIA and municipality combination, as long as there is more than _n_ data points for a given combination of HIA and municipality.

Here a general workflow for the calculation of the indicator.

1.  Import [Nature type data](#NTM2) data set (incl. GRUK) and [ANO data
    set](#ANO-import2)

2.  Identify the [relevant](#naturtype) nature types and [subset](#NTM)
    the data

3.  Convert [ANO points to polygons](#ANO-points-to-poly)

5.  [Combine](#combine-nt-ano) data sets

6.  [Scale](#scale-alien-ind) the `alien species` variable based on
    reference values.

7.  Define [homogeneous impact areas](#HIA) (HIA) based on an infrastructure index and municipalities

8.  [Aggregate and spread](#spread-slitasje) indicator value across HIAs
    (and municipalities)

9.  Confirm relationship between infrastructure index and indicator
    values to justify the extrapolation

10. TO DO: Prepare ecosystem delineation maps and us ethese to mask the extrapolated indicator maps

11. Spatial aggregation of indicator values and uncertainties to accounting areas

12. Export indicator maps and regional extrapolated maps

## About the underlying data

The indicator uses a data set from a standardised field survey of nature
types. You can read more about the preliminary analyses
[here](#naturtype). See also the [official
site](https://www.miljodirektoratet.no/ansvarsomrader/overvaking-arealplanlegging/naturkartlegging/naturtyper/)
of the Environment Agency. I also import a data set called
[ANO](#ANO-import2), which you can read about
[here](https://www.miljodirektoratet.no/ansvarsomrader/overvaking-arealplanlegging/miljoovervaking/overvakingsprogrammer/natur-pa-land/arealrepresentativ-naturovervakning-ano/)

### Representativity in time and space

The nature type mapping is not random and cannot be said to be area
representative. The ANO data set however, is area representative. The
data is from 2018 to present. The data from one field season usually
becomes available early the following year.

### Original units

The variables are recorded on a seven-step ordinal scale (Fig.
\@ref(fig:R7)) or a %cover of alien species.

### Temporal coverage

The data goes back to 2018. I therefore bulk all the data from 2018 to 2022 into one time step. I then use the mean date for the raw data,
and define the variable as belonging to the year 2020 (read more
[here](#scaled-alien-variable)).

### Aditional comments about the dataset

For a run through of the nature type data set, see [here](#naturtype).

## Ecosystem characteristic

### Norwegain standard

The indicator is tagged to the *Økologisk egenskap* called
**Funksjonelt viktige arter og strukturer** (Functionally important species and structures).

### UN standard

The indicator is tagged as a **B1 Compositional state characteristics**
indicator.

## Collinearities with other indicators

_Alien species_ is not thought to exhibit collinearity with any other indicator at
the present.

## Reference condition and values

### Reference condition

The reference condition is one with minimal negative human impact. This
is also true for semi-natural ecosystems. In a sense, the reference condition is pre 1800s,
since alien species are defined as species established post 1800.

### Reference values, thresholds for defining *good ecological condition*, minimum and/or maximum values

* Upper = 1 (R7 scale), corresponding to 0% alien species

* Threshold = 4 (R7 scale), corresponding to about 5-30% alien species cover.

* Lower = 7 (R7 scale), corresponding to 100% alien species cover.

Read about the normalisation [here](##scaled-alien-variable).

## Uncertainties

Uncertainties/errors are estimated for aggregated indicator values by
bootstrapping individual indicator values 1000 times and calculating a
distribution of area weighted means. This uncertainty is different from 
the spatial variation which we could get more straight forward without 
bootstrapping. When aggregating a second time,
from homogeneous impact areas to accounting areas, we assume a normal
distribution around the indicator values, with the already mentioned
errors, and sample _n_ times from these and combine the resamples into an
a new, area weighted, distribution. The errors for the accounting areas
thus represents both the spatial variation, and the precision, of the
indicator values within the accounting areas.

## References

[Nature type data](https://www.miljodirektoratet.no/ansvarsomrader/overvaking-arealplanlegging/naturkartlegging/naturtyper/)

[ANO](https://www.miljodirektoratet.no/ansvarsomrader/overvaking-arealplanlegging/miljoovervaking/overvakingsprogrammer/natur-pa-land/arealrepresentativ-naturovervakning-ano/)

Evju, M., Skrindo, A.B. & Solstad, H. (red.) 2022. Overvåking av åpen grunnlendt kalkmark 2021‒2024. Årsrapport 2022. NINA Rapport 2195. Norsk institutt for naturforskning.

## Analyses

### Data sets

#### Nature type mapping {#NTM2}

This indicator uses the data set
`Naturtyper etter Miljødirektoratets Instruks`, which can be found
[here](https://kartkatalog.geonorge.no/metadata/naturtyper-miljoedirektoratets-instruks/eb48dd19-03da-41e1-afd9-7ebc3079265c).
See also [here](#naturtype) for a detailed description of the data set.

We also have a separate [summary file](##exp-natureType-summary) where
the nature types are manually mapped the NiN variables and to the correct NiN-main types. 
We can use this to find the nature types associated with the NiN main types
that we are interested in.

```{r}
naturetypes_summary_import <- readRDS("data/naturetypes/natureType_summary.rds")
```

We are only interested in a mapping units that include our target variables. The variable _fremmedartsinnslag_ is coded as 7FA.

```{r}
myVars <- '7FA'
```

```{r}
naturetypes_summary <- naturetypes_summary_import %>%
  rowwise() %>%
  mutate(keepers = sum(c_across(
    all_of(myVars))>0, na.rm=T)) %>%
  filter(keepers >0) %>%
  select(Nature_type, NiN_mainType, Year)
```

This deleted `r nrow(naturetypes_summary_import) - nrow(naturetypes_summary)` nature types and left us with these `r nrow(naturetypes_summary)` nature types where 7FA was recorded:

```{r}
DT::datatable(naturetypes_summary)
```

Importing and sub-setting the main data file, fix duplicate _hovedøkosystem_, calculate area, split one column in two, make numeric, and select target variables:
```{r, message=F}
naturetypes <- sf::st_read(dsn = path) %>%
  filter(naturtype %in% naturetypes_summary$Nature_type) %>%
  mutate(hovedøkosystem = case_match(hovedøkosystem,
                                  "naturligÅpneOmråderILavlandet" ~ "naturligÅpneOmråderUnderSkoggrensa",
                                  .default = hovedøkosystem),
         area = st_area(.)) %>%
  separate_rows(ninBeskrivelsesvariable, sep=",") %>%
  separate(col = ninBeskrivelsesvariable,
           into = c("NiN_variable_code", "NiN_variable_value"),
           sep = "_",
           remove=F) %>%
  mutate(NiN_variable_value = as.numeric(NiN_variable_value)) %>%
  filter(NiN_variable_code %in% myVars)
```

```{r fig.cap="An overview of the naturetypes for which we will calculate the indicator.", fig.height=8, fig.width=6}
ggplot(data = naturetypes, aes(x = naturtype, fill = hovedøkosystem))+
  geom_bar()+
  coord_flip()+
  theme_bw(base_size = 10)+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.direction = "vertical")+
  guides(fill = "none")+
  xlab("")+
  ylab("Number of localities")+
  facet_wrap(.~hovedøkosystem, ncol=1, scale="free")
```

Column names starting with a number is problematic, so adding a prefix

```{r}
naturetypes$NiN_variable_code <- paste0("var_", naturetypes$NiN_variable_code)
```

Removing NA's

```{r}
naturetypes <- naturetypes[!is.na(naturetypes$NiN_variable_value),]
```

Shift variable to start from 1 rather than from 0.

```{r}
naturetypes$NiN_variable_value <- naturetypes$NiN_variable_value+1
```

```{r alien-naturetype, fig.cap="Alien species variable scores."}
ggplot(naturetypes, aes(x = NiN_variable_value))+
  geom_histogram(fill="grey",
           colour = "black",
           binwidth = 1)+
  theme_bw(base_size = 12)+
  labs(x = "Alien species variable score") +
  facet_wrap(.~hovedøkosystem, scales="free_y")
```

It appears most localities are in good condition, and most localities have no alien species at all.

#### GRUK

This variable is also recorded in
[GRUK](https://www.nina.no/Naturmangfold/Trua-natur/%C3%85pen-grunnlendt-kalkmark).
The nature type data set I'm working on here includes this data already
(presently only 2021 included). GRUK also records a related variable:
`% cover in 5m radii circles`, which is much more detailed. This data is
not published. In any case it is better to use the harmonized data set
in our case.

_NOTE: I'm not sure, but I wonder if the 2021 GRUK data was a pilot project and that most GRUK data, and all futute GRUK data, will not be included in this data set._

#### ANO {#ANO-import2}

Arealrepresentativ Naturovervåking (ANO) consist of 1000 systematically
placed locations each with 18 sample points. In each sample point a
circle of 250 m2 is visualised, and the main ecosystem is recorded.

The %cover of alien vascular plant is estimated, independent on main ecosystem type. 
In contrast to the nature type data set, ANO only looks at vascular plants, and then also only the three strictest categories
in the alien plant list (SE, HI, PH). It is therefore not completely unproblematic to combine these to data sets,
but we will do so anyway. We expect plants to be the by far most common species group to influence the 7FA variable as well.
And similarly, we expect the three strictest  alien plant categories to contain the most aggressive species which will
drive the 7FA variable more so than the other categories.

```{r, message=F}
ano_eco <- c("vaatmark", "naturlig_apne", "semi_naturlig_mark")
ano <- sf::st_read(paste0(pData, "Naturovervaking_eksport.gdb"),
                   layer = "ANO_SurveyPoint") %>%
  dplyr::filter(hovedoekosystem_punkt %in% ano_eco) %>%
   mutate(hovedoekosystem_punkt = case_match(hovedoekosystem_punkt,
                                       "vaatmark" ~ "våtmark",
                                       "naturlig_apne" ~ "naturligÅpneOmråderUnderSkoggrensa",
                                       "semi_naturlig_mark" ~ "semi-naturligMark",
                                       .default = hovedoekosystem_punkt))
```

```{r}
table(ano$aar)
```

This data set only contains data from year 2019 and 2021. We need to
update this data set later. It is not clear why there is no data from
2020.

Each point/row here is 250 square meters. The data also contains
information about how big a proportion of this area is made up of the
dominant main ecosystem. However, there are
`r (NAS <- length(ano$andel_hovedoekosystem_punkt[is.na(ano$andel_hovedoekosystem_punkt)]))`
NA's here, which is `r round(NAS/nrow(ano), 2)*100`% of the data.

It appears the proportion of each circle that is made up of the dominant
ecosystem was only recorded after year 2019. In fact, the main ecosystem
was not recorded at all in 2019:
```{r}
# Hard coding name change
#unique(ano$hovedtype_250m2)[3]
#ano$hovedtype_250m2[ano$hovedtype_250m2==unique(ano$hovedtype_250m2)[3]] <- "Aapen jordvassmyr"
```

```{r}
table(ano$hovedtype_250m2, ano$aar)
```

I can remove the NA's, and thus the 2019 data.

> *All of 2019 ANO data is excluded because of missing information*

```{r}
ano <- ano[!is.na(ano$andel_hovedoekosystem_punkt),]
```

Let's look at the variation in the recorded proportion of ecosystem
cover

```{r, fig.cap="Distribution of the ANO variable andel_hovedoekosystem_punkt for wetland localities."}
par(mar=c(5,6,4,2))
plot(ano$andel_hovedoekosystem_punkt[order(ano$andel_hovedoekosystem_punkt)],
     ylab="Percentage of the 250 m2 area\ncovered by the main ecosystem")
```

The zero in there is an obvious mistake.

```{r}
ano <- ano[ano$andel_hovedoekosystem_punkt>20,]
```

Here's another plot of the distribution of the same variable:

```{r, warning=F, fig.cap="Percentage cover of the Våtmark main ecosystem in the 250m2 circle"}
ggplot(ano, aes(x = andel_hovedoekosystem_punkt))+
         geom_histogram(fill = "grey",
                        colour="black",
                        binwidth = 1)+
  theme_bw(base_size = 12)+
  xlab("Percentage cover of the Våtmark main ecosystem\n in the 250m2 circle")+
  scale_x_continuous(limits = c(0,101),
                     breaks = seq(0,100,10))
```

We can see that people tend to record the variable in steps of 5%, and
that most sample points are 100% belonging to the same main ecosystem.

We want to use area weighting in this indicator, so we can use this
percentage cover data to calculate the area. Note that both data sets
use m^2^ as area units.

```{r}
ano$area <- (ano$andel_hovedoekosystem_punkt/100)*250
```

Let's now look at the distribution of the variable. It is coded as `fa_total_dekning`. 
First, there are `r length(ano[is.na(ano$fa_total_dekning),])` NA's in this column which we can remove.

```{r}
ano <- ano[!is.na(ano$fa_total_dekning),]
```

Let us plot the distribution of values.

```{r ano-alien-dist, fig.cap"Distribution of %cover of alien plants variable in the ANO data set"}
ggplot(ano, aes(x = fa_total_dekning))+
  geom_histogram(fill="grey",
           colour = "black",
           binwidth = 1)+
  theme_bw(base_size = 12)+
  labs(x = "% cover alien vascular plants") +
  facet_wrap(.~hovedoekosystem_punkt, scales="free_y")
```

The ANO localities seemingly have less alien plants in them than the nature type localities. This can be du to sampling biases in the latter.

Now I will convert these %cover values into 7FA classes. This is a critical step, since there is not one-to-one mapping here, 
but rather there is considerable overlap, as can be seen in Fig. \@ref(fig:fremmedart-prosentVSinnslag).

```{r}
ano <- ano %>%
  mutate(var_7FA = case_when(
             fa_total_dekning>90 ~ 7,
             fa_total_dekning>25 ~ 6,
             fa_total_dekning>10 ~ 5,
             fa_total_dekning>5 ~ 4,
             fa_total_dekning>2 ~ 3,
             fa_total_dekning>1 ~ 2,
             fa_total_dekning<1 ~ 1))
```

```{r ano-alien-dist-R7, fig.cap"Distribution of R7 values in the ANO data set"}
ggplot(ano, aes(x = var_7FA))+
  geom_histogram(fill="grey",
           colour = "black",
           binwidth = 1)+
  theme_bw(base_size = 12)+
  labs(x = "% cover alien vascular plants\n converted to the R7 scale") +
  facet_wrap(.~hovedoekosystem_punkt, scales="free_y")
```

##### Combine Naturtype data and ANO {#combine-nt-ano2}

We need to combine the nature type data set with the ANO data set. I
will add a column `origin` to show where the data comes from. I will
also add a column with the main ecosystem.

```{r}
ano$origin <- "ANO"
naturetypes$origin <- "Nature type mapping"
ano$hovedøkosystem <- ano$hovedoekosystem_punkt
ano$kartleggingsår <- ano$aar
```

Fix class

```{r}
naturetypes$kartleggingsår <- as.numeric(naturetypes$kartleggingsår)
naturetypes$area <- units::drop_units(naturetypes$area)
```

I use `dplyr::select` to reduce the number of columns to keep things a
bit more tidy.

```{r}
alien_data <- dplyr::bind_rows(select(ano,
                                  GlobalID,
                                  origin,
                                  kartleggingsår,
                                  hovedøkosystem,
                                  area,
                                  var_7FA,
                                  SHAPE), 
                           select(naturetypes,
                                  identifikasjon_lokalId,
                                  origin,
                                  hovedøkosystem,
                                  kartleggingsår,
                                  area,
                                  var_7FA = NiN_variable_value,
                                  SHAPE))
```

###### Points to polygons {#ANO-points-to-poly}

The ANO data is point data and the nature type data is *multipolygon*:
```{r}
unique(st_geometry_type(alien_data))
```

Because later we might want to save this as a shape file, we cannot have
a mixed class type. I will therefore convert the points to polygons. I
use the area column to calculate a radii that gives that area.

```{r}
alien_data_points <- alien_data %>%
  mutate(g_type = st_geometry_type(.)) %>%
  filter(g_type =="POINT") %>%
  st_buffer(sqrt(alien_data$area/pi))
```

Checking now that the new polygons have the area corresponding to the
proportion of the point that was part of the same main ecosystem:

```{r new-area2, fig.cap="Checking that the area of the new polygons fall in line with the proportion of each point which is part of the main ecosystem."}
alien_data_points$area2 <- st_area(alien_data_points)
plot(alien_data_points$area, alien_data_points$area2,
     xlab = "Target area",
     ylab = "Area of the new polygons")
abline(0,1)
```

The area calculation seems to have worked fine. Checking that the new
data set contains **only polygons**.

```{r}
alien_data_polygons <- alien_data %>%
  mutate(g_type = st_geometry_type(.)) %>%
  filter(g_type !="POINT")

alien_data <- bind_rows(alien_data_points, alien_data_polygons)

unique(st_geometry_type(alien_data))
```

Ok.

#### Distribution of 7FA scores

```{r, fig.cap="7FA (alien species) scores (ANO Naturetype (and GRUK) data combined)."}
temp <- as.data.frame(table(alien_data$var_7FA))
ggplot(temp, aes(x = Var1,
                 y = Freq))+
  geom_bar(stat="identity",
           fill="grey",
           colour = "black")+
  theme_bw(base_size = 12)+
  labs(x = "7FA score",
       y = "Number of localities")
```

Let's see the proportion of data points (not area) origination from each
data set

```{r, fig.cap="Barplot show the contribution (number of localities) of different data sets to the alien species indicator.", fig.width=4}
temp <- as.data.frame(table(alien_data$origin))

ggplot(temp, aes(x = Var1,
                 y = Freq))+
  geom_bar(stat="identity",
           fill="grey",
           colour = "black")+
  theme_bw(base_size = 12)+
  labs(x = "Data origin",
       y = "Number of localities")
```

So the ANO data is not very important here, but it can become more
important in the future so good to have them included in the workflow.

#### Outline of Norway and regions

The GIS layers are used to crop out marine areas, and to define
accounting areas, respectively.

```{r}
outline <- sf::read_sf("data/outlineOfNorway_EPSG25833.shp")
```

```{r}
regions <- sf::read_sf("data/regions.shp", options = "ENCODING=UTF8")
unique(regions$region)
```

Municipalities

```{r}
path1 <- ifelse(dir == "C:", 
               "R:/",
               "/data/R/")

path <- paste0(path1, "GeoSpatialData/AdministrativeUnits/Norway_AdministrativeUnits/Original/Norway_Municipalities/Basisdata_0000_Norge_25833_Kommuner_FGDB.gdb")

# find the correct layer
#st_layers(path)
muni <- sf::read_sf(path, options = "ENCODING=UTF8", layer = "kommune")

```

```{r, include=F, eval=FALSE}
# If I import the manuicipality names:
muni_names <- sf::read_sf(path, options = "ENCODING=UTF8", layer = "administrativenhetnavn")
#The encoding option fail for gdb-files, so I fix the names again here
muni_names$navn <- stringi::stri_encode(muni_names$navn, from = "UTF-8", to = "UTF-8")

# ... there is no common ID to match these with the municipality polygons!!
```

There are some multi-surfaces in here that I will convert to multi-polygons before plotting.

```{r kommuner, fig.cap="Municipalities in Norway."}
muni <- st_cast(muni, "MULTIPOLYGON")
tm_shape(muni)+
  tm_polygons(col="kommunenummer")+
  tm_layout(legend.outside = T)
```

### Scaled indicator values {#scale-alien-ind}

I can scale the indicator for each polygon, or I can chose to aggregate
them first. If the scaled value is representative and precise at the
polygon level, then I could scale at that level. I think they are.

However, the combined surveyed area is a very small fraction of the
total area of Norway, so that only producing indicator values for the
mapped areas leaves the indicator without much value for regional
assessments. When we do regional assessments and calculate regional
indicator values, we cannot simply do an area weighting of the polygons
in each region. This is because we don't want to assume that the
polygons are representative far outside of the mapped area. But perhaps
we can assume them to be representative inside *homogeneous ecological
areas,* or *Homoegeneous Impact Areas* (HIAs). That's where the
infrastructure index comes in. Here's the plan:

1.  normalise the indicators at the polygon level.

2.  take a simplified infrastructure index (vector data set with HIAs)
    and extract the corresponding indicator values that intersect with a
    given HIA, and extrapolate the area weighted mean of those values to
    the entire HIA withing a given geographical region (municipality). Errors are
    calculated from bootstrapping.

3.  Take the new area weighted indicator values for the HIAs and match
    them with ecosystem occurrences of the same HIA (_possibly using
    randomly sampled points from the available ecosystem occurrences_).
    The errors are the same as for the HIAs.

4.  calculate an indicator value for a region (accounting area) by doing
    a weighted average based of the relative area of ecosystem
    occurrences. Errors should be carried somehow, perhaps via weighted
    resampling.

#### Scaled variable {#scaled-slitasje-variable}

I will use the same reference levels/values for all of Norway:

```{r}
low <- 1
high <- 7
threshold <- 4
```

```{r alien-normalise, fig.cap="Performing a linear break-point type normalisation of the slitasje variable.", fig.width=4}
eaTools::ea_normalise(data = alien_data,
                      vector = "var_7FA",
                      upper_reference_level = high,
                      lower_reference_level = low,
                      break_point = threshold,
                      plot=T,
                      reverse = T
                      )
```

There is no point yet making this a time series, and I will
assign all the indicator value to the average year of the data.

```{r}
mean(alien_data$kartleggingsår)
```

Assigning the indicator to year 2020.

```{r}
alien_data$i_2020 <- eaTools::ea_normalise(data = alien_data,
                      vector = "var_7FA",
                      upper_reference_level = high,
                      lower_reference_level = low,
                      break_point = threshold,
                      reverse = T
                      )
```

### Homogeneous impact areas {#HIA}

I want to use the [Homogeneous Impact Areas](#HIA) (HIA) to define
smaller regions into which I can extrapolate the indicator values. This
data is generated by discretizing the Norwegian Infrastructure Index. I
refer to the ordinal values of the four HIA classes as their *Human
Impact Factor* (HIF). This is just to keep the approach separate from
the Norwegian Infrastructure Index.

```{r}
HIA <- readRDS(paste0(pData, "infrastrukturindeks/homogeneous_impact_areas.rds"))
```

I want to check that HIF is in fact a good predictor for *alien species*.

I also want to split these four HIA classes based on municipality. To do
this I need the two layers to have the same CRS.

```{r}
st_crs(HIA) == st_crs(muni)
```

Then we get the intersections (unique combinations)

```{r, eval=F}
HIA_muni <- eaTools::ea_homogeneous_area(HIA,
                             muni,
                             keep1 = "infrastructureIndex",
                             keep2 = "kommunenummer")
saveRDS(HIA_muni, "P:/41201785_okologisk_tilstand_2022_2023/data/cache/HIA_muni.rds")
```

```{r, include=F}
# silently read cached file
HIA_muni <- readRDS(paste0(pData, "cache/HIA_muni.rds"))
```

Create a new column by crossing municipality number and HIF

```{r}
HIA_muni <- HIA_muni %>%
  mutate(muni_HIF = paste("ID", kommunenummer, infrastructureIndex, sep="_"))
```

#### Validate

I now have `r length(unique(HIA_muni$muni_HIF))` unique areas (for each main ecosystem) that I will, given
there is data, extrapolate indicator values over.

###### Subset ETs

I will subset the `alien_data` into the three ecosystems. Note that
only for open wetland do we have good ecosystem delineation maps to base
a spatial averaging on.

```{r}
#make valid
alien_data <- st_make_valid(alien_data)

#subset
wetlands <- alien_data[alien_data$hovedøkosystem == "våtmark",]
seminat <- alien_data[alien_data$hovedøkosystem == "semi-naturligMark",]
natOpen <- alien_data[alien_data$hovedøkosystem == "naturligÅpneOmråderUnderSkoggrensa",]
```

Creating some summary statistics.

```{r, eval=F}
wetland_stats <- ea_spread(indicator_data = wetlands,
          indicator = i_2020,
          regions = HIA_muni,
          groups = muni_HIF,
          summarise = TRUE)

seminat_stats <- ea_spread(indicator_data = seminat,
          indicator = i_2020,
          regions = HIA_muni,
          groups = muni_HIF,
          summarise = TRUE)

natOpen_stats <- ea_spread(indicator_data = natOpen,
          indicator = i_2020,
          regions = HIA_muni,
          groups = muni_HIF,
          summarise = TRUE)

wetland_stats <- wetland_stats %>%
  add_column(eco = "wetland")

seminat_stats <- seminat_stats %>%
  add_column(eco = "semi-natural")

natOpen_stats <- natOpen_stats %>%
  add_column(eco = "Naturally-open")

all_stats <- rbind(wetland_stats,
                   seminat_stats,
                   natOpen_stats)
all_stats <- all_stats %>%
  separate(muni_HIF,
           into = c("tempLink", "municipalityNumber", "HIF"),
           sep = "_")

#saveRDS(all_stats, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/all_stats_alienSpecies.rds")
saveRDS(all_stats, "P:/41201785_okologisk_tilstand_2022_2023/data/cache/all_stats_alienSpecies.rds")

```

```{r, include=F}
#silently read cached file
all_stats <- readRDS(paste0(pData, "cache/all_stats_alienSpecies.rds"))
```

We have data from `r length(unique(all_stats$municipalityNumber))` out of `r nrow(muni)` municipalities. This figure shows how many municipalities we have at least one data point for, specific for each human impact factor level:

```{r, fig.cap="Number of municipalities we have at least one data point for, specific for each human impact factor level."}
all_stats %>%
  group_by(HIF, eco) %>%
  summarise(n = n()) %>%
  ggplot()+
  geom_bar(aes(x = HIF, y = n), stat ="identity")+
  theme_bw()+
  ylab("Number of municipalities")+
  facet_wrap(.~eco)
```

If we say that we need 20 data points for each HIA and municipality combintaion in order to extrapolate to the entire HIA for that municipality, we get much less to work with:

```{r, fig.cap="Number of municipalities we have at least 20 data points for, specific for each human impact factor level."}
all_stats %>%
  filter(n>20) %>%
  group_by(HIF, eco) %>%
  summarise(n = n()) %>%
  ggplot()+
  geom_bar(aes(x = HIF, y = n), stat ="identity")+
  theme_bw()+
  ylab("Number of municipalities")+
  facet_wrap(.~eco)

```

```{r, fig.cap="Indicator to pressure relationship across all ecosystems."}
all_stats %>%
  filter(n > 20) %>%
  ggplot(aes(x = HIF, y = w_mean))+
  geom_point(size=2, position = position_dodge2(.1))+
  geom_violin(alpha=0)+
  theme_bw()
```

There is a trend here, but not very strong.

Let us look at the effect of sample size on the indicator uncertainty.

```{r, fig.cap="Sample size against indicator uncertainty."}
all_stats %>%
  filter(n > 5) %>%
  ggplot(aes(x = n, y = sd))+
  geom_point(size=2, position = position_dodge2(.1))+
  theme_bw()+
    facet_wrap(.~eco, scales="free")
```

This shows that the uncertainty is really inflated with sample sizes less than about 50-100.

```{r, fig.cap="Summary statistics for the alien species indicator."}
DT::datatable(all_stats) %>%
  formatRound(columns=3:8, digits=2)
```

In addition to looking at the area weighted means and its relationship
with the pressure indicator (HIF), we can do the same at the polygon
level. Because the indicator is pseudo-continuous, boxplots for example
become useless, so I will present a relative frequency plot.

```{r, eval=F}
corrCheck <- st_intersection(alien_data, HIA_muni)
saveRDS(corrCheck, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/corrCheck.rds")
```

```{r, include=F}
#Silently read cached data
corrCheck <- readRDS(paste0(pData, "cache/corrCheck.rds"))
```

```{r slitasje-precentageplot, fig.height=12, fig.width=12, fig.cap="Relative frequency plot (conditioned on region and main ecosystem) showing the distribution of polygons with different indicator values."}
ggplot(corrCheck, aes(x = factor(infrastructureIndex), fill = factor(round(i_2020,2))))+
  geom_bar(position="fill")+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Sliatsje indicator"))+
  ylab("Fraction of data points")+
  xlab("HIF")+
  scale_fill_brewer(palette = "RdYlGn")+
  facet_grid(hovedøkosystem~region)
```

The figure above I think supports the indicator-pressure relationship
and justifies using the HIA+region intersections as local reference
areas. Especially when down-weighting the importance of those
groups/categories that have very little data.

#### Looking at the HIA

Here is a view of the data zooming in on Trondheim

```{r}
myBB <- st_bbox(c(xmin=260520.12, xmax = 278587.56,
                ymin = 7032142.5, ymax = 7045245.27),
                crs = st_crs(HIA_regions))
```

Cropping the raster to the bbox

```{r}
HIA_trd <- sf::st_crop(HIA_regions, myBB)
```

Get map of major roads, for context

```{r}
hw_utm <- readRDS("data/cache/highways_trondheim.rds")
```

```{r, fig.cap="A closer look at the HIA designation over Trondheim"}
(HIA_trd <- tm_shape(HIA_trd)+
  tm_polygons(col = "infrastructureIndex",
    title="Infrastructure index\n(modified 4-step scale)",
    palette = "-viridis",
    style="cat")+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="red")+
  tm_shape(outline)+
  tm_borders(col = "black", lwd=2))
```

Let's calculate the areas of these polygons and compare the HIF in the
five regions.

```{r}
HIA_regions$area <- sf::st_area(HIA_regions)
```

```{r HIF-region, fig.cap="Stacked barplot showing the distribution of human impact factor across five regions in Norway."}
temp <- as.data.frame(HIA_regions) %>%
  group_by(region, infrastructureIndex) %>%
  summarise(area = mean(area))

ggplot(temp, aes(x = region, y = area, fill = factor(infrastructureIndex)))+
  geom_bar(position = "stack", stat = "identity")+
  guides(fill = guide_legend("HIF"))
```

The figure above shows that _Nord-Norge_ for example, has a lot of
relatively untouched areas, and that _Østlandet_ has the highest
proportion of impacted areas. This is expected.

### Aggregate and spread (extrapolate) {#spread-slitasje}

I now want to find the mean indicator value for each HIA (i.e. to
aggregate) and to spread these out spatially to the entire HIA within
each region (i.e. to extrapolate).

I want to add a threshold so that we don't end up over extrapolating
based on too few data points. I have previousl used 150 data pionts, and I think that was about right.

```{r, eval=F}
wetlands_slitasje_extr <- ea_spread(indicator_data = wetlands,
                         indicator = i_2020,
                         regions = HIA_regions,
                         groups = region_HIF,
                         threshold = 150)

seminat_slitasje_extr <- ea_spread(indicator_data = seminat,
                         indicator = i_2020,
                         regions = HIA_regions,
                         groups = region_HIF,
                         threshold = 150)

natOpen_slitasje_extr <- ea_spread(indicator_data = natOpen,
                         indicator = i_2020,
                         regions = HIA_regions,
                         groups = region_HIF,
                         threshold = 150)
#saveRDS(wetlands_slitasje_extr, "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/wetlands_slitasje_extr.rds")
#saveRDS(seminat_slitasje_extr,  "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/seminat_slitasje_extr.rds")
#saveRDS(natOpen_slitasje_extr,  "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/natOpen_slitasje_extr.rds")
```

```{r, include=F}
wetlands_slitasje_extr <- readRDS(paste0(pData, "cache/wetlands_slitasje_extr.rds"))
seminat_slitasje_extr  <- readRDS(paste0(pData, "cache/seminat_slitasje_extr.rds"))
natOpen_slitasje_extr  <- readRDS(paste0(pData, "cache/natOpen_slitasje_extr.rds"))
```

It's easier to see what's happening if we zoom in a bit. Lets get a
boundary box around Trondheim.

```{r}
myBB <- st_bbox(c(xmin=260520.12, xmax = 278587.56,
                ymin = 7032142.5, ymax = 7045245.27),
                crs = st_crs(wetlands_slitasje_extr))
```

Cropping the data to the bbox

```{r, warning=F}
wetlands_slitasje_extr_trd <- sf::st_crop(wetlands_slitasje_extr, myBB)
seminat_slitasje_extr_trd <- sf::st_crop(seminat_slitasje_extr, myBB)
natOpen_slitasje_extr_trd <- sf::st_crop(natOpen_slitasje_extr, myBB)
```

```{r extrapolated-slitasje, fig.cap="Slitasje indicator extrapolated over Trondheim"}
myCol <- "RdYlGn"
tmap_arrange(
tm_shape(wetlands_slitasje_extr_trd)+
  tm_polygons(col = "w_mean",
    title="Slitasje indicator",
    palette = myCol,
    style="cont",
    breaks = c(0,1))+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="blue")+
  tm_layout(title = "Wetland"),

tm_shape(seminat_slitasje_extr_trd)+
  tm_polygons(col = "w_mean",
    title="Slitasje indicator",
    palette = myCol,
    style="cont",
    breaks = c(0,1))+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="blue")+
  tm_layout(title = "Semi-natural areas"),

tm_shape(natOpen_slitasje_extr_trd)+
  tm_polygons(col = "w_mean",
    title="Slitasje indicator",
    palette = myCol,
    style="cont",
    breaks = c(0,1))+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="blue")+
  tm_layout(title = "Naturally open areas"),
HIA_trd)
```

Note that for wetland and semi-natural areas we did not
extrapolate values for the most urban areas, and that for naturally open
areas we did not extrapolate to the least affected areas.
For the other regions we generally excluded more HIFs.

> *The maps above are easy to misunderstand, and they should therefor be
> communicated with caution. They should be considered as temporary
> files or as by-products.*

The next step now is to

1. Prepare ecosystem delineation maps in EPSG25833 and perfectly aligned to a master grid

2. Rasterize the extrapolated indicator map, using the ET map as a template

3. and mask it using the perfectly aligned ET map.

A faster way could be to:

1.  Draw *n* random point samples from the ET map, and

2.  extract indicator values at those points (from the extrapolated
    indicator map (a stars object)).

... but his will not provide spatially explicit indicator values for the
entire ecosystem delineation.

### Ecosystem map

Importing the wetland map.

```{r, eval=T}
#takes about one minute without the proxy
(mireET <- stars::read_stars("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Myrmodell/myrmodell90pros.tif",
                             proxy=T))
```

This is in the wrong CRS. We want to stay with 25833.

We have a 50x50 m [master grid](#master-grid) with the correct CRS, and I would warp mireET to fit that grid like this:
```{r, eval=F}
grid <- stars::read_stars(paste0(pData, "masterGrid_50m.tiff"))
mireET_warped <- stars::st_warp(mireET, grid)
```

... but I don't want to down sample the mire data, since we will loose a lot of the smaller ecosystem occurrences.

I could make a 10x10 master grid, but I have not found a way to stop this from being just a massive file.

I can transform the ET map like this, and not worry about the rasters being perfectly aligned, but even this operation is too heavy for the R Studio Server:

```{r, eval=F}
mireET_trans <- sf::st_transform(mireET, 25833)
```

I will therefore ask GIS engineers to prepare ecosystem maps to the standard that we chose, so that we can simply combine the stars object with `stars::c()`, or perhaps side-step that operatio and just mask directly like this: `wetlands_slitasje_extr[mireET==0] <- NA`.

### Aggregate to region

We need to aggreate the ET occurence and their indicator values, to a
regional delineation. This functionality can be developed in `eaTools`,
but will in any case build around `exactextratr::exact_extract()`.

To carry the errors, we can for example
produce distributions for the indicator values for each ET
occurrence (or groups of them) and resample the distributions with an
area weighting.

### Uncertainty

The calculations of indicator uncertainty is done with bootstrapping, as
is described in eaTools.

## Eksport files (final product)

According to the workflow specifications, I will export the indicator
map at its highest resolution, for use in various web interfaces etc.
For the finest resolution indicator map I will not include columns for
uncertainty (because this will only be estimated at the aggregated
level), or reference levels (because there is no spatial variation).

I can also export a map of the indicator where we extrapolate the
indicator values to the homogeneous areas (e.g. Fig. \@ref(fig:extrapolated-slitasje) . This map should be
interpreted with caution, since they appear as providing explicit indicator
values for areas where there is no data. This map is mainly for
aggregation purposes. I will mark this map with the prefix
*homogeneous-areas*. I will also export this same map after masking it
with the ET map, and this output must come with the same disclaimers.

Finally I will export a map of the five accounting areas with
wall-to-wall indicator value and errors (e.g. Fig \@ref(fig:wall-to-wall-summer-temp-indicator)). This map can also contain data
covering other ETs. Otherwise the maps would not we visually
interpretative.

### Slitasje in wetlands

This is how I plan to export the final product.

```{r, eval=F}
# keep col names short and standardized
exp_wetland <- ...
```

```{r, eval=F}
exp_wetland_HIAextrapolated <- …
```

```{r, eval=F}
exp_wetland_AAextrapolated <- …
```

Write to file

```{r, eval=F}
st_write(exp_wetland, "indicators/indicatorMaps/slitasje/wetland_slitasje.shp",
         append = F)

st_write(exp_wetlandHIAextrapolated, 
         "indicators/indicatorMaps/extrapolated_to_homogeneous_areas/slitasje/homogeneous_areas_wetland_slitasje.shp",
         append = F)

st_write(exp_wetland, "indicators/indicatorMaps/extrapolated_to_accounting_areas/slitasje/accounting_areas_wetland_slitasje.shp",
         append = F)
```