### 2023-06-14 DRAFT_ndvi_wetland.Rmd
markdown text

# Functional plant indicators, wetland ecosystems (Moisture, Light, pH, Nitrogen) {#Functional-plant-indicators-wetland}

<br />

_Author and date:_
Joachim Töpper, Tessa Bargmann

June 2023

<br />

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(sf)
library(plyr)
library(stringr, lib.loc = "/usr/lib/R/site-library")
library(tidyverse, lib.loc = "/usr/local/lib/R/site-library")
library(data.table)
library(lubridate)
library(RColorBrewer)
library("gridExtra") 
library(ggridges)
library(tmap)
library(tmaptools)
library(raster)
library(stars)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
Ecosystem <- "våtmark"
Egenskap  <- c("Primærproduksjon")
ECT       <- "Structural state characteristics"
Contact   <- "Joachim Töpper"
```

```{r, echo=FALSE}
metaData <- data.frame(Ecosystem,
                       "Økologisk egenskap" = Egenskap,
                       "ECT class" = ECT)
knitr::kable(metaData)
```

<!-- Don't remove these three html lines -->
<br />
<br />
<hr />

## Introduction
The Normalized difference vegetation index (NDVI) can be used to describe the greenness of an area, and may thus represent a proxy for the amount of chlorophyll and plant productivity in the given area. Different ecosystem types naturally have greenness signals of different intensity but also with a certain variance. Deviations beyond the natural variation in either direction may indicate a reduction in ecological condition of a given ecosystem. E.g. lower than normal NDVI may indicate excessive disturbance, while higher than normal NDVI may indicate regrowth and transitions towards more densely vegetated systems. In recent assessments of ecological condition for forest and mountain ecosystems, two different NDVI-based indicators were developed and applied. For forests, NDVI-values across the country were compared to a spatial reference NDVI-signal from protected areas. For mountains, the extent and consistence of the temporal change of NDVI-values over the last twenty years was assessed. In both cases, the NDVI-product from the MODIS satellite reflectance imagery was used based on available maps of spatial ecosystem extent in Norway (i.e. national maps for the forest and mountain ecosystems). Here, we explore the potential for developing an NDVI-indicator for ecological condition representing open wetland ecosystems (i.e. excluding forest swamps). For this task, two initial challenges can be identified early on:
1) We don't yet have a national wetland map, i.e. we don't know the exact locations of all wetlands in Norway, which is a prerequisite to extract satellite data for the right locations and ecosystems, and thus achieve a good and unbiased spatial representation for the assessment area.
2) The spatial extent of the single wetland ecosystem occurrences (hereafter 'polygons') includes many wetland areas which are way smaller than the pixel size in the MODIS imagery (250 x 250 m) which was used for forests and mountains and has a nice time series back to year 2000.

In this draft development for a wetland NDVI indicator for ecological condition we approach these challenges in the following ways:
1) We use ecosystem occurrence data from NiN-mapping to guide the extraction of satellite NDVI data for pixels representing the right ecosystem. However, the ecosystem occurrences in the NiN data are highly spatially biased and thus cannot provide a good and unbiased spatial representation for the assessment area. Our work therefore represents a concept and method development which may be applied in an assessment of ecological condition once spatially unbiased occurrence data or a national map of wetland ecosystems are available.
2) We include Sentinel-II imagery for wetland ecosystems as the resolution for the bands underlying the NDVI product here is 10 x 10 m which is smaller than almost all wetland polygons in the NiN-data. However, since Sentinel-II only goes back to 2015, we lose much of the time series aspect as compared to the MODIS data. In addition, we still explore NDVI data from the MODIS and Landsat satellites (see more details below).

## About the underlying data
In the NDVI project for wetlands, we use four sets of data for building indicators for ecological condition:

- NiN ecosystem data from 'Naturtypekartlegging etter Miljødirektoratets innstruks' to guide extraction of NDVI data for the right locations and to provide information on the field-assessed ecological condition of mapped polygons
- Sentinel-II satellite NDVI data
- MODIS satellite NDVI data
- Landsat satellite NDVI data

(1) NiN ecosystem data:

(2) Sentinel-II satellite NDVI data:

(3) MODIS satellite NDVI data:

(4) Landsat satellite NDVI data:

have to exclude all polygons smaller than the MODIS pixel size of 250 x 250 m, which reduces the number of available polygons greatly. To explore temporal trends even further back in time, we also include Landsat NDVI data, which span a time frame from ??? to ???, and have a resolution of 30 x 30 m, thus excluding only few polygons of insufficient size.

### Representativity in time and space
For wetlands, the NiN data contain 6503 wetland polygons of known area size across mainland Norway. However, NiN-mapping is not planned and performed in a spatially representative way and thus both the ecosystem occurrences and their spatial extent have to be treated as spatially biased. The 6503 wetland polygons are distributed across regions and main ecosystem types in the following way:

----build a matrix-table----

The NDVI data from either satellite follow the spatial distribution and bias status of the NiN-data. Due to the larger number of excluded polygons in the NDVI data from MODIS (because of the large pixel size in MODIS, see above), the MODIS NDVI data for the wetland occurrences in the NiN data likely are even more spatially biased than it is the case for the NDVI data from Sentinel-II and Landsat.

### Temporal coverage
The currently available NiN data on wetland ecosystems span a time period from 2018-2021 and thus represent a contemporary sample of ecosystem occurrences. Note that none of the polygons in the NiN-data has been visited more than once during that period.

Sentinel-II satellite NDVI data span the years 2015-2022.
MODIS satellite NDVI data span the years ????-????.
Landsat satellite NDVI data span the years ????-????.

## Collinearities with other indicators
The NDVI signal may likely be co-linear with the re-growth indicator based on LiDAr derived canopy height in drained wetland areas under shrub and tree encroachment.

## Reference state and values
### Reference state
This indicator project is mainly exploratory, mapping the possibilities of and lacking prerequisites for satellite derived NDVI data to inform ecological condition assessments. Therefore, we lack a concise strategy for how to define the reference state, especially prior to having access to an ecosystem map for wetland ecosystems. In this exploratory work, we will use the condition classification in the NiN-data to explore its applicability for defining a reference state. We will use statistical models to test whether or not NDVI varies systematically with condition and across time.

### Reference values, thresholds for defining _good ecological condition_, minimum and/or maximum values
We will explore (i) if the condition classes in the NiN data can be used to define scaling values, and (ii) the potential for taking a probabilistic approach to scaling original NDVI values to a condition index based on probability density curves derived from above named models.

## Uncertainties
Given a condition index can be achieved, we can calculate a mean indicator value (after scaling) for every region (or any other delimited area of interest) as well as its corresponding standard error and standard deviation as a measure of spatial uncertainty for a geographical area.

## References

----not referred to in text yet, but likely will :)----

Framstad, E., Kolstad, A. L., Nybø, S., Töpper, J. & Vandvik, V. 2022. The condition of forest and mountain ecosystems in Norway. Assessment by the IBECA method. NINA Report 2100. Norwegian Institute for Nature Research.

Halvorsen, R., Skarpaas, O., Bryn, A., Bratli, H., Erikstad, L., Simensen, T., & Lieungh, E. (2020). Towards a systematics of ecodiversity: The EcoSyst framework. Global Ecology and Biogeography, 29(11), 1887-1906. doi:10.1111/geb.13164

Jakobsson, S., Töpper, J.P., Evju, M., Framstad, E., Lyngstad, A., Pedersen, B., Sickel, H., Sverdrup-Thygeson, A., Vandvik. V., Velle, L.G., Aarrestad, P.A. & Nybø, S. 2020. Setting reference levels and limits for good ecological condition in terrestrial ecosystems. Insights from a case study based on the IBECA approach. Ecological Indicators 116: 106492.

Töpper, J. & Jakobsson, S. 2021. The Index-Based Ecological Condition Assessment (IBECA) - Technical protocol, version 1.0. NINA Report 1967. Norwegian Institute for Nature Research.

## Analyses
### Data sets

#### NiN data
We read the NiN data for this document from a cache at the NINA server...
```{r, echo = F}
# Add data from cache
nin<-readRDS(paste0(here::here(),"/data/cache/nin.RDS"))
```

...but it can also be directly downloaded from Miiljødirektoratets kartkatalog (https://kartkatalog.miljodirektoratet.no/Dataset)
```{r, echo = T, eval = F}

# st_layers(dsn = "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Naturovervaking_eksport.gdb")
# ANO.sp <- st_read("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Naturovervaking_eksport.gdb",
#                    layer="ANO_Art")
# ANO.geo <- st_read("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Naturovervaking_eksport.gdb",
#                   layer="ANO_SurveyPoint")
# head(ANO.sp)
# head(ANO.geo)
```

Plant indicators from Tyler et al. (2021)
```{r, echo = F}
## Tyler indicator data
## data from cache
ind.dat <-readRDS(paste0(here::here(),"/data/cache/ind.dat.RDS"))
# ind.dat <- read.table("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/functional plant indicators/Tyler et al_Swedish plant indicators.txt",
#                         sep = '\t', header=T, quote = '')
# head(ind.dat)

```

Generalized species lists NiN
```{r, echo = F}
# data from cache
Eco_State<-readRDS(paste0(here::here(),"/data/cache/Eco_State.RDS"))
# load("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/functional plant indicators/reference from NiN/Eco_State.RData")
# str(Eco_State)
```

#### Data handling
- Checking for errors
- Checking species nomenclature in the different species lists to make species and indicator data possible to merge
- Merging indicator data with monitoring data and indicator data with reference data
(not shown here, but documented in the code)

```{r, include = F}

### Plant indicator data
names(ind.dat)
names(ind.dat)[1] <- 'species'
ind.dat$species <- as.factor(ind.dat$species)
summary(ind.dat$species)
ind.dat <- ind.dat[!is.na(ind.dat$species),] # removing species-NAs
ind.dat[,'species.orig'] <- ind.dat[,'species'] # make a backup of the original species variable
ind.dat[,'species'] <- word(ind.dat[,'species'], 1,2) # trimming away sub-species & co, and descriptor info

# the trimming above leaves some duplicates that need to be taken care of
ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat.dup <- ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat[ind.dat$species %in% ind.dat.dup,c("Light","Moisture","Soil_reaction_pH","Nitrogen","species.orig","species")]
ind.dat <- ind.dat %>% filter( !(species.orig %in% list("Ammophila arenaria x Calamagrostis epigejos",
                                                            "Anemone nemorosa x ranunculoides",
                                                            "Armeria maritima ssp. elongata",
                                                            "Asplenium trichomanes ssp. quadrivalens",
                                                            "Calystegia sepium ssp. spectabilis",
                                                            "Campanula glomerata 'Superba'",
                                                            "Dactylorhiza maculata ssp. fuchsii",
                                                            "Erigeron acris ssp. droebachensis",
                                                            "Erigeron acris ssp. politus",
                                                            "Erysimum cheiranthoides L. ssp. alatum",
                                                            "Euphrasia nemorosa x stricta var. brevipila",
                                                            "Galium mollugo x verum",
                                                            "Geum rivale x urbanum",
                                                            "Hylotelephium telephium (ssp. maximum)",
                                                            "Juncus alpinoarticulatus ssp. rariflorus",
                                                            "Lamiastrum galeobdolon ssp. argentatum",
                                                            "Lathyrus latifolius ssp. heterophyllus",
                                                            "Medicago sativa ssp. falcata",
                                                            "Medicago sativa ssp. x varia",
                                                            "Monotropa hypopitys ssp. hypophegea",
                                                            "Ononis spinosa ssp. hircina",
                                                            "Ononis spinosa ssp. procurrens",
                                                            "Pilosella aurantiaca ssp. decolorans",
                                                            "Pilosella aurantiaca ssp. dimorpha",
                                                            "Pilosella cymosa ssp. gotlandica",
                                                            "Pilosella cymosa ssp. praealta",
                                                            "Pilosella officinarum ssp. peleteranum",
                                                            "Poa x jemtlandica (Almq.) K. Richt.",
                                                            "Poa x herjedalica Harry Sm.",
                                                            "Ranunculus peltatus ssp. baudotii",
                                                            "Sagittaria natans x sagittifolia",
                                                            "Salix repens ssp. rosmarinifolia",
                                                            "Stellaria nemorum L. ssp. montana",
                                                            "Trichophorum cespitosum ssp. germanicum")
) )

# checking duplicates again
ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat.dup <- ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat[ind.dat$species %in% ind.dat.dup,c("Light","Moisture","Soil_reaction_pH","Nitrogen","species.orig","species")]
# getting rid of sect. for Hieracium
ind.dat <- ind.dat %>% mutate(species=gsub("sect. ","",species.orig))
ind.dat[,'species'] <- word(ind.dat[,'species'], 1,2)

ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat.dup <- ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat[ind.dat$species %in% ind.dat.dup,c("Light","Moisture","Soil_reaction_pH","Nitrogen","species.orig","species")]
# only hybrids left -> get rid of these
ind.dat <- ind.dat[!duplicated(ind.dat[,'species']),]
ind.dat[duplicated(ind.dat[,'species']),"species"]

ind.dat$species <- as.factor(ind.dat$species)
summary(ind.dat$species)
# no duplicates left

head(ind.dat)

### ANO monitoring data

head(ANO.sp)
head(ANO.geo)

## fix NiN information
ANO.geo$hovedtype_rute <- substr(ANO.geo$kartleggingsenhet_1m2,1,3) # take the 3 first characters
ANO.geo$hovedtype_rute <- gsub("-", "", ANO.geo$hovedtype_rute) # remove hyphen
unique(as.factor(ANO.geo$hovedtype_rute))

## fix NiN-variables
colnames(ANO.geo)
colnames(ANO.geo)[42:47] <- c("groeftingsintensitet",
                              "bruksintensitet",
                              "beitetrykk",
                              "slatteintensitet",
                              "tungekjoretoy",
                              "slitasje")
head(ANO.geo)

# remove variable code in the data
ANO.geo$groeftingsintensitet <- gsub("7GR-GI_", "", ANO.geo$groeftingsintensitet) 
unique(ANO.geo$groeftingsintensitet)
ANO.geo$groeftingsintensitet <- gsub("X", "NA", ANO.geo$groeftingsintensitet)
unique(ANO.geo$groeftingsintensitet)
ANO.geo$groeftingsintensitet <- as.numeric(ANO.geo$groeftingsintensitet)
unique(ANO.geo$groeftingsintensitet)

ANO.geo$bruksintensitet <- gsub("7JB-BA_", "", ANO.geo$bruksintensitet) 
unique(ANO.geo$bruksintensitet)
ANO.geo$bruksintensitet <- gsub("X", "NA", ANO.geo$bruksintensitet)
unique(ANO.geo$bruksintensitet)
ANO.geo$bruksintensitet <- as.numeric(ANO.geo$bruksintensitet)
unique(ANO.geo$bruksintensitet)

ANO.geo$beitetrykk <- gsub("7JB-BT_", "", ANO.geo$beitetrykk) 
unique(ANO.geo$beitetrykk)
ANO.geo$beitetrykk <- gsub("X", "NA", ANO.geo$beitetrykk)
unique(ANO.geo$beitetrykk)
ANO.geo$beitetrykk <- as.numeric(ANO.geo$beitetrykk)
unique(ANO.geo$beitetrykk)

ANO.geo$slatteintensitet <- gsub("7JB-SI_", "", ANO.geo$slatteintensitet) 
unique(ANO.geo$slatteintensitet)
ANO.geo$slatteintensitet <- gsub("X", "NA", ANO.geo$slatteintensitet)
unique(ANO.geo$slatteintensitet)
ANO.geo$slatteintensitet <- as.numeric(ANO.geo$slatteintensitet)
unique(ANO.geo$slatteintensitet)

ANO.geo$tungekjoretoy <- gsub("7TK_", "", ANO.geo$tungekjoretoy) 
unique(ANO.geo$tungekjoretoy)
ANO.geo$tungekjoretoy <- gsub("X", "NA", ANO.geo$tungekjoretoy)
unique(ANO.geo$tungekjoretoy)
ANO.geo$tungekjoretoy <- as.numeric(ANO.geo$tungekjoretoy)
unique(ANO.geo$tungekjoretoy)

ANO.geo$slitasje <- gsub("7SE_", "", ANO.geo$slitasje) 
unique(ANO.geo$slitasje)
ANO.geo$slitasje <- gsub("X", "NA", ANO.geo$slitasje)
unique(ANO.geo$slitasje)
ANO.geo$slitasje <- as.numeric(ANO.geo$slitasje)
unique(ANO.geo$slitasje)

## check that every point is present only once
length(levels(as.factor(ANO.geo$ano_flate_id)))
length(levels(as.factor(ANO.geo$ano_punkt_id)))
summary(as.factor(ANO.geo$ano_punkt_id))
# there's many double presences, probably some wrong registrations of point numbers,
# but also double registrations (e.g. ANO0159_55)
# CHECK THIS when preparing ecosystem-datasets for scaling

# fix species names
ANO.sp$Species <- ANO.sp$art_navn
unique(as.factor(ANO.sp$Species))
ANO.sp[,'Species'] <- word(ANO.sp[,'Species'], 1,2) # lose subspecies
ANO.sp$Species <- str_to_title(ANO.sp$Species) # make first letter capital
ANO.sp$Species <- gsub("( .*)","\\L\\1",ANO.sp$Species,perl=TRUE) # make capital letters after hyphen to lowercase
ANO.sp$Species <- gsub("( .*)","\\L\\1",ANO.sp$Species,perl=TRUE) # make capital letters after space to lowercase

## merge species data with indicators
ANO.sp.ind <- merge(x=ANO.sp[,c("Species", "art_dekning", "ParentGlobalID")], 
                y= ind.dat[,c("species", "Light", "Moisture", "Soil_reaction_pH", "Nitrogen")],
                by.x="Species", by.y="species", all.x=T)
summary(ANO.sp.ind)

## checking which species didn't find a match
unique(ANO.sp.ind[is.na(ANO.sp.ind$Moisture),'Species'])

# fix species name issues
ind.dat <- ind.dat %>% 
  mutate(species=str_replace(species,"Aconitum lycoctonum", "Aconitum septentrionale")) %>% 
  mutate(species=str_replace(species,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  mutate(species=str_replace(species,"Carex myosuroides", "Kobresia myosuroides")) %>%
  mutate(species=str_replace(species,"Clinopodium acinos", "Acinos arvensis")) %>%
  mutate(species=str_replace(species,"Artemisia rupestris", "Artemisia norvegica")) %>%
  mutate(species=str_replace(species,"Cherleria biflora", "Minuartia biflora"))

ANO.sp <- ANO.sp %>% 
  mutate(Species=str_replace(Species,"Arctous alpinus", "Arctous alpina")) %>%
  mutate(Species=str_replace(Species,"Betula tortuosa", "Betula pubescens")) %>%
  mutate(Species=str_replace(Species,"Blysmopsis rufa", "Blysmus rufus")) %>%
  mutate(Species=str_replace(Species,"Cardamine nymanii", "Cardamine pratensis")) %>%
  mutate(Species=str_replace(Species,"Carex adelostoma", "Carex buxbaumii")) %>%
  mutate(Species=str_replace(Species,"Carex leersii", "Carex echinata")) %>%
  mutate(Species=str_replace(Species,"Carex paupercula", "Carex magellanica")) %>%
  mutate(Species=str_replace(Species,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  mutate(Species=str_replace(Species,"Carex viridula", "Carex flava")) %>%
  mutate(Species=str_replace(Species,"Chamaepericlymenum suecicum", "Cornus suecia")) %>%
  mutate(Species=str_replace(Species,"Cicerbita alpina", "Lactuca alpina")) %>%
  mutate(Species=str_replace(Species,"Empetrum hermaphroditum", "Empetrum nigrum")) %>%
  mutate(Species=str_replace(Species,"Festuca prolifera", "Festuca rubra")) %>%
  mutate(Species=str_replace(Species,"Galium album", "Galium mollugo")) %>%
  mutate(Species=str_replace(Species,"Galium elongatum", "Galium palustre")) %>%
  mutate(Species=str_replace(Species,"Helictotrichon pratense", "Avenula pratensis")) %>%
  mutate(Species=str_replace(Species,"Helictotrichon pubescens", "Avenula pubescens")) %>%
  mutate(Species=str_replace(Species,"Hieracium alpina", "Hieracium Alpina")) %>%
  mutate(Species=str_replace(Species,"Hieracium alpinum", "Hieracium Alpina")) %>%
  mutate(Species=str_replace(Species,"Hieracium hieracium", "Hieracium Hieracium")) %>%
  mutate(Species=str_replace(Species,"Hieracium hieracioides", "Hieracium umbellatum")) %>%
  mutate(Species=str_replace(Species,"Hieracium murorum", "Hieracium Vulgata")) %>%
  mutate(Species=str_replace(Species,"Hieracium oreadea", "Hieracium Oreadea")) %>%
  mutate(Species=str_replace(Species,"Hieracium prenanthoidea", "Hieracium Prenanthoidea")) %>%
  mutate(Species=str_replace(Species,"Hieracium vulgata", "Hieracium Vulgata")) %>%
  mutate(Species=str_replace(Species,"Hieracium pilosella", "Pilosella officinarum")) %>%
  mutate(Species=str_replace(Species,"Hieracium vulgatum", "Hieracium umbellatum")) %>%
  mutate(Species=str_replace(Species,"Hierochloã« alpina", "Hierochloë alpina")) %>%
  mutate(Species=str_replace(Species,"Hierochloã« hirta", "Hierochloë hirta")) %>%
  mutate(Species=str_replace(Species,"Hierochloã« odorata", "Hierochloë odorata")) %>%
  mutate(Species=str_replace(Species,"Listera cordata", "Neottia cordata")) %>%
  mutate(Species=str_replace(Species,"Leontodon autumnalis", "Scorzoneroides autumnalis")) %>%
  mutate(Species=str_replace(Species,"Loiseleuria procumbens", "Kalmia procumbens")) %>%
  mutate(Species=str_replace(Species,"Mycelis muralis", "Lactuca muralis")) %>%
  mutate(Species=str_replace(Species,"Omalotheca supina", "Gnaphalium supinum")) %>%
  mutate(Species=str_replace(Species,"Omalotheca norvegica", "Gnaphalium norvegicum")) %>%
  mutate(Species=str_replace(Species,"Omalotheca sylvatica", "Gnaphalium sylvaticum")) %>%
  mutate(Species=str_replace(Species,"Oreopteris limbosperma", "Thelypteris limbosperma")) %>%
  mutate(Species=str_replace(Species,"Oxycoccus microcarpus", "Vaccinium microcarpum")) %>%
  mutate(Species=str_replace(Species,"Oxycoccus palustris", "Vaccinium oxycoccos")) %>%
  mutate(Species=str_replace(Species,"Phalaris minor", "Phalaris arundinacea")) %>%
  mutate(Species=str_replace(Species,"Pinus unicinata", "Pinus mugo")) %>%
  mutate(Species=str_replace(Species,"Poa alpigena", "Poa pratensis")) %>%
  mutate(Species=str_replace(Species,"Poa angustifolia", "Poa pratensis")) %>%
  mutate(Species=str_replace(Species,"Pyrola grandiflora", "Pyrola rotundifolia")) %>%
  mutate(Species=str_replace(Species,"Rumex alpestris", "Rumex acetosa")) %>%
  mutate(Species=str_replace(Species,"Syringa emodi", "Syringa vulgaris")) %>%
  mutate(Species=str_replace(Species,"Taraxacum crocea", "Taraxacum officinale")) %>%
  mutate(Species=str_replace(Species,"Taraxacum croceum", "Taraxacum officinale")) %>%
  mutate(Species=str_replace(Species,"Trientalis europaea", "Lysimachia europaea")) %>%
  mutate(Species=str_replace(Species,"Trifolium pallidum", "Trifolium pratense"))

## merge species data with indicators
ANO.sp.ind <- merge(x=ANO.sp[,c("Species", "art_dekning", "ParentGlobalID")], 
                    y= ind.dat[,c("species", "Light", "Moisture", "Soil_reaction_pH", "Nitrogen")],
                    by.x="Species", by.y="species", all.x=T)
summary(ANO.sp.ind)
# checking which species didn't find a match
unique(ANO.sp.ind[is.na(ANO.sp.ind$Moisture),'Species'])
# don't find synonyms for these in the ind lists

## trimming away the points without information on NiN, species or cover
ANO.sp.ind <- ANO.sp.ind[!is.na(ANO.sp.ind$Species),]
ANO.sp.ind <- ANO.sp.ind[!is.na(ANO.sp.ind$art_dekning),]

summary(ANO.sp.ind)
head(ANO.sp.ind)
rm(ANO.sp)

### NiN reference data - data handling

## generalized species lists from NiN
str(Eco_State)

# species
Eco_State$Concept_Data$Species$Species_List$species
# environments
t(Eco_State$Concept_Data$Env$Env_Data)
# abundances
t(Eco_State$Concept_Data$Species$Species_Data)

## transposing abundance data
NiN.sp <- t(Eco_State$Concept_Data$Species$Species_Data)
NiN.sp <- as.data.frame(NiN.sp)
NiN.sp$sp <- as.factor(as.vector(Eco_State$Concept_Data$Species$Species_List$species))
# only genus and species name
NiN.sp$sp <- word(NiN.sp$sp, 1,2)
NiN.sp$spgr <- as.factor(as.vector(Eco_State$Concept_Data$Species$Species_List$art.code))
# if relevant, trimming to desired species groups (for forests e.g. removing trees)
#NiN.sp <- NiN.sp[NiN.sp$spgr!="a1a",]

## environment data
NiN.env <- Eco_State$Concept_Data$Env$Env_Data

## merging with indicator values
NiN.sp.ind <- merge(NiN.sp,ind.dat, by.x="sp", by.y="species", all.x=T)
summary(NiN.sp.ind)

NiN.sp.ind[NiN.sp.ind==999] <- NA

## checking which species didn't find a match
unique(NiN.sp.ind[is.na(NiN.sp.ind$Moisture) & NiN.sp.ind$spgr %in% list("a1a","a1b","a1c"),'sp'])

## fix species name issues
ind.dat <- ind.dat %>% 
  mutate(species=str_replace(species,"Aconitum lycoctonum", "Aconitum septentrionale")) %>% 
  mutate(species=str_replace(species,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  mutate(species=str_replace(species,"Carex myosuroides", "Kobresia myosuroides")) %>%
  mutate(species=str_replace(species,"Clinopodium acinos", "Acinos arvensis")) %>%
  mutate(species=str_replace(species,"Artemisia rupestris", "Artemisia norvegica")) %>%
  mutate(species=str_replace(species,"Cherleria biflora", "Minuartia biflora"))

NiN.sp <- NiN.sp %>% 
  mutate(sp=str_replace(sp,"Aconitum lycoctonum", "Aconitum septentrionale")) %>% 
  mutate(sp=str_replace(sp,"Anagallis minima", "Lysimachia minima")) %>% 
  mutate(sp=str_replace(sp,"Arctous alpinus", "Arctous alpina")) %>%
  mutate(sp=str_replace(sp,"Betula tortuosa", "Betula pubescens")) %>%
  mutate(sp=str_replace(sp,"Blysmopsis rufa", "Blysmus rufus")) %>%
  mutate(sp=str_replace(sp,"Cardamine nymanii", "Cardamine pratensis")) %>%
  mutate(sp=str_replace(sp,"Carex adelostoma", "Carex buxbaumii")) %>%
  mutate(sp=str_replace(sp,"Carex leersii", "Carex echinata")) %>%
  mutate(sp=str_replace(sp,"Carex paupercula", "Carex magellanica")) %>%
  mutate(sp=str_replace(sp,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  mutate(sp=str_replace(sp,"Carex _vacillans", "Carex vacillans")) %>%
  mutate(sp=str_replace(sp,"Carex viridula", "Carex flava")) %>%
  mutate(sp=str_replace(sp,"Chamaepericlymenum suecicum", "Cornus suecia")) %>%
  mutate(sp=str_replace(sp,"Cornus suecia", "Cornus suecica")) %>%
  mutate(sp=str_replace(sp,"Cicerbita alpina", "Lactuca alpina")) %>%
  mutate(sp=str_replace(sp,"Dactylorhiza sphagnicola", "Dactylorhiza majalis")) %>%
  mutate(sp=str_replace(sp,"Diphasiastrum complanatum", "Lycopodium complanatum")) %>%
  mutate(sp=str_replace(sp,"Elymus alaskanus", "Elymus kronokensis")) %>%
  mutate(sp=str_replace(sp,"Empetrum hermaphroditum", "Empetrum nigrum")) %>%
  mutate(sp=str_replace(sp,"Erigeron eriocephalus", "Erigeron uniflorus")) %>%
  mutate(sp=str_replace(sp,"Festuca prolifera", "Festuca rubra")) %>%
  mutate(sp=str_replace(sp,"Galium album", "Galium mollugo")) %>%
  mutate(sp=str_replace(sp,"Galium elongatum", "Galium palustre")) %>%
  mutate(sp=str_replace(sp,"Glaux maritima", "Lysimachia maritima")) %>%
  mutate(sp=str_replace(sp,"Helictotrichon pratense", "Avenula pratensis")) %>%
  mutate(sp=str_replace(sp,"Helictotrichon pubescens", "Avenula pubescens")) %>%
  mutate(sp=str_replace(sp,"Hieracium alpina", "Hieracium Alpina")) %>%
  mutate(sp=str_replace(sp,"Hieracium alpinum", "Hieracium Alpina")) %>%
  mutate(sp=str_replace(sp,"Hieracium aurantiacum", "Pilosella aurantiaca")) %>%
  mutate(sp=str_replace(sp,"Hieracium hieracium", "Hieracium Hieracium")) %>%
  mutate(sp=str_replace(sp,"Hieracium hieracioides", "Hieracium umbellatum")) %>%
  mutate(sp=str_replace(sp,"Hieracium lactucella", "Pilosella lactucella")) %>%
  mutate(sp=str_replace(sp,"Hieracium murorum", "Hieracium Vulgata")) %>%
  mutate(sp=str_replace(sp,"Hieracium oreadea", "Hieracium Oreadea")) %>%
  mutate(sp=str_replace(sp,"Hieracium prenanthoidea", "Hieracium Prenanthoidea")) %>%
  mutate(sp=str_replace(sp,"Hieracium vulgata", "Hieracium Vulgata")) %>%
  mutate(sp=str_replace(sp,"Hieracium pilosella", "Pilosella officinarum")) %>%
  mutate(sp=str_replace(sp,"Hieracium vulgatum", "Hieracium umbellatum")) %>%
  mutate(sp=str_replace(sp,"Hierochloã« alpina", "Hierochloë alpina")) %>%
  mutate(sp=str_replace(sp,"Hierochloã« hirta", "Hierochloë hirta")) %>%
  mutate(sp=str_replace(sp,"Hierochloã« odorata", "Hierochloë odorata")) %>%
  mutate(sp=str_replace(sp,"Huperzia appressa", "Huperzia selago")) %>%
  mutate(sp=str_replace(sp,"Hylotelephium maximum", "Hylotelephium telephium")) %>%
  mutate(sp=str_replace(sp,"Lappula myosotis", "Lappula squarrosa")) %>%
  mutate(sp=str_replace(sp,"Lepidotheca suaveolens", "Matricaria discoidea")) %>%
  mutate(sp=str_replace(sp,"Listera cordata", "Neottia cordata")) %>%
  mutate(sp=str_replace(sp,"Leontodon autumnalis", "Scorzoneroides autumnalis")) %>%
  mutate(sp=str_replace(sp,"Loiseleuria procumbens", "Kalmia procumbens")) %>%
  mutate(sp=str_replace(sp,"Logfia arvensis", "Filago arvensis")) %>%
  mutate(sp=str_replace(sp,"Mentha _verticillata", "Mentha verticillata")) %>%
  mutate(sp=str_replace(sp,"Minuartia rubella", "Sabulina rubella")) %>%
  mutate(sp=str_replace(sp,"Minuartia stricta", "Sabulina stricta")) %>%
  mutate(sp=str_replace(sp,"Mycelis muralis", "Lactuca muralis")) %>%
  mutate(sp=str_replace(sp,"Omalotheca supina", "Gnaphalium supinum")) %>%
  mutate(sp=str_replace(sp,"Omalotheca norvegica", "Gnaphalium norvegicum")) %>%
  mutate(sp=str_replace(sp,"Omalotheca sylvatica", "Gnaphalium sylvaticum")) %>%
  mutate(sp=str_replace(sp,"Ononis arvensis", "Ononis spinosa")) %>%
  mutate(sp=str_replace(sp,"Oreopteris limbosperma", "Thelypteris limbosperma")) %>%
  mutate(sp=str_replace(sp,"Oxycoccus microcarpus", "Vaccinium microcarpum")) %>%
  mutate(sp=str_replace(sp,"Oxycoccus palustris", "Vaccinium oxycoccos")) %>%
  mutate(sp=str_replace(sp,"Phalaris minor", "Phalaris arundinacea")) %>%
  mutate(sp=str_replace(sp,"Phalaroides arundinacea", "Phalaris arundinacea")) %>%
  mutate(sp=str_replace(sp,"Pinus unicinata", "Pinus mugo")) %>%
  mutate(sp=str_replace(sp,"Platanthera montana", "Platanthera chlorantha")) %>%
  mutate(sp=str_replace(sp,"Poa alpigena", "Poa pratensis")) %>%
  mutate(sp=str_replace(sp,"Poa angustifolia", "Poa pratensis")) %>%
  mutate(sp=str_replace(sp,"Poa laxa", "Poa flexuosa")) %>%
  mutate(sp=str_replace(sp,"Poa _herjedalica", "Poa herjedalica")) %>%
  mutate(sp=str_replace(sp,"Poa _jemtlandica", "Poa jemtlandica")) %>%
  mutate(sp=str_replace(sp,"Poa lindebergii", "Poa arctica")) %>%
  mutate(sp=str_replace(sp,"Pyrola grandiflora", "Pyrola rotundifolia")) %>%
  mutate(sp=str_replace(sp,"Rhamnus catharticus", "Rhamnus cathartica")) %>%
  mutate(sp=str_replace(sp,"Rumex alpestris", "Rumex acetosa")) %>%
  mutate(sp=str_replace(sp,"Salix _fragilis", "Salix fragilis")) %>%
  mutate(sp=str_replace(sp,"Saxifraga _opdalensis", "Saxifraga opdalensis")) %>%
  mutate(sp=str_replace(sp,"Spergularia salina", "Spergularia marina")) %>%
  mutate(sp=str_replace(sp,"Syringa emodi", "Syringa vulgaris")) %>%
  mutate(sp=str_replace(sp,"Taraxacum crocea", "Taraxacum officinale")) %>%
  mutate(sp=str_replace(sp,"Taraxacum croceum", "Taraxacum officinale")) %>%
  mutate(sp=str_replace(sp,"Taraxacum erythrospermum", "Taraxacum officinale")) %>%
  mutate(sp=str_replace(sp,"Taraxacum hamatum", "Taraxacum officinale")) %>%
  mutate(sp=str_replace(sp,"Trientalis europaea", "Lysimachia europaea")) %>%
  mutate(sp=str_replace(sp,"Trifolium pallidum", "Trifolium pratense")) %>%
  mutate(sp=str_replace(sp,"Vicia orobus", "Vicia cassubica"))

## merge species data with indicators
NiN.sp.ind <- merge(NiN.sp,ind.dat, by.x="sp", by.y="species", all.x=T)
summary(NiN.sp.ind)

NiN.sp.ind[NiN.sp.ind==999] <- NA

# checking which species didn't find a match
unique(NiN.sp.ind[is.na(NiN.sp.ind$Moisture) & NiN.sp.ind$spgr %in% list("a1a","a1b","a1c"),'sp'])
# ok now

## matching with NiN ecosystem types - for wetlands
# NB! beware of rogue spaces in the 'Nature_type' & 'Sub_Type' variables, e.g. "Spring_Forest "
NiN.wetland <- NiN.sp.ind[,c("sp",paste(NiN.env[NiN.env$Nature_Type=="Mire","ID"]),colnames(ind.dat)[15:18])]   # Light, Moisture, Soil_reaction_pH, Nitrogen
NiN.wetland[1,]
names(NiN.wetland)

cbind(colnames(NiN.wetland),
      c("",
        'V3-C1a','V3-C1b','V3-C1c','V3-C1d','V3-C1e',
        'V1-C1a','V1-C1b','V1-C1c','V1-C1d','V1-C1e',
        'V1-C2a','V1-C2b','V1-C2c','V1-C2d',
        'V1-C3a','V1-C3b','V1-C3c','V1-C3d',
        'V1-C4a','V1-C4b','V1-C4c','V1-C4d',
        'V1-C4e','V1-C4f','V1-C4g','V1-C4h',
        'V3-C2','V1-C5',
        'V1-C6a','V1-C6b',
        'V1-C7a','V1-C7b',
        'V1-C8a','V1-C8b',
        'V2-C1a','V2-C1b',
        'V2-C2a','V2-C2b',
        'V2-C3a','V2-C3b',
        "V4-C2","V4-C3",
        "","",
        'V8-C1','V8-C2','V8-C3',
        rep("",10),
        rep("",4)
      )
)

NiN.wetland <- NiN.wetland[,c(1:43,46:48,59:62)]
colnames(NiN.wetland)[2:46] <- c('V3-C1a','V3-C1b','V3-C1c','V3-C1d','V3-C1e',
                              'V1-C1a','V1-C1b','V1-C1c','V1-C1d','V1-C1e',
                              'V1-C2a','V1-C2b','V1-C2c','V1-C2d',
                              'V1-C3a','V1-C3b','V1-C3c','V1-C3d',
                              'V1-C4a','V1-C4b','V1-C4c','V1-C4d',
                              'V1-C4e','V1-C4f','V1-C4g','V1-C4h',
                              'V3-C2','V1-C5',
                              'V1-C6a','V1-C6b',
                              'V1-C7a','V1-C7b',
                              'V1-C8a','V1-C8b',
                              'V2-C1a','V2-C1b',
                              'V2-C2a','V2-C2b',
                              'V2-C3a','V2-C3b',
                              'V4-C2','V4-C3',
                              'V8-C1','V8-C2','V8-C3'
)
head(NiN.wetland)

# translating the abundance classes into %-cover
coverscale <- data.frame(orig=0:6,
                         cov=c(0,1/32,1/8,3/8,0.6,4/5,1)
)

NiN.wetland.cov <- NiN.wetland
colnames(NiN.wetland.cov)
for (i in 2:46) {
  NiN.wetland.cov[,i] <- coverscale[,2][ match(NiN.wetland[,i], 0:6 ) ]
}

NiN.wetland.cov$sp <- as.factor(NiN.wetland.cov$sp)
```
leaving us with the monitoring data including plant indicators (ANO.sp.ind) and the reference data including plant indicators (NiN.wetland.cov)

```{r}
head(ANO.sp.ind)
head(NiN.wetland.cov)
```
For each ecosystem type with a NiN species list, we can calculate a community weighted mean (CWM) for the relevant functional plant indicators.
For wetland ecosystem we are testing "Light", "Moisture", "Soil_reaction_pH", and "Nitrogen". In order to get distributions of CWMs rather than one single value (for comparison with the empirical testing data) the NiN lists can be bootstrapped.

##### bootstrap function for frequency abundance
- function to calculate community weighted means of selected indicator values (ind)
- for species lists (sp) with given abundances in percent (or on a scale from 0 to 1) in one or more 'sites' (abun)
- with a given number of iterations (iter),
- with species given a certain minimum abundance occurring in all bootstraps (obl), and
- with a given re-sampling ratio of the original species list (rat)
- in every bootstrap iteration the abundance of the sampled species can be randomly changed by a limited amount if wished by introducing a re-sampling of abundance values from adjacent abundance steps with a certain probability (var.abun)

```{r, echo = F}
indBoot.freq <- function(sp,abun,ind,iter,obl,rat=2/3,var.abun=F) {

  ind.b <- matrix(nrow=iter,ncol=length(colnames(abun)))
  colnames(ind.b) <- colnames(abun)
  ind.b <- as.data.frame(ind.b)

  ind <- as.data.frame(ind)
  ind.list <- as.list(1:length(colnames(ind)))
  names(ind.list) <- colnames(ind)

  for (k in 1:length(colnames(ind)) ) {
    ind.list[[k]] <- ind.b }

  for (j in 1:length(colnames(abun)) ) {

    dat <- cbind(sp,abun[,j],ind)
    dat <- dat[dat[,2]>0,]            # only species that are present in the ecosystem
    dat <- dat[!is.na(dat[,3]),]      # only species that have indicator values

    for (i in 1:iter) {

      speciesSample <- sample(dat$sp[dat[,2] < obl], size=round( (length(dat$sp)-length(dat$sp[dat[,2]>=obl])) *rat,0), replace=F)  
      dat.b <- rbind(dat[dat[,2] >= obl,],
                     dat[match(speciesSample,dat$sp),]
      )

      if (var.abun==T) {
        for (m in 1:nrow(coverscale[-1,]) ) {
          xxx <- dat.b[dat.b[,2]==coverscale[-1,][m,2],2]
          if ( m==1 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.5, 0.5, 0.0, 0.0, 0.0, 0.0, 0.0) ,size=length(xxx),replace=T) }
          if ( m==2 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.2, 0.3, 0.5, 0.0, 0.0, 0.0, 0.0) ,size=length(xxx),replace=T) }
          if ( m==3 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.0, 0.2, 0.3, 0.5, 0.0, 0.0, 0.0) ,size=length(xxx),replace=T) }
          if ( m==4 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.0, 0.0, 0.2, 0.3, 0.5, 0.0, 0.0) ,size=length(xxx),replace=T) }
          if ( m==5 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.0, 0.0, 0.0, 0.2, 0.3, 0.5, 0.0) ,size=length(xxx),replace=T) }
          if ( m==6 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.0, 0.0, 0.0, 0.0, 0.2, 0.3, 0.5) ,size=length(xxx),replace=T) }
        }
        dat.b[!is.na(dat.b[,2]) & dat.b[,2]<=(0),2] <- 0.01
        dat.b[!is.na(dat.b[,2]) & dat.b[,2]>1,2] <- 1
      }

      for (k in 1:length(colnames(ind))) {

        if ( nrow(dat.b)>2 ) {

          ind.b <- sum(dat.b[!is.na(dat.b[,2+k]),2] * dat.b[!is.na(dat.b[,2+k]),2+k] , na.rm=T) / sum(dat.b[!is.na(dat.b[,2+k]),2],na.rm=T)
          ind.list[[k]][i,j] <- ind.b

        } else {ind.list[[k]][i,j] <- NA}

      }

#      print(paste(i,"",j)) 
    }

  }
  return(ind.list)
}
```

Running the bootstraps

```{r, echo = T, eval = F}
colnames(NiN.wetland)
# 1st column is the species
# 2nd-46th column is the abundances of sp in different ecosystem types
# 47th-50th column is the indicator values of the respective species
# we choose 1000 iterations
# species with abundance 1 (i.e. a max of 100%, must be included in each sample)
# each sample re-samples 1/3 of the number of species
# the abundance of the re-sampled species may vary (see bootstrap function for details)
wetland.ref.cov <- indBoot.freq(sp=NiN.wetland.cov[,1],abun=NiN.wetland.cov[,2:46],ind=NiN.wetland.cov[,47:50],
                          iter=1000,obl=1,rat=1/3,var.abun=T)

# fixing NaNs
for (i in 1:length(wetland.ref.cov) ) {
  for (j in 1:ncol(wetland.ref.cov[[i]]) ) {
    v <- wetland.ref.cov[[i]][,j]
    v[is.nan(v)] <- NA
    wetland.ref.cov[[i]][,j] <- v
  }
}
```

```{r, include = F}
# Data from cache
wetland.ref.cov<-readRDS(paste0(here::here(), "/data/cache/wetland.ref.cov.RDS"))

#load("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/FPI_output large files for markdown/ref_lists_wetland.RData")
```

```{r}
head(wetland.ref.cov[[1]])
```

This results in an R-list with a slot for every selected indicator, and in every slot there's a data frame with as many columns as there are NiN species lists and as many rows as there were iterations in the bootstrap.
Next, we need to derive scaling values from these bootstrap-lists (the columns) for every mapping unit in NiN. Here, we define things in the following way:

- Median = reference values
- 0.025 and 0.975 quantiles = lower and upper limit values
- min and max of the respective indicator's scale = min/max values

```{r, echo = F}

# every NiN-type is represented by one 'generalisert artsliste'
# some NiN-types are represented by two such species lists
# in some cases two NiN-types are represented by the same species list
head(wetland.ref.cov[[1]])
wetland.ref.cov[[1]][0,]

# checking the actual NiN-types in the wetland lists
wetland.NiNtypes <- colnames(wetland.ref.cov[["Light"]])
wetland.NiNtypes <- substr(wetland.NiNtypes,1,5)
unique(wetland.NiNtypes)

# 4 indicator-value indicators: Tyler's Light, Moisture, Soil_reaction_pH, "Nitrogen"
indEll.n=4
# creating a table to hold:
# Tyler: the 0.5 quantile (median), 0.05 quantile and  0.95 quantile for each NiN-type
# for every nature type (nrows)
tab <- matrix(ncol=3*indEll.n, nrow=length(unique(wetland.NiNtypes)) ) # 43 basic ecosystem types
# coercing the values into the table
# NiN-types where each type is represented by one species list (including when one species list represents two NiN-types)
names(wetland.ref.cov[["Light"]])
x <- c(27,28,41:45)

for (i in 1:length(x) ) {
  tab[i,1:3] <- quantile(as.matrix(wetland.ref.cov[["Light"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[i,4:6] <- quantile(as.matrix(wetland.ref.cov[["Moisture"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[i,7:9] <- quantile(as.matrix(wetland.ref.cov[["Soil_reaction_pH"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[i,10:12] <- quantile(as.matrix(wetland.ref.cov[["Nitrogen"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)
}

tab <- as.data.frame(tab)
tab$NiN <- NA
tab$NiN[1:length(x)] <- names(wetland.ref.cov[[1]])[x]

# NiN-types represented by several species lists
wetland.NiNtypes2 <- wetland.NiNtypes[-x]
unique(wetland.NiNtypes2)
grep(pattern=unique(wetland.NiNtypes2)[1], x=wetland.NiNtypes) # finds columns in e.g. colnames(wetland.ref.cov[["Continentality"]]) that match the first NiN-type

for (i in 1:length(unique(wetland.NiNtypes2)) ) {
  tab[length(x)+i,1:3] <- quantile(as.matrix(wetland.ref.cov[["Light"]][,grep(pattern=unique(wetland.NiNtypes2)[i], x=wetland.NiNtypes)]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[length(x)+i,4:6] <- quantile(as.matrix(wetland.ref.cov[["Moisture"]][,grep(pattern=unique(wetland.NiNtypes2)[i], x=wetland.NiNtypes)]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[length(x)+i,7:9] <- quantile(as.matrix(wetland.ref.cov[["Soil_reaction_pH"]][,grep(pattern=unique(wetland.NiNtypes2)[i], x=wetland.NiNtypes)]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[length(x)+i,10:12] <- quantile(as.matrix(wetland.ref.cov[["Nitrogen"]][,grep(pattern=unique(wetland.NiNtypes2)[i], x=wetland.NiNtypes)]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab$NiN[length(x)+i] <- unique(wetland.NiNtypes2)[i]

}

tab

# making it a proper data frame
round(tab[,1:12],digits=2)

colnames(tab) <- c("Light_q2.5","Light_q50","Light_q97.5",
                   "Moist_q2.5","Moist_q50","Moist_q97.5",
                   "pH_q2.5","pH_q50","pH_q97.5",
                   "Nitrogen_q2.5","Nitrogen_q50","Nitrogen_q97.5",
                   "NiN")
summary(tab)
tab$NiN <- gsub("C", "C-", tab$NiN) # add extra hyphon after C for NiN-types
tab

# restructuring into separate indicators for lower (q2.5) and higher (q97.5) than reference value (=median, q50)
y.Light <- numeric(length=nrow(tab)*2)
y.Light[((1:dim(tab)[1])*2)-1] <- tab$Light_q2.5 
y.Light[((1:dim(tab)[1])*2)] <- tab$Light_q97.5

y.Moist <- numeric(length=nrow(tab)*2)
y.Moist[((1:dim(tab)[1])*2)-1] <- tab$Moist_q2.5 
y.Moist[((1:dim(tab)[1])*2)] <- tab$Moist_q97.5

y.pH <- numeric(length=nrow(tab)*2)
y.pH[((1:dim(tab)[1])*2)-1] <- tab$pH_q2.5 
y.pH[((1:dim(tab)[1])*2)] <- tab$pH_q97.5

y.Nitrogen <- numeric(length=nrow(tab)*2)
y.Nitrogen[((1:dim(tab)[1])*2)-1] <- tab$Nitrogen_q2.5 
y.Nitrogen[((1:dim(tab)[1])*2)] <- tab$Nitrogen_q97.5

# creating final objects holding the reference and limit values for all indicators

# ref object for indicators
wetland.ref.cov.val <- data.frame(N1=rep('wetland',(nrow(tab)*2*indEll.n)),
                              hoved=c(rep('NA',(nrow(tab)*2*indEll.n))),
                              grunn=c(rep(rep(tab$NiN,each=2),indEll.n)),
                              county=rep('all',(nrow(tab)*2*indEll.n)),
                              region=rep('all',(nrow(tab)*2*indEll.n)),
                              Ind=c(rep(c('Light1','Light2'),nrow(tab)),
                                    rep(c('Moist1','Moist2'),nrow(tab)),
                                    rep(c('pH1','pH2'),nrow(tab)),
                                    rep(c('Nitrogen1','Nitrogen2'),nrow(tab))
                              ),
                              Rv=c(rep(tab$Light_q50,each=2),
                                   rep(tab$Moist_q50,each=2),
                                   rep(tab$pH_q50,each=2),
                                   rep(tab$Nitrogen_q50,each=2)
                              ),
                              Gv=c(y.Light,y.Moist,y.pH,y.Nitrogen),
                              maxmin=c(rep(c(1,7),nrow(tab)), # 7 levels of light
                                       rep(c(1,12),nrow(tab)), # 12 levels of moisture
                                       rep(c(1,8),nrow(tab)), # 8 levels of soil reaction pH
                                       rep(c(1,9),nrow(tab))  # 9 levels of nitrogen
                              )
)

wetland.ref.cov.val
wetland.ref.cov.val$grunn <- as.factor(wetland.ref.cov.val$grunn)
wetland.ref.cov.val$Ind <- as.factor(wetland.ref.cov.val$Ind)
summary(wetland.ref.cov.val)

```

```{r}
head(wetland.ref.cov.val)
```

Once test data (ANO) and the scaling values from the reference data (NiN) are in place, we can calculate CWMs of the selected indicators for the ANO community data and scale them against the scaling values from the reference distribution. Note that we scale each ANO plot's CWM against either the lower threshold value and the min value OR the upper threshold value and the max value based on whether the CWM is smaller or higher than the reference value. Since the scaled values for both sides range between 0 and 1, we generate separate lower and upper condition indicators for each functional plant indicator. An ANO plot can only have a scaled value in either the lower or the upper indicator (the other one will be 'NA'), except for the unlikely event that the CWM exactly matches the reference value, in which case both lower and upper indicator will receive a scaled indicator value of 1.

Here is the scaling function
```{r}

#### scaled values ####
r.s <- 1    # reference value
l.s <- 0.6  # limit value
a.s <- 0    # abscence of indicator, or indicator at maximum

#### function for calculating scaled values for measured value ####

## scaling function including truncation
scal <- function() {
  # place to hold the result
   x <- numeric()
  if (maxmin < ref) {
    # values >= the reference value equal 1
    if (val >= ref) {x <- 1}
    # values < the reference value and >= the limit value can be deducted from the linear relationship between these two
    if (val < ref & val >= lim) {x <- (l.s + (val-lim) * ( (r.s-l.s) / (ref-lim) ) )}
    # values < the limit value and > maxmin can be deducted from the linear relationship between these two
    if (val < lim & val > maxmin) {x <- (a.s + (val-maxmin) * ( (l.s-a.s) / (lim-maxmin) ) )}
    # value equals or lower than maxmin
    if (val <= maxmin) {x <-0}
  } else {
    # values <= the reference value equal 1
    if (val <= ref) {x <- 1}
    # values > the reference value and <= the limit value can be deducted from the linear relationship between these two
    if (val > ref & val <= lim) {x <- ( r.s - ( (r.s - l.s) * (val - ref) / (lim - ref) ) )}
    # values > the limit value and < maxmin can be deducted from the linear relationship between these two
    if (val > lim) {x <- ( l.s - (l.s * (val - lim) / (maxmin - lim) ) )}
    # value equals or larger than maxmin
    if (val >= maxmin) {x <-0}
  }
  return(x)

}

```

We then can prepare a list of data frames to hold the results and perform the scaling according to the principles described in NINA report 1967 (Töpper and Jakobsson 2021)
```{r, include=FALSE}

unique(ANO.geo$hovedtype_rute) # NiN types in data
unique(substr(wetland.ref.cov.val$grunn,1,2)) # NiN types in reference
#### creating dataframe to hold the results for wetlands ####
# all ANO points
nrow(ANO.geo)
# all wetland ANO points
nrow(ANO.geo[ANO.geo$hovedtype_rute %in% list("V1","V2","V3","V4","V5","V6","V7","V8","V9","V10","V11","V12","V13"),])
# all wetland ANO points with a NiN-type represented in the reference
nrow(ANO.geo[ANO.geo$hovedtype_rute %in% unique(substr(wetland.ref.cov.val$grunn,1,2)),])
# ok, we'll be losing 70 (our of 1349) that are not covered by the reference
ANO.wetland <- ANO.geo[ANO.geo$hovedtype_rute %in% list("V1","V2","V3","V4","V5","V6","V7","V8","V9","V10","V11","V12","V13"),]

head(ANO.wetland)
# update row-numbers
row.names(ANO.wetland) <- 1:nrow(ANO.wetland)
head(ANO.wetland)
dim(ANO.wetland)
colnames(ANO.wetland)

length(levels(as.factor(ANO.wetland$ano_flate_id)))
length(levels(as.factor(ANO.wetland$ano_punkt_id)))
summary(as.factor(ANO.wetland$ano_punkt_id))
# four points that are double
ANO.wetland[ANO.wetland$ano_punkt_id=="ANO0159_55",] # double registration, said so in comment. -> choose row 207 over 206
ANO.wetland <- ANO.wetland[-206,]
row.names(ANO.wetland) <- 1:nrow(ANO.wetland) # update row-numbers
ANO.wetland[ANO.wetland$ano_punkt_id=="ANO0283_22",] # 2019 & 2021. Lot of NA's in 2019 -> omit 2019
ANO.wetland <- ANO.wetland[-156,]
row.names(ANO.wetland) <- 1:nrow(ANO.wetland) # update row-numbers
ANO.wetland[ANO.wetland$ano_punkt_id=="ANO0363_24",]
ANO.wetland[ANO.wetland$ano_flate_id=="ANO0363","ano_punkt_id"] # point-ID 15 is missing, but 24 is double. Likely that registrations are valid, but wrong point-ID.  -> keep both, call the second obs the one that's missing
ANO.wetland[311,"ano_punkt_id"] <- "ANO0363_15"
ANO.wetland[ANO.wetland$ano_punkt_id=="ANO1550_64",] # point-ID 66 is missing, but 64 is double. Likely that registrations are valid, but wrong point-ID.  -> keep both
ANO.wetland[ANO.wetland$ano_flate_id=="ANO1550","ano_punkt_id"] # point-ID 66 is missing, but 64 is double. Likely that registrations are valid, but wrong point-ID.  -> keep both
ANO.wetland[1273,"ano_punkt_id"] <- "ANO1550_66"

unique(ANO.wetland$hovedoekosystem_punkt)
unique(ANO.wetland$hovedtype_rute)
unique(ANO.wetland$kartleggingsenhet_1m2)
ANO.wetland$hovedtype_rute <- factor(ANO.wetland$hovedtype_rute)
ANO.wetland$kartleggingsenhet_1m2 <- factor(ANO.wetland$kartleggingsenhet_1m2)
summary(ANO.wetland$Hovedtype_rute)
summary(ANO.wetland$Kartleggingsenhet_rute)

results.wetland <- list()
ind <- unique(wetland.ref.cov.val$Ind)
# choose columns for site description
colnames(ANO.wetland)
results.wetland[['original']] <- ANO.wetland
# drop geometry
st_geometry(results.wetland[['original']]) <- NULL
results.wetland[['original']]

# add columns for indicators
nvar.site <- ncol(results.wetland[['original']])
for (i in 1:length(ind) ) {results.wetland[['original']][,i+nvar.site] <- NA}
colnames(results.wetland[['original']])[(nvar.site+1):(length(ind)+nvar.site)] <- paste(ind)
for (i in (nvar.site+1):(length(ind)+nvar.site) ) {results.wetland[['original']][,i] <- as.numeric(results.wetland[['original']][,i])}
summary(results.wetland[['original']])
#results.wetland[['original']]$Region <- as.factor(results.wetland[['original']]$Region)
results.wetland[['original']]$GlobalID <- as.factor(results.wetland[['original']]$GlobalID)
results.wetland[['original']]$ano_flate_id <- as.factor(results.wetland[['original']]$ano_flate_id)
results.wetland[['original']]$ano_punkt_id <- as.factor(results.wetland[['original']]$ano_punkt_id)
results.wetland[['original']]$hovedoekosystem_punkt <- as.factor(results.wetland[['original']]$hovedoekosystem_punkt)
#results.wetland[['original']]$Hovedoekosystem_rute  <- as.factor(results.wetland[['original']]$Hovedoekosystem_rute )
results.wetland[['original']]$kartleggingsenhet_1m2 <- as.factor(results.wetland[['original']]$kartleggingsenhet_1m2)
results.wetland[['original']]$hovedtype_rute    <- as.factor(results.wetland[['original']]$hovedtype_rute)

# roll out
results.wetland[['scaled']] <- results.wetland[['non-truncated']] <- results.wetland[['original']]
```

```{r, echo = T, eval = F}

#### calculating scaled and non-truncated values for the indicators based on the dataset ####
for (i in 1:nrow(ANO.wetland) ) {  #
  tryCatch({
    print(i)
    print(paste(ANO.wetland$ano_flate_id[i]))
    print(paste(ANO.wetland$ano_punkt_id[i]))
#    ANO.wetland$Hovedoekosystem_sirkel[i]
#    ANO.wetland$Hovedoekosystem_rute[i]

    # if the ANO.hovedtype exists in the reference
    if (ANO.wetland$hovedtype_rute[i] %in% unique(substr(wetland.ref.cov.val$grunn,1,2)) ) {

      # if there is any species present in current ANO point  
      if ( length(ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.wetland$GlobalID[i]),'Species']) > 0 ) {

          # Light
          dat <- ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.wetland$GlobalID[i]),c('art_dekning','Light')]
          results.wetland[['original']][i,'richness'] <- nrow(dat)
          dat <- dat[!is.na(dat$Light),]

          if ( nrow(dat)>0 ) {

            val <- sum(dat[,'art_dekning'] * dat[,'Light'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
            # lower part of distribution
            ref <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Light1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Light1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            maxmin <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Light1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.wetland dataframe
            results.wetland[['scaled']][i,'Light1'] <- scal() 
            results.wetland[['non-truncated']][i,'Light1'] <- scal.2() 
            results.wetland[['original']][i,'Light1'] <- val

            # upper part of distribution
            ref <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Light2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Light2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            maxmin <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Light2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.wetland dataframe
            results.wetland[['scaled']][i,'Light2'] <- scal() 
            results.wetland[['non-truncated']][i,'Light2'] <- scal.2() 
            results.wetland[['original']][i,'Light2'] <- val
          }

          # Moisture
          dat <- ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.wetland$GlobalID[i]),c('art_dekning','Moisture')]
          results.wetland[['original']][i,'richness'] <- nrow(dat)
          dat <- dat[!is.na(dat$Moisture),]

          if ( nrow(dat)>0 ) {

            val <- sum(dat[,'art_dekning'] * dat[,'Moisture'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
            # lower part of distribution
            ref <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Moist1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Moist1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            maxmin <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Moist1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.wetland dataframe
            results.wetland[['scaled']][i,'Moist1'] <- scal() 
            results.wetland[['non-truncated']][i,'Moist1'] <- scal.2() 
            results.wetland[['original']][i,'Moist1'] <- val

            # upper part of distribution
            ref <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Moist2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Moist2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            maxmin <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Moist2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.wetland dataframe
            results.wetland[['scaled']][i,'Moist2'] <- scal() 
            results.wetland[['non-truncated']][i,'Moist2'] <- scal.2() 
            results.wetland[['original']][i,'Moist2'] <- val
          }

          # Soil_reaction_pH
          dat <- ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.wetland$GlobalID[i]),c('art_dekning','Soil_reaction_pH')]
          results.wetland[['original']][i,'richness'] <- nrow(dat)
          dat <- dat[!is.na(dat$Soil_reaction_pH),]

          if ( nrow(dat)>0 ) {

            val <- sum(dat[,'art_dekning'] * dat[,'Soil_reaction_pH'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
            # lower part of distribution
            ref <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='pH1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='pH1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            maxmin <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='pH1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.wetland dataframe
            results.wetland[['scaled']][i,'pH1'] <- scal() 
            results.wetland[['non-truncated']][i,'pH1'] <- scal.2() 
            results.wetland[['original']][i,'pH1'] <- val

            # upper part of distribution
            ref <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='pH2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='pH2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            maxmin <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='pH2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.wetland dataframe
            results.wetland[['scaled']][i,'pH2'] <- scal() 
            results.wetland[['non-truncated']][i,'pH2'] <- scal.2() 
            results.wetland[['original']][i,'pH2'] <- val
          }

          # Nitrogen
          dat <- ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.wetland$GlobalID[i]),c('art_dekning','Nitrogen')]
          results.wetland[['original']][i,'richness'] <- nrow(dat)
          dat <- dat[!is.na(dat$Nitrogen),]

          if ( nrow(dat)>0 ) {

            val <- sum(dat[,'art_dekning'] * dat[,'Nitrogen'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
            # lower part of distribution
            ref <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Nitrogen1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Nitrogen1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            maxmin <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Nitrogen1' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.wetland dataframe
            results.wetland[['scaled']][i,'Nitrogen1'] <- scal() 
            results.wetland[['non-truncated']][i,'Nitrogen1'] <- scal.2() 
            results.wetland[['original']][i,'Nitrogen1'] <- val

            # upper part of distribution
            ref <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Nitrogen2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Nitrogen2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            maxmin <- wetland.ref.cov.val[wetland.ref.cov.val$Ind=='Nitrogen2' & wetland.ref.cov.val$grunn==as.character(results.wetland[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.wetland dataframe
            results.wetland[['scaled']][i,'Nitrogen2'] <- scal() 
            results.wetland[['non-truncated']][i,'Nitrogen2'] <- scal.2() 
            results.wetland[['original']][i,'Nitrogen2'] <- val
          }

        }
      }

  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

# for using both sides of the plant indicators
results.wetland[['2-sided']] <- results.wetland[['non-truncated']]

# remove values >1 for 2-sided indicators
results.wetland[['2-sided']]$Light1[results.wetland[['2-sided']]$Light1>1] <- NA
results.wetland[['2-sided']]$Light2[results.wetland[['2-sided']]$Light2>1] <- NA

results.wetland[['2-sided']]$Moist1[results.wetland[['2-sided']]$Moist1>1] <- NA
results.wetland[['2-sided']]$Moist2[results.wetland[['2-sided']]$Moist2>1] <- NA

results.wetland[['2-sided']]$pH1[results.wetland[['2-sided']]$pH1>1] <- NA
results.wetland[['2-sided']]$pH2[results.wetland[['2-sided']]$pH2>1] <- NA

results.wetland[['2-sided']]$Nitrogen1[results.wetland[['2-sided']]$Nitrogen1>1] <- NA
results.wetland[['2-sided']]$Nitrogen2[results.wetland[['2-sided']]$Nitrogen2>1] <- NA

```

```{r, include = F}
# Use cache data
results.wetland<-readRDS(paste0(here::here(),"/data/cache/results.wetland.RDS"))

#load("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/FPI_output large files for markdown/results.wetland.RData")
```

```{r}
head(results.wetland[['2-sided']])
```
#### Scaled value analyses

In order to visualize the results we need to rearrange the results-objects from wide to long format (note that there is both a lower and an upper condition indicator for each of the functional plant indicators).

```{r, message = FALSE}
#### plotting scaled values by main ecosystem type ####
## continuing with 2-sided
res.wetland <- results.wetland[['2-sided']]

# make long version of the scaled value part
res.wetland <-
  res.wetland %>% 
  pivot_longer(
    cols = c("Light1","Light2","Moist1","Moist2","pH1","pH2","Nitrogen1","Nitrogen2"),
    names_to = "fp_ind",
    values_to = "scaled_value",
    values_drop_na = FALSE
  )

# add original values as well
res.wetland <- 
  res.wetland %>% add_column(original = results.wetland[['original']] %>% 
                               pivot_longer(
                                 cols =  c("Light1","Light2","Moist1","Moist2","pH1","pH2","Nitrogen1","Nitrogen2"),
                                 names_to = NULL,
                                 values_to = "original",
                                 values_drop_na = FALSE
                               ) %>%
                               pull(original)
  )

```

### Ecosystem sub-types
And we can show the resulting scaled values as Violin plots for each indicator and main ecosystem type 
```{r,echo=FALSE,message=FALSE, warning=FALSE}

# making the plot
ggplot(res.wetland[!is.na(res.wetland$scaled_value),], aes(x=factor(hovedtype_rute), y=scaled_value, fill=fp_ind)) + 
  geom_hline(yintercept=0.6, linetype="dashed") + 
  geom_violin() +
#  geom_boxplot(width=0.2, color="grey") +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~factor(fp_ind,levels=c("Light1","Moist1","pH1","Nitrogen1","Light2","Moist2","pH2","Nitrogen2")), ncol = 4) + 
  xlab("Main ecosystem type") + 
  ylab("Scaled indicator value") 
```
The functional signature from the plant communities in wetland ecosystems looks largely good, except for V2 (swamp forests), which show rather low index scores for the lower pH and Nitrogen indicators, i.e. suggesting too acidic conditions and too low nitrogen availability.

We can look at V2 specifically to see if this is the case for all sub-types.

```{r}
# making the plot for V2 only
ggplot(res.wetland[res.wetland$hovedtype_rute=="V2",], aes(x=factor(kartleggingsenhet_1m2), y=scaled_value, fill=fp_ind)) + 
  geom_hline(yintercept=0.6, linetype="dashed") + 
  geom_violin() +
  #  geom_boxplot(width=0.2, color="grey") +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~factor(fp_ind,levels=c("Light1","Moist1","pH1","Nitrogen1","Light2","Moist2","pH2","Nitrogen2")), ncol = 4) + 
  xlab("Main ecosystem type") + 
  ylab("Scaled indicator value") 
```
The deviations towards low pH and low nitrogen are mainly an issue for V2-C1, limestone-poor swamp forests.

### Indicator index maps
We can also show the results as a map, for instance for pH1 (the lower pH indicator), either by directly plotting the data onto the map...
```{r, echo=FALSE, warning=FALSE,message = FALSE}
#### scaled value maps ####
# keep wide format and add geometry again
res.wetland2 <- results.wetland[['2-sided']]
st_geometry(res.wetland2) <- st_geometry(ANO.wetland)

#nor <- readRDS('P:/41201785_okologisk_tilstand_2022_2023/data/rds/norway_outline.RDS')%>%
#  st_as_sf() %>%
#  st_transform(crs = crs(ANO.geo))

nor <- st_read("data/outlineOfNorway_EPSG25833.shp")%>%
  st_as_sf() %>%
  st_transform(crs = crs(ANO.geo))

#reg <- st_read("P:/41201785_okologisk_tilstand_2022_2023/data/regioner/regNorway_wgs84 - MERGED.shp")%>%
#  st_as_sf() %>%
#  st_transform(crs = crs(ANO.geo))

reg <- st_read("data/regions.shp")%>%
  st_as_sf() %>%
  st_transform(crs = crs(ANO.geo))

# change region names to something R-friendly
reg$region
reg$region <- c("Northern Norway","Central Norway","Eastern Norway","Western Norway","Southern Norway")

regnor <- st_intersection(reg,nor)

## scaled value maps
# pH1 (lower indicator)
tm_shape(regnor) +
  tm_fill('GID_0', labels="", title="", legend.show = FALSE) + 
  tm_borders() +
  tm_shape(res.wetland2) +
  tm_dots('pH1',midpoint=NA, palette=tmaptools::get_brewer_pal("YlOrRd", 7, plot = FALSE), scale=1, legend.show = FALSE) + # 
  tm_layout(main.title = "pH (lower), wetland",legend.position = c("right", "bottom"), main.title.size=1.2) + 
  tm_add_legend(type = "fill", 
                col = c(tmaptools::get_brewer_pal("YlOrRd", 5, plot = FALSE),'grey'),
                labels = c("0.0 - 0.2", "0.2 - 0.4", "0.4 - 0.6", 
                           "0.6 - 0.8", "0.8 - 1.0", "NA"),
                title = "index values")

```
...but here the colors and values of the data points on the map are hard to make out.

### Regions - maps and statistics
Alternatively we can calculate and show the region-wise means and their related standard errors. But note that calculating a simple mean would be inappropriate for these data. This is because:
(i) the scaled data are bound between 0 and 1, and thus follow a beta-distribution rather than a Gaussian one
(ii) the ANO dataset has a nested structure

Therefore, we need to (i) use a beta-model, that (ii) can account for the nested structure of the data.
Here, we apply the following function using either a glmmTMB null-model with a beta-distribution, logit link, and the nesting as a random intercept, or a simple betareg null-model with logit link if the nesting is not extensive enough for a mixed model.
```{r}
library(betareg)
library(glmmTMB)

expit <- function(L) exp(L) / (1+exp(L)) # since the beta-models use a logit link, we need to calculate the estimates back to the identity scale

# the function performs a glmmTMB if there's >= 5 random levels in the nesting structure
# if that is not the case, then the function performs a betareg if theres >= 2 observations
# if that is not the case either, then the function returns the value of the single observation with NA standard error

indmean.beta <- function(df) {

  st_geometry(df) <- NULL
  colnames(df) <- c("y","ran")

  if ( nrow(df[!is.na(df[,1]),]) >= 2 ) {

    if ( length(unique(df[!is.na(df[,1]),2])) >=5 ) {

      mod1 <- glmmTMB(y ~ 1 +(1|ran), family=beta_family(), data=df)

      return(c(
        expit(summary( mod1 )$coefficients$cond[1]),

        expit( summary( mod1 )$coefficients$cond[1] + 
                 summary( mod1 )$coefficients$cond[2] )-
          expit( summary( mod1 )$coefficients$cond[1] ),

        nrow(df[!is.na(df$y),]),
        summary( mod1 )$coefficients$cond[1],
        summary( mod1 )$coefficients$cond[2]
      ))

    } else {

      mod2 <- betareg(y ~ 1, data=df)

      return(c(
        expit(summary( mod2 )$coefficients$mean[1]),
        expit( summary( mod2 )$coefficients$mean[1] + 
                 summary( mod2 )$coefficients$mean[2] )-
          expit( summary( mod2 )$coefficients$mean[1] ),
        nrow(df[!is.na(df$y),]),
        summary( mod2 )$coefficients$mean[1],
        summary( mod2 )$coefficients$mean[2]
      ))

    }

  } else {

    return(c(df$y,NA,1,NA,NA))

  }

}

```

```{r}

res.wetland2 = st_join(res.wetland2, regnor, left = TRUE)

regnor <- regnor %>%
  mutate(
    pH1.reg.mean = c(indmean.beta(df=res.wetland2[res.wetland2$region=="Northern Norway",c("pH1","ano_flate_id")])[1],
                                 indmean.beta(df=res.wetland2[res.wetland2$region=="Central Norway",c("pH1","ano_flate_id")])[1],
                                 indmean.beta(df=res.wetland2[res.wetland2$region=="Eastern Norway",c("pH1","ano_flate_id")])[1],
                                 indmean.beta(df=res.wetland2[res.wetland2$region=="Western Norway",c("pH1","ano_flate_id")])[1],
                                 indmean.beta(df=res.wetland2[res.wetland2$region=="Southern Norway",c("pH1","ano_flate_id")])[1]
    ),
    pH1.reg.se = c(indmean.beta(df=res.wetland2[res.wetland2$region=="Northern Norway",c("pH1","ano_flate_id")])[2],
                               indmean.beta(df=res.wetland2[res.wetland2$region=="Central Norway",c("pH1","ano_flate_id")])[2],
                               indmean.beta(df=res.wetland2[res.wetland2$region=="Eastern Norway",c("pH1","ano_flate_id")])[2],
                               indmean.beta(df=res.wetland2[res.wetland2$region=="Western Norway",c("pH1","ano_flate_id")])[2],
                               indmean.beta(df=res.wetland2[res.wetland2$region=="Southern Norway",c("pH1","ano_flate_id")])[2]
    ),
    pH1.reg.n = c(nrow(res.wetland2[res.wetland2$region=="Northern Norway" & !is.na(res.wetland2$pH1),]),
                              nrow(res.wetland2[res.wetland2$region=="Central Norway" & !is.na(res.wetland2$pH1),]),
                              nrow(res.wetland2[res.wetland2$region=="Eastern Norway" & !is.na(res.wetland2$pH1),]),
                              nrow(res.wetland2[res.wetland2$region=="Western Norway" & !is.na(res.wetland2$pH1),]),
                              nrow(res.wetland2[res.wetland2$region=="Southern Norway" & !is.na(res.wetland2$pH1),])
    )
  )

## scaled value maps for pH1 (lower indicator)
# mean
tm_shape(regnor) +
  tm_polygons(col="pH1.reg.mean", title="pH (lower), mean", style="quantile", palette=rev(get_brewer_pal(palette="OrRd", n=5, plot=FALSE))) +
  tm_text("pH1.reg.n",col="black",bg.color="grey")
```
Mean index value by region for the lower pH indicator (i.e. index shows deviations towards a more acidic environment). Numbers in grey fields show the number of observations in the respective region.

```{r}
# se
tm_shape(regnor) +
  tm_polygons(col="pH1.reg.se", title="pH (lower), sd", style="quantile", palette=get_brewer_pal(palette="OrRd", n=5, plot=FALSE)) +
  tm_text("pH1.reg.n",col="black",bg.color="grey")

```
Standard error to the mean index value by region for the lower pH indicator. Numbers in grey fields show the number of observations in the respective region.

### unscaled values vs. reference
We can also compare the unscaled values to the reference distribution in order to identify ecosystem types and functional plant indicators showing a deviation from the expectation. Since pH and nitrogen show some deviation of the lower pH indicator we exemplify this with these indicators for unscaled values.

```{r}
summary(res.wetland$kartleggingsenhet_1m2)
length(unique(res.wetland$kartleggingsenhet_1m2))
# 16 NiN-types to plot
colnames(wetland.ref.cov[['Soil_reaction_pH']])

### pH
par(mfrow=c(2,4))
## V1s
# V1-C-1
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V1-C1a","V1-C1b","V1-C1c","V1-C1d","V1-C1e")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C1',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original)),
       col="red")

# V1-C-2
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V1-C2a","V1-C2b","V1-C2c","V1-C2d")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C2',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-2",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-2",]$original)),
       col="red")

# V1-C-3
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V1-C3a","V1-C3b","V1-C3c","V1-C3d")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C3',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-3",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-3",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-3",]$original)),
       col="red")

# V1-C-4
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V1-C4a","V1-C4b","V1-C4c","V1-C4d","V1-C4e","V1-C4f","V1-C4g","V1-C4h")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C4',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-4",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-4",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-4",]$original)),
       col="red")

# V1-C-5
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V1-C5")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C5',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-5",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-5",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-5",]$original)),
       col="red")

# V1-C-6
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V1-C6a","V1-C6b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C6',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-6",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-6",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-6",]$original)),
       col="red")

# V1-C-7
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V1-C7a","V1-C7b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C7',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-7",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-7",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-7",]$original)),
       col="red")

# V1-C-8
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V1-C8a","V1-C8b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C8',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original)),
       col="red")

par(mfrow=c(2,4))
## V2s
# V2-C-1
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V2-C1a","V2-C1b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V2-C1',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original)),
       col="red")

# V2-C-2
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V2-C2a","V2-C2b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V2-C2',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-2",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-2",]$original)),
       col="red")

# V2-C-3
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V2-C3a","V2-C3b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V2-C3',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-3",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-3",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-3",]$original)),
       col="red")

## V3s
# V3-C-1
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V3-C1a","V3-C1b","V3-C1c","V3-C1d","V3-C1e")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V3-C1',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V3-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V3-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V3-C-1",]$original)),
       col="red")

# V3-C-2
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V3-C2")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V3-C2',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V3-C-2",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V3-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V3-C-2",]$original)),
       col="red")

## V4s
# V4-C-2
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V4-C2")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V4-C2',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V4-C-2",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V4-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V4-C-2",]$original)),
       col="red")

## V8s
# V8-C-1
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V8-C1")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V8-C1',xlab='pH value')
#points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-1",]$original,na.rm=T),
#       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-1",]$original)),
       col="red")

# V8-C-2
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V8-C2")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V8-C2',xlab='pH value')
#points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original,na.rm=T),
#       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original)),
       col="red")
legend("topleft", legend=c("reference","field data"), pch=c(NA,1), lty=1, col=c("black","red"))

### Nitrogen
par(mfrow=c(2,4))
## V1s
# V1-C-1
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C1a","V1-C1b","V1-C1c","V1-C1d","V1-C1e")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C1',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original)),
       col="red")

# V1-C-2
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C2a","V1-C2b","V1-C2c","V1-C2d")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C2',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-2",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-2",]$original)),
       col="red")

# V1-C-3
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C3a","V1-C3b","V1-C3c","V1-C3d")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C3',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-3",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-3",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-3",]$original)),
       col="red")

# V1-C-4
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C4a","V1-C4b","V1-C4c","V1-C4d","V1-C4e","V1-C4f","V1-C4g","V1-C4h")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C4',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-4",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-4",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-4",]$original)),
       col="red")

# V1-C-5
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C5")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C5',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-5",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-5",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-5",]$original)),
       col="red")

# V1-C-6
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C6a","V1-C6b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C6',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-6",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-6",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-6",]$original)),
       col="red")

# V1-C-7
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C7a","V1-C7b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C7',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-7",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-7",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-7",]$original)),
       col="red")

# V1-C-8
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C8a","V1-C8b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C8',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original)),
       col="red")

par(mfrow=c(2,4))
## V2s
# V2-C-1
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V2-C1a","V2-C1b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V2-C1',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original)),
       col="red")

# V2-C-2
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V2-C2a","V2-C2b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V2-C2',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-2",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-2",]$original)),
       col="red")

# V2-C-3
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V2-C3a","V2-C3b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V2-C3',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-3",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-3",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-3",]$original)),
       col="red")

## V3s
# V3-C-1
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V3-C1a","V3-C1b","V3-C1c","V3-C1d","V3-C1e")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V3-C1',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V3-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V3-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V3-C-1",]$original)),
       col="red")

# V3-C-2
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V3-C2")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V3-C2',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V3-C-2",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V3-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V3-C-2",]$original)),
       col="red")

## V4s
# V4-C-2
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V4-C2")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V4-C2',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V4-C-2",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V4-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V4-C-2",]$original)),
       col="red")

## V8s
# V8-C-1
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V8-C1")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V8-C1',xlab='Nitrogen value')
#points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V8-C-1",]$original,na.rm=T),
#       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V8-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V8-C-1",]$original)),
       col="red")

# V8-C-2
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V8-C2")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V8-C2',xlab='Nitrogen value')
#points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original,na.rm=T),
#       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original)),
       col="red")

legend("topleft", legend=c("reference","field data"), pch=c(NA,1), lty=1, col=c("black","red"))

```
THe figure shows that deviations mainly occur in V2-C1, which represents limestone-poor swamp forests. According to the functional signature from the plant community composition many of the occurrences of this ecosystem type are too acidic and may have too low availability of nitrogen.

### Eksport file (final product)
<!-- Export final file. Ideally a georeferenced shape or raster wit indicators values (raw and normalised), reference values and errors. -->