### 2023-08-28 DRAFT_functional_plant_indicators_natopen.Rmd
analysis bit of natopen markdown

# Functional plant indicators, naturally open ecosystems below the tree line (Grime's CSR values, Light, Nitrogen, Soil disturbance) {#Functional-plant-indicators-seminat}

<br />

_Author and date:_
Joachim Töpper

August 2023

<br />

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(readxl)
library(sf)
library(plyr)
library(tidyverse)
library(RColorBrewer)
library("gridExtra") 
library(ggridges)
library(tmap)
library(tmaptools)
library(raster)
library(stars)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
Ecosystem <- "naturally open below tree line"
Egenskap  <- c("Primærproduksjon","Abiotiske forhold")
ECT       <- "Functional state characteristic"
Contact   <- "Joachim Töpper"
```

```{r, echo=FALSE}
metaData <- data.frame(Ecosystem,
                       "Økologisk egenskap" = Egenskap,
                       "ECT class" = ECT)
knitr::kable(metaData)
```

<!-- Don't remove these three html lines -->
<br />
<br />
<hr />

## Introduction
Functional plant indicators can be used to describe the functional signature of plant communities by calculating community-weighted means of plant indicator values for plant communities (Diekmann 2003). The functional signature of plant communities may be indicative of ecosystem identity, depending on which functional plant indicator we look at (cf. Töpper et al. 2018). For instance, using an indicator for moisture one would find a functional signature with higher moisture values for plant communities in mires compared to e.g. grasslands or forests. Deviations in the functional signature of such an indicator beyond a certain range of indicator values (as there of course is natural variation of functional signatures within an ecosystem type) may be related to a reduction in ecological condition. Here, we combine functional plant indicator data with field sampled plant community data from the Norwegian nature monitoring programs ANO (Tingstad et al. 2019) and GRUK (Evju et al. 2020) for naturally open ecosystems below tree line (abbreviated as 'nat-open' henceforth). We calculate the functional signature of plant communities in the monitored sites with respect to Grime's CSR values, light, nitrogen, and soil disturbance. These functional signatures are then compared to reference distributions of functional signature, separately for each nat-open ecosystem type, calculated from 'generalized species lists' developed for ecosystem types in the Norwegian categorization system for eco-diversity (Halvorsen et al. 2020). These plant functional condition indicators are developed following the principles and technical protocol of the IBECA framework (Jakobsson et al. 2021, Töpper & Jakobsson 2021). Note that deviations from the reference may occur in both directions, e.g. the nitrogen signature from the testing data may be higher or lower than in the reference. Deviations in these two directions indicate very different environmental phenomena and thus have to be treated separately. Therefore, we develop two condition indicators for each functional plant indicator, a lower one and an upper one (see further down for more details).

## About the underlying data
In the 'functional plant indicator' project for nat-open ecosystems, we use five sets of data for building indicators for ecological condition:

- as test data we use plant community data from (1) the ANO monitoring scheme (cf. Tingstad et al. 2019) and (2) from the GRUK monitoring scheme (cf. Evju et al. 2020).
- as reference data we use (3) generalized species lists developed by Evju et al. 2023 for nat-open ecosystem types.
- (4) Swedish plant indicator data for light, nitrogen, ans soil disturbance from Tyler et al. (2021), and (5) Grime's CSR values for plant species' strategies (towards competition, stress, and ruderal conditions)

(1) ANO monitoring data:
ANO stands for 'areal-representativ naturovervåking', i.e. 'area representative nature monitoring'. 1000 sites are randomly distributed across mainland Norway and visited within a 5-year cycle. Each ANO site spans a 500 x 500 m grid cell, and the data collection at each ANO site includes up to 12 evenly distributed vegetation analyses in 1 x 1 m plots (up to 12, because some of these evenly distributed points may be in water or otherwise inaccessible). For the vegetation analyses, the cover of each vascular plant species in the plot is recorded. Every vegetation analysis is accompanied by an assessment of the ecosystem the plot lies in, including ecosystem type and some additional variables related to ecosystem-specific drivers of state. In the analysis in this document, we only use the plots which were classified as one of the nat-open ecosystem types in the Norwegian categorization system for eco-diversity (NiN) and which are not registered as lying in mountain areas above the tree line.
In the analysis in this document, we use the data available on Miljødirektoratets kartkatalog (https://kartkatalog.miljodirektoratet.no/Dataset/Details/2054), which comprises data from the first three ANO-years, 2019-2021, and a total of 8887 plots in 498 sites.

(2) GRUK monitoring data
GRUK stands for 'Grunnlendt åpen kalkmark', and represents a monitoring of open, limestone rich ecosystems with soils too shallow for forest establishment. The limestone rich focus of the scheme is motivated from a high biodiversity value and places all monitoring sites in an area around the Oslofjord. As of spring 2023, the GRUK data comprise 596 vegetation plots in 146 sites, all of which are mapped to the ecosystem type "åpen grunnlendt mark" (T2) and therein to the limestone-rich subtypes T2-C-7 and T2-C-8 in the Norwegian classification system for ecosystem types.
The vegetation analysis in GRUK record percent cover for every species of vascular plants in 1 sqm plots. In addition, a range of site variables related to ecosystem-specific drivers of state are recorded.
In the analysis in this document, we use all GRUK monitoring data data, which span the years 2020-22. The 2022-data also include an assessment of site condition, which is included in the analysis of results in this document. In the future, these data data may become available in Miljødirektoratets kartkatalog.

(3) NiN reference data:
The generalized species lists underlying the ecosystem categorization in NiN represent expert-compiled species lists based on empirical evidence from the literature and expert knowledge of the systems and their species. In these lists, every species is assigned an abundance value on a 6-step scale, with each step representing a range for the 'expected combination of frequency and cover' of occurrence (1: < 1/32, 2: 1/32 - 1/8 , 3: 1/8 - 3/8, 4: 3/8 - 4/5, 5: 3/8 - 4/5 + dominance, 6: > 4/5). For the purpose of this project, these steps are simplified to maximum 'expected combination of frequency and cover', whereby steps 4 & 5 are assigned 0.6 and 0.8, respectively, in order to distinguish between them.

(4) The Swedish plant indicator set published by Tyler et al. (2021) contains a large collection of plant indicators based on the Swedish flora, which is well representative of the Norwegian flora as well. From this set, we decided to include indicator data for light, moisture, pH, nitrogen, phosphorus, grazing_mowing, and soil disturbance for semi-natural ecosystems as these are thought to be subject to potential change due to abandonment, drainage/flooding, pollution, and erosion.

(5) Grime's system of plant strategy scores (Grime 1974) comprises relative (too one another) scores for the competition-, stress-, and disturbance("ruderality")-related life strategy of plant species. In the analysis, in this document we use all three variables, C, S and R, as different pressures acting on the ecosystem might change every one of the strategies (e.g. alien species for competition, climate change for stress, land use change for ruderality).

### Representativity in time and space
For nat-open ecosystems, the ANO data in this analysis contain 143 plots in 52 sites, in principle distributed randomly across the country. As nat-open ecosystems occur more often in certain regions of Norway than in others, the amount of plots and sites is not equal among Norway's five regions. The 143 plots are distributed across regions in the following way:

- Northern Norway: 29
- Central Norway:  39
- Eastern Norway:  34
- Western Norway:  24
- Southern Norway: 17

For GRUK, this analysis covers 1103 plots in 146 sites.

The 1103 plots are distributed across regions in the following way:

- Northern Norway: 0
- Central Norway:  0
- Eastern Norway:  1042
- Western Norway:  0
- Southern Norway: 61

### Temporal coverage
The ANO evaluation data cover the first three years, 2019-2021, of the first 5-year-cycle in the ANO monitoring scheme. GRUK covers 2020-2022. Thus, there is no actual time series to these data, and the indicator evaluation does therefore not include any temporal analyses.

## Collinearities with other indicators
<!-- Text here -->

## Reference state and values
### Reference state
The reference state is defined via the functional signature of the generalized species lists for NiN ecosystem types (see also Töpper et al. 2018). For the nat-open ecosystem types these lists have been newly prepared by Evju et al. 2023. By bootstrapping the species lists (see details further below) and calculating community-weighted means of functional plant indicators for every re-sampled community, we describe the reference state as a distribution of indicator values for each respective plant functional indicator. These distributions are calculated for ecosystem sub-types ("grunntyper" or "kartleggingsenheter" at a 1:5000 mapping scale) within the main wetland types in NiN. A more extensive discussion on the use of reference communities can be found in Jakobsson et al. (2020).

### Reference values, thresholds for defining _good ecological condition_, minimum and/or maximum values
In this analysis, we derive scaling values from statistical (here, non-parametric) distributions (see Jakobsson et al. 2010). For each ecosystem sub-type and plant functional indicator, the reference value is defined as the median value of the indicator value distribution. As in most cases the distributions naturally are two-sided (but see the Heat-requirement indicator in the mountain assessment for an example of a one-sided functional plant indicator, Fremstad et al. 2022), and deviation from the optimal state thus may occur in both direction (e.g. indicating too low or too high pH), we need to define two threshold values for good ecological condition as well as both a minimum and maximum value. In line with previous assessments of ecological condition for Norwegian forests and mountains, we define a lower and an upper threshold value via the 95% confidence interval of the reference distribution, i.e. its 0.025 and 0.975 quantiles. The minimum and maximum values are given by the minimum and maximum of the possible indicator values for each respective plant functional indicator. For details on the scaling principles in IBECA, please see Töpper & Jakobsson (2021).

## Uncertainties
We can calculate a mean indicator value (after scaling) for every region (or any other delimited area of interest) as well as its corresponding standard error as a measure of spatial uncertainty for a geographical area.

## References
Diekmann, M. 2003. Species indicator values as an important tool in applied plant ecology - a review. Basic and Applied Ecology 4: 493-506, doi:10.1078/1439-1791-00185

Evju, M., Stabbetorp, O.E., Olsen, S.L., Bratli, H., Often, A. & Bakkestuen, V. 2020. Dry calcareous grasslands in the Oslofjord region. A test of monitoring protocols and results for 2020. NINA Report 1910. Norwegian Institute for Nature Research.

Evju, M., Stabbetorp, O.E., Olsen, S.L., Bratli, ... . 2023. Generalized species lists for naturally open ecosystem types below the tree line in Norway. in prep.

Framstad, E., Kolstad, A. L., Nybø, S., Töpper, J. & Vandvik, V. 2022. The condition of forest and mountain ecosystems in Norway. Assessment by the IBECA method. NINA Report 2100. Norwegian Institute for Nature Research.

Grime J.P. 1974. Vegetation classification by reference to strategies. Nature 250(5461):26-31.

Halvorsen, R., Skarpaas, O., Bryn, A., Bratli, H., Erikstad, L., Simensen, T., & Lieungh, E. (2020). Towards a systematics of ecodiversity: The EcoSyst framework. Global Ecology and Biogeography, 29(11), 1887-1906. doi:10.1111/geb.13164

Jakobsson, S., Töpper, J.P., Evju, M., Framstad, E., Lyngstad, A., Pedersen, B., Sickel, H., Sverdrup-Thygeson, A., Vandvik. V., Velle, L.G., Aarrestad, P.A. & Nybø, S. 2020. Setting reference levels and limits for good ecological condition in terrestrial ecosystems. Insights from a case study based on the IBECA approach. Ecological Indicators 116: 106492.

Jakobsson, S., Evju, M., Framstad, E., Imbert, A., Lyngstad, A., Sickel, H., Sverdrup-Thygeson, A., Töpper, J., Vandvik, V., Velle, L.G., Aarrestad, P.A. & Nybø, S. 2021. An index-based assessment of ecological condition and its links to international frameworks. Ecological Indicators 124: 107252.

Tingstad, L., Evju, M., Sickel, H., & Töpper, J. 2019. Utvikling av nasjonal arealrepresentativ naturovervåking (ANO). Forslag til gjennomføring, protokoller og kostnadsvurderinger med utgangspunkt i erfaringer fra uttesting i Trøndelag. NINA Rapport 1642.

Töpper, J. & Jakobsson, S. 2021. The Index-Based Ecological Condition Assessment (IBECA) - Technical protocol, version 1.0. NINA Report 1967. Norwegian Institute for Nature Research.

Töpper, J., Velle, L.G. & Vandvik, V. 2018. Developing a method for assessment of ecological state based on indicator values after Ellenberg and Grime (revised edition). NINA Report 1529b. Norwegian Institute for Nature Research.

Tyler, T., Herbertsson, L., Olofsson, J., & Olsson, P. A. 2021. Ecological indicator and traits values for Swedish vascular plants. Ecological In-dicators, 120. doi:10.1016/j.ecolind.2020.106923

## Analyses
### Data sets

ANO data: ANO.sp contains the species data, ANO.geo contains site data
```{r, echo = F}
# Add data from cache
ANO.sp<-readRDS(paste0(here::here(),"/data/cache/ANO.sp.RDS"))
ANO.geo<-readRDS(paste0(here::here(),"/data/cache/ANO.geo.RDS"))
```

GRUK data: GRUK.species contains the species data, GRUK.variable contains site data, GRUK2021.condition contains a field-based condition assessment from the 2021 season
```{r, echo = F}
# Add data from cache
GRUK.variables<-readRDS(paste0(here::here(),"/data/cache/GRUK.variables.RDS"))
GRUK.species<-readRDS(paste0(here::here(),"/data/cache/GRUK.species.RDS"))
GRUK2021.condition<-readRDS(paste0(here::here(),"/data/cache/GRUK2021.condition.RDS"))
```

Plant indicators from Tyler et al. (2021) and Grime (1974)
```{r, echo = F}
## data from cache
ind.Tyler<-readRDS(paste0(here::here(),"/data/cache/ind.Tyler.RDS"))
ind.Grime<-readRDS(paste0(here::here(),"/data/cache/ind.Grime.RDS"))
```

Generalized species lists (reference communities): natopen_NiN_ref contains the reference species lists, natopen_NiN_ref_spInfo contains additional taxonomic information for each species
```{r, echo = F}
# data from cache
natopen_NiN_ref<-readRDS(paste0(here::here(),"/data/cache/natopen_NiN_ref.RDS"))
natopen_NiN_ref_spInfo<-readRDS(paste0(here::here(),"/data/cache/natopen_NiN_ref_spInfo.RDS"))
```

#### Data handling
- Checking for errors
- Checking species nomenclature in the different species lists to make species and indicator data possible to merge
- Merging indicator data with monitoring data and indicator data with reference data
(not shown here, but documented in the code)

```{r, include = F}

### Plant indicator data
# trimming away sub-species & co, and descriptor info
ind.Grime[,'species.orig'] <- ind.Grime[,'species']
ind.Grime[,'species'] <- word(ind.Grime[,'species'], 1,2)
names(ind.Tyler)[1] <- 'species'
ind.Tyler$species <- as.factor(ind.Tyler$species)
#summary(ind.Tyler$species)
ind.Tyler <- ind.Tyler[!is.na(ind.Tyler$species),]
ind.Tyler[,'species.orig'] <- ind.Tyler[,'species']
ind.Tyler[,'species'] <- word(ind.Tyler[,'species'], 1,2)

# dealing with 'duplicates'
ind.Grime[duplicated(ind.Grime[,'species']),"species"]
ind.Grime.dup <- ind.Grime[duplicated(ind.Grime[,'species']),c("species")]
ind.Grime[ind.Grime$species %in% ind.Grime.dup,]
# getting rid of the duplicates
ind.Grime <- ind.Grime %>% filter( !(species.orig %in% list("Carex viridula brachyrrhyncha",
                                                            "Dactylorhiza fuchsii praetermissa",
                                                            "Medicago sativa varia",
                                                            "Montia fontana chondrosperma",
                                                            "Papaver dubium lecoqii",
                                                            "Sanguisorba minor muricata")
) )
#ind.Grime[duplicated(ind.Grime[,'species']),"species"]

names(ind.Tyler)[1] <- 'species'
ind.Tyler$species <- as.factor(ind.Tyler$species)
#summary(ind.Tyler$species)
ind.Tyler <- ind.Tyler[!is.na(ind.Tyler$species),]
ind.Tyler[,'species.orig'] <- ind.Tyler[,'species']
ind.Tyler[,'species'] <- word(ind.Tyler[,'species'], 1,2)
#ind.Tyler2 <- ind.Tyler
#ind.Tyler <- ind.Tyler2
#ind.Tyler[duplicated(ind.Tyler[,'species']),"species"]
ind.Tyler.dup <- ind.Tyler[duplicated(ind.Tyler[,'species']),"species"]
#ind.Tyler[ind.Tyler$species %in% ind.Tyler.dup,c("Light","Moisture","Soil_reaction_pH","Nitrogen","species.orig","species")]
ind.Tyler <- ind.Tyler %>% filter( !(species.orig %in% list("Ammophila arenaria x Calamagrostis epigejos",
                                                            "Anemone nemorosa x ranunculoides",
                                                            "Armeria maritima ssp. elongata",
                                                            "Asplenium trichomanes ssp. quadrivalens",
                                                            "Calystegia sepium ssp. spectabilis",
                                                            "Campanula glomerata 'Superba'",
                                                            "Dactylorhiza maculata ssp. fuchsii",
                                                            "Erigeron acris ssp. droebachensis",
                                                            "Erigeron acris ssp. politus",
                                                            "Erysimum cheiranthoides L. ssp. alatum",
                                                            "Euphrasia nemorosa x stricta var. brevipila",
                                                            "Galium mollugo x verum",
                                                            "Geum rivale x urbanum",
                                                            "Hylotelephium telephium (ssp. maximum)",
                                                            "Juncus alpinoarticulatus ssp. rariflorus",
                                                            "Lamiastrum galeobdolon ssp. argentatum",
                                                            "Lathyrus latifolius ssp. heterophyllus",
                                                            "Medicago sativa ssp. falcata",
                                                            "Medicago sativa ssp. x varia",
                                                            "Monotropa hypopitys ssp. hypophegea",
                                                            "Ononis spinosa ssp. hircina",
                                                            "Ononis spinosa ssp. procurrens",
                                                            "Pilosella aurantiaca ssp. decolorans",
                                                            "Pilosella aurantiaca ssp. dimorpha",
                                                            "Pilosella cymosa ssp. gotlandica",
                                                            "Pilosella cymosa ssp. praealta",
                                                            "Pilosella officinarum ssp. peleteranum",
                                                            "Poa x jemtlandica (Almq.) K. Richt.",
                                                            "Poa x herjedalica Harry Sm.",
                                                            "Ranunculus peltatus ssp. baudotii",
                                                            "Sagittaria natans x sagittifolia",
                                                            "Salix repens ssp. rosmarinifolia",
                                                            "Stellaria nemorum L. ssp. montana",
                                                            "Trichophorum cespitosum ssp. germanicum")
) )
#ind.Tyler[duplicated(ind.Tyler[,'species']),"species"]
ind.Tyler.dup <- ind.Tyler[duplicated(ind.Tyler[,'species']),"species"]
#ind.Tyler[ind.Tyler$species %in% ind.Tyler.dup,c("Light","Moisture","Soil_reaction_pH","Nitrogen","species.orig","species")]
# getting rid of sect. for Hieracium
ind.Tyler <- ind.Tyler %>% mutate(species=gsub("sect. ","",species.orig))
ind.Tyler[,'species'] <- word(ind.Tyler[,'species'], 1,2)

#ind.Tyler[duplicated(ind.Tyler[,'species']),"species"]
ind.Tyler.dup <- ind.Tyler[duplicated(ind.Tyler[,'species']),"species"]
#ind.Tyler[ind.Tyler$species %in% ind.Tyler.dup,c("Light","Moisture","Soil_reaction_pH","Nitrogen","species.orig","species")]
# only hybrids left -> get rid of these
ind.Tyler <- ind.Tyler[!duplicated(ind.Tyler[,'species']),]
#ind.Tyler[duplicated(ind.Tyler[,'species']),"species"]

ind.Tyler$species <- as.factor(ind.Tyler$species)
#summary(ind.Tyler$species)
# no duplicates left

# merge indicator data
ind.dat <- merge(ind.Grime,ind.Tyler, by="species", all=T)
#summary(ind.dat)
#ind.dat[duplicated(ind.dat[,'species']),"species"]
ind.dat$species <- as.factor(ind.dat$species)
#summary(ind.dat$species)
head(ind.dat)

### ANO monitoring data

head(ANO.sp)
head(ANO.geo)

## fix NiN information
ANO.geo$hovedtype_rute <- substr(ANO.geo$kartleggingsenhet_1m2,1,3) # take the 3 first characters
ANO.geo$hovedtype_rute <- gsub("-", "", ANO.geo$hovedtype_rute) # remove hyphen
unique(as.factor(ANO.geo$hovedtype_rute))

## fix NiN-variables
colnames(ANO.geo)
colnames(ANO.geo)[42:47] <- c("groeftingsintensitet",
                              "bruksintensitet",
                              "beitetrykk",
                              "slatteintensitet",
                              "tungekjoretoy",
                              "slitasje")
head(ANO.geo)

# remove variable code in the data
ANO.geo$groeftingsintensitet <- gsub("7GR-GI_", "", ANO.geo$groeftingsintensitet) 
unique(ANO.geo$groeftingsintensitet)
ANO.geo$groeftingsintensitet <- gsub("X", "NA", ANO.geo$groeftingsintensitet)
unique(ANO.geo$groeftingsintensitet)
ANO.geo$groeftingsintensitet <- as.numeric(ANO.geo$groeftingsintensitet)
unique(ANO.geo$groeftingsintensitet)

ANO.geo$bruksintensitet <- gsub("7JB-BA_", "", ANO.geo$bruksintensitet) 
unique(ANO.geo$bruksintensitet)
ANO.geo$bruksintensitet <- gsub("X", "NA", ANO.geo$bruksintensitet)
unique(ANO.geo$bruksintensitet)
ANO.geo$bruksintensitet <- as.numeric(ANO.geo$bruksintensitet)
unique(ANO.geo$bruksintensitet)

ANO.geo$beitetrykk <- gsub("7JB-BT_", "", ANO.geo$beitetrykk) 
unique(ANO.geo$beitetrykk)
ANO.geo$beitetrykk <- gsub("X", "NA", ANO.geo$beitetrykk)
unique(ANO.geo$beitetrykk)
ANO.geo$beitetrykk <- as.numeric(ANO.geo$beitetrykk)
unique(ANO.geo$beitetrykk)

ANO.geo$slatteintensitet <- gsub("7JB-SI_", "", ANO.geo$slatteintensitet) 
unique(ANO.geo$slatteintensitet)
ANO.geo$slatteintensitet <- gsub("X", "NA", ANO.geo$slatteintensitet)
unique(ANO.geo$slatteintensitet)
ANO.geo$slatteintensitet <- as.numeric(ANO.geo$slatteintensitet)
unique(ANO.geo$slatteintensitet)

ANO.geo$tungekjoretoy <- gsub("7TK_", "", ANO.geo$tungekjoretoy) 
unique(ANO.geo$tungekjoretoy)
ANO.geo$tungekjoretoy <- gsub("X", "NA", ANO.geo$tungekjoretoy)
unique(ANO.geo$tungekjoretoy)
ANO.geo$tungekjoretoy <- as.numeric(ANO.geo$tungekjoretoy)
unique(ANO.geo$tungekjoretoy)

ANO.geo$slitasje <- gsub("7SE_", "", ANO.geo$slitasje) 
unique(ANO.geo$slitasje)
ANO.geo$slitasje <- gsub("X", "NA", ANO.geo$slitasje)
unique(ANO.geo$slitasje)
ANO.geo$slitasje <- as.numeric(ANO.geo$slitasje)
unique(ANO.geo$slitasje)

## check that every point is present only once
length(levels(as.factor(ANO.geo$ano_flate_id)))
length(levels(as.factor(ANO.geo$ano_punkt_id)))
summary(as.factor(ANO.geo$ano_punkt_id))
# there's many double presences, probably some wrong registrations of point numbers,
# CHECK THIS when preparing ecosystem-datasets for scaling

# fix species names
ANO.sp$Species <- ANO.sp$art_navn
unique(as.factor(ANO.sp$Species))
#ANO.sp$Species <- sub(".*?_", "", ANO.sp$Species) # lose the Norwegian name in the front
ANO.sp[,'Species'] <- word(ANO.sp[,'Species'], 1,2) # lose subspecies
ANO.sp$Species <- str_to_title(ANO.sp$Species) # make first letter capital
#ANO.sp$Species <- gsub("_", " ", ANO.sp$Species) # replace underscore with space
ANO.sp$Species <- gsub("( .*)","\\L\\1",ANO.sp$Species,perl=TRUE) # make capital letters after hyphon to lowercase
ANO.sp$Species <- gsub("( .*)","\\L\\1",ANO.sp$Species,perl=TRUE) # make capital letters after space to lowercase

## merge species data with indicators
ANO.sp.ind <- merge(x=ANO.sp[,c("Species", "art_dekning", "ParentGlobalID")], 
                    y= ind.dat[,c("species","CC", "SS", "RR","Light", "Nitrogen", "Soil_disturbance")],
                    by.x="Species", by.y="species", all.x=T)
summary(ANO.sp.ind)

## checking which species didn't find a match
unique(ANO.sp.ind[is.na(ANO.sp.ind$Light),'Species'])

# fix species name issues
ind.dat <- ind.dat %>% 
  mutate(species=str_replace(species,"Aconitum lycoctonum", "Aconitum septentrionale")) %>% 
  mutate(species=str_replace(species,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  mutate(species=str_replace(species,"Carex myosuroides", "Kobresia myosuroides")) %>%
  mutate(species=str_replace(species,"Clinopodium acinos", "Acinos arvensis")) %>%
  mutate(species=str_replace(species,"Artemisia rupestris", "Artemisia norvegica")) %>%
  mutate(species=str_replace(species,"Cherleria biflora", "Minuartia biflora"))

ANO.sp <- ANO.sp %>% 
  mutate(Species=str_replace(Species,"Arctous alpinus", "Arctous alpina")) %>%
  mutate(Species=str_replace(Species,"Betula tortuosa", "Betula pubescens")) %>%
  mutate(Species=str_replace(Species,"Blysmopsis rufa", "Blysmus rufus")) %>%
  mutate(Species=str_replace(Species,"Cardamine nymanii", "Cardamine pratensis")) %>%
  mutate(Species=str_replace(Species,"Carex adelostoma", "Carex buxbaumii")) %>%
  mutate(Species=str_replace(Species,"Carex leersii", "Carex echinata")) %>%
  mutate(Species=str_replace(Species,"Carex paupercula", "Carex magellanica")) %>%
  mutate(Species=str_replace(Species,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  mutate(Species=str_replace(Species,"Carex viridula", "Carex flava")) %>%
  mutate(Species=str_replace(Species,"Chamaepericlymenum suecicum", "Cornus suecia")) %>%
  mutate(Species=str_replace(Species,"Cicerbita alpina", "Lactuca alpina")) %>%
  mutate(Species=str_replace(Species,"Empetrum hermaphroditum", "Empetrum nigrum")) %>%
  mutate(Species=str_replace(Species,"Festuca prolifera", "Festuca rubra")) %>%
  mutate(Species=str_replace(Species,"Galium album", "Galium mollugo")) %>%
  mutate(Species=str_replace(Species,"Galium elongatum", "Galium palustre")) %>%
  mutate(Species=str_replace(Species,"Helictotrichon pratense", "Avenula pratensis")) %>%
  mutate(Species=str_replace(Species,"Helictotrichon pubescens", "Avenula pubescens")) %>%
  mutate(Species=str_replace(Species,"Hieracium alpina", "Hieracium Alpina")) %>%
  mutate(Species=str_replace(Species,"Hieracium alpinum", "Hieracium Alpina")) %>%
  mutate(Species=str_replace(Species,"Hieracium hieracium", "Hieracium Hieracium")) %>%
  mutate(Species=str_replace(Species,"Hieracium hieracioides", "Hieracium umbellatum")) %>%
  mutate(Species=str_replace(Species,"Hieracium murorum", "Hieracium Vulgata")) %>%
  mutate(Species=str_replace(Species,"Hieracium oreadea", "Hieracium Oreadea")) %>%
  mutate(Species=str_replace(Species,"Hieracium prenanthoidea", "Hieracium Prenanthoidea")) %>%
  mutate(Species=str_replace(Species,"Hieracium vulgata", "Hieracium Vulgata")) %>%
  mutate(Species=str_replace(Species,"Hieracium pilosella", "Pilosella officinarum")) %>%
  mutate(Species=str_replace(Species,"Hieracium vulgatum", "Hieracium umbellatum")) %>%
  mutate(Species=str_replace(Species,"Hierochloã« alpina", "Hierochloë alpina")) %>%
  mutate(Species=str_replace(Species,"Hierochloã« hirta", "Hierochloë hirta")) %>%
  mutate(Species=str_replace(Species,"Hierochloã« odorata", "Hierochloë odorata")) %>%
  mutate(Species=str_replace(Species,"Listera cordata", "Neottia cordata")) %>%
  mutate(Species=str_replace(Species,"Leontodon autumnalis", "Scorzoneroides autumnalis")) %>%
  mutate(Species=str_replace(Species,"Loiseleuria procumbens", "Kalmia procumbens")) %>%
  mutate(Species=str_replace(Species,"Mycelis muralis", "Lactuca muralis")) %>%
  mutate(Species=str_replace(Species,"Omalotheca supina", "Gnaphalium supinum")) %>%
  mutate(Species=str_replace(Species,"Omalotheca norvegica", "Gnaphalium norvegicum")) %>%
  mutate(Species=str_replace(Species,"Omalotheca sylvatica", "Gnaphalium sylvaticum")) %>%
  mutate(Species=str_replace(Species,"Oreopteris limbosperma", "Thelypteris limbosperma")) %>%
  mutate(Species=str_replace(Species,"Oxycoccus microcarpus", "Vaccinium microcarpum")) %>%
  mutate(Species=str_replace(Species,"Oxycoccus palustris", "Vaccinium oxycoccos")) %>%
  mutate(Species=str_replace(Species,"Phalaris minor", "Phalaris arundinacea")) %>%
  mutate(Species=str_replace(Species,"Pinus unicinata", "Pinus mugo")) %>%
  mutate(Species=str_replace(Species,"Poa alpigena", "Poa pratensis")) %>%
  mutate(Species=str_replace(Species,"Poa angustifolia", "Poa pratensis")) %>%
  mutate(Species=str_replace(Species,"Pyrola grandiflora", "Pyrola rotundifolia")) %>%
  mutate(Species=str_replace(Species,"Rumex alpestris", "Rumex acetosa")) %>%
  mutate(Species=str_replace(Species,"Syringa emodi", "Syringa vulgaris")) %>%
  mutate(Species=str_replace(Species,"Taraxacum crocea", "Taraxacum officinale")) %>%
  mutate(Species=str_replace(Species,"Taraxacum croceum", "Taraxacum officinale")) %>%
  mutate(Species=str_replace(Species,"Trientalis europaea", "Lysimachia europaea")) %>%
  mutate(Species=str_replace(Species,"Trifolium pallidum", "Trifolium pratense"))

## merge species data with indicators
ANO.sp.ind <- merge(x=ANO.sp[,c("Species", "art_dekning", "ParentGlobalID")], 
                    y= ind.dat[,c("species","CC", "SS", "RR","Light", "Nitrogen", "Soil_disturbance")],
                    by.x="Species", by.y="species", all.x=T)
summary(ANO.sp.ind)
# checking which species didn't find a match
unique(ANO.sp.ind[is.na(ANO.sp.ind$Light),'Species'])
# don't find synonyms for these in the ind lists

# trimming away the points without information on NiN, species or cover
ANO.sp.ind <- ANO.sp.ind[!is.na(ANO.sp.ind$Species),]
ANO.sp.ind <- ANO.sp.ind[!is.na(ANO.sp.ind$art_dekning),]

summary(ANO.sp.ind)
head(ANO.sp.ind)
rm(ANO.sp)

### GRUK data handling
names(GRUK.variables)
names(GRUK.species)
names(GRUK2021.condition)

colnames(GRUK.species)[5] <- "art_dekning"

# fix species names
GRUK.species <- as.data.frame(GRUK.species)

GRUK.species$Species <- GRUK.species$Navn
unique(as.factor(GRUK.species$Species))
GRUK.species$Species <- sub(".*?_", "", GRUK.species$Species) # lose the Norwegian name in the front
GRUK.species$Species <- gsub("_", " ", GRUK.species$Species) # replace underscore with space
GRUK.species$Species <- str_to_title(GRUK.species$Species) # make first letter capital
GRUK.species$Species <- gsub("( .*)","\\L\\1",GRUK.species$Species,perl=TRUE) # make capital letters after hyphon to lowercase
GRUK.species$Species <- gsub("( .*)","\\L\\1",GRUK.species$Species,perl=TRUE) # make capital letters after space to lowercase
GRUK.species[,'Species'] <- word(GRUK.species[,'Species'], 1,2) # lose subspecies

## merge species data with indicators
GRUK.species.ind <- merge(x=GRUK.species[,c("Species", "art_dekning", "ParentGlobalID")], 
                          y= ind.dat[,c("species","CC", "SS", "RR","Light", "Nitrogen", "Soil_disturbance")],
                          by.x="Species", by.y="species", all.x=T)
summary(GRUK.species.ind)

# checking which species didn't find a match
unique(GRUK.species.ind[is.na(GRUK.species.ind$Light & 
                                is.na(GRUK.species.ind$RR)),'Species'])

# fix species name issues
ind.dat <- ind.dat %>% 
  #  mutate(species=str_replace(species,"Aconitum lycoctonum", "Aconitum septentrionale")) %>% 
  #  mutate(species=str_replace(species,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  #  mutate(species=str_replace(species,"Carex myosuroides", "Kobresia myosuroides")) %>%
  #  mutate(species=str_replace(species,"Clinopodium acinos", "Acinos arvensis")) %>%
  #  mutate(species=str_replace(species,"Artemisia rupestris", "Artemisia norvegica")) %>%
  mutate(species=str_replace(species,"Rosa vosagica", "Rosa vosagiaca"))

GRUK.species <- GRUK.species %>% 
  mutate(Species=str_replace(Species,"Arabis wahlenbergii", "Arabis hirsuta")) %>%
  #  mutate(Species=str_replace(Species,"Arctous alpinus", "Arctous alpina")) %>%
  #  mutate(Species=str_replace(Species,"Betula tortuosa", "Betula pubescens")) %>%
  #  mutate(Species=str_replace(Species,"Blysmopsis rufa", "Blysmus rufus")) %>%
  #  mutate(Species=str_replace(Species,"Cardamine nymanii", "Cardamine pratensis")) %>%
  #  mutate(Species=str_replace(Species,"Carex adelostoma", "Carex buxbaumii")) %>%
  #  mutate(Species=str_replace(Species,"Carex leersii", "Carex echinata")) %>%
  mutate(Species=str_replace(Species,"Carex paupercula", "Carex magellanica")) %>%
  #  mutate(Species=str_replace(Species,"Carex simpliciuscula", "Kobresia simpliciuscula")) %>%
  mutate(Species=str_replace(Species,"Carex viridula", "Carex flava")) %>%
  #  mutate(Species=str_replace(Species,"Chamaepericlymenum suecicum", "Cornus suecia")) %>%
  #  mutate(Species=str_replace(Species,"Cicerbita alpina", "Lactuca alpina")) %>%
  mutate(Species=str_replace(Species,"Cotoneaster scandinavicus", "Cotoneaster integerrimus")) %>%
  mutate(Species=str_replace(Species,"Cotoneaster symondsii", "Cotoneaster integrifolius")) %>%
  mutate(Species=str_replace(Species,"Cyanus montanus", "Centaurea montana")) %>%
  #  mutate(Species=str_replace(Species,"Empetrum hermaphroditum", "Empetrum nigrum")) %>%
  mutate(Species=str_replace(Species,"Erysimum virgatum", "Erysimum strictum")) %>%
  #  mutate(Species=str_replace(Species,"Festuca prolifera", "Festuca rubra")) %>%
  mutate(Species=str_replace(Species,"Festuca trachyphylla", "Festuca brevipila")) %>%
  mutate(Species=str_replace(Species,"Galium album", "Galium mollugo")) %>%
  #  mutate(Species=str_replace(Species,"Galium elongatum", "Galium palustre")) %>%
  mutate(Species=str_replace(Species,"Helictotrichon pratense", "Avenula pratensis")) %>%
  mutate(Species=str_replace(Species,"Helictotrichon pubescens", "Avenula pubescens")) %>%
  #  mutate(Species=str_replace(Species,"Hieracium alpina", "Hieracium Alpina")) %>%
  #  mutate(Species=str_replace(Species,"Hieracium alpinum", "Hieracium Alpina")) %>%
  #  mutate(Species=str_replace(Species,"Hieracium hieracium", "Hieracium Hieracium")) %>%
  #  mutate(Species=str_replace(Species,"Hieracium hieracioides", "Hieracium umbellatum")) %>%
  mutate(Species=str_replace(Species,"Hieracium murorum", "Hieracium Vulgata")) %>%
  #  mutate(Species=str_replace(Species,"Hieracium oreadea", "Hieracium Oreadea")) %>%
  #  mutate(Species=str_replace(Species,"Hieracium prenanthoidea", "Hieracium Prenanthoidea")) %>%
  #  mutate(Species=str_replace(Species,"Hieracium vulgata", "Hieracium Vulgata")) %>%
  #  mutate(Species=str_replace(Species,"Hieracium pilosella", "Pilosella officinarum")) %>%
  #  mutate(Species=str_replace(Species,"Hieracium vulgatum", "Hieracium umbellatum")) %>%
  #  mutate(Species=str_replace(Species,"Hierochloã« alpina", "Hierochloë alpina")) %>%
  #  mutate(Species=str_replace(Species,"Hierochloã« hirta", "Hierochloë hirta")) %>%
  #  mutate(Species=str_replace(Species,"Hierochloã« odorata", "Hierochloë odorata")) %>%
  mutate(Species=str_replace(Species,"Hylotelephium maximum", "Sedum telephium")) %>%
  #  mutate(Species=str_replace(Species,"Listera cordata", "Neottia cordata")) %>%
  #  mutate(Species=str_replace(Species,"Leontodon autumnalis", "Scorzoneroides autumnalis")) %>%
  mutate(Species=str_replace(Species,"Lepidotheca suaveolens", "Matricaria discoidea")) %>%
  #  mutate(Species=str_replace(Species,"Loiseleuria procumbens", "Kalmia procumbens")) %>%
  mutate(Species=str_replace(Species,"Malus ×domestica", "Malus domestica")) %>%
  #  mutate(Species=str_replace(Species,"Mycelis muralis", "Lactuca muralis")) %>%
  #  mutate(Species=str_replace(Species,"Omalotheca supina", "Gnaphalium supinum")) %>%
  #  mutate(Species=str_replace(Species,"Omalotheca norvegica", "Gnaphalium norvegicum")) %>%
  #  mutate(Species=str_replace(Species,"Omalotheca sylvatica", "Gnaphalium sylvaticum")) %>%
  #  mutate(Species=str_replace(Species,"Oreopteris limbosperma", "Thelypteris limbosperma")) %>%
  #  mutate(Species=str_replace(Species,"Oxycoccus microcarpus", "Vaccinium microcarpum")) %>%
  #  mutate(Species=str_replace(Species,"Oxycoccus palustris", "Vaccinium oxycoccos")) %>%
  #  mutate(Species=str_replace(Species,"Phalaris minor", "Phalaris arundinacea")) %>%
  #  mutate(Species=str_replace(Species,"Pinus unicinata", "Pinus mugo")) %>%
  #  mutate(Species=str_replace(Species,"Poa alpigena", "Poa pratensis")) %>%
  mutate(Species=str_replace(Species,"Poa angustifolia", "Poa pratensis")) %>%
  mutate(Species=str_replace(Species,"Poa humilis", "Poa pratensis")) %>%
  #  mutate(Species=str_replace(Species,"Pyrola grandiflora", "Pyrola rotundifolia")) %>%
  mutate(Species=str_replace(Species,"Rosa dumalis", "Rosa vosagiaca")) %>%
  #  mutate(Species=str_replace(Species,"Rumex alpestris", "Rumex acetosa")) %>%
  #  mutate(Species=str_replace(Species,"Syringa emodi", "Syringa vulgaris")) %>%
  #  mutate(Species=str_replace(Species,"Taraxacum crocea", "Taraxacum officinale")) %>%
  #  mutate(Species=str_replace(Species,"Taraxacum croceum", "Taraxacum officinale")) %>%
  #  mutate(Species=str_replace(Species,"Trientalis europaea", "Lysimachia europaea")) %>%
  mutate(Species=str_replace(Species,"Trifolium pallidum", "Trifolium pratense"))

## merge species data with indicators
GRUK.species.ind <- merge(x=GRUK.species[,c("Species", "art_dekning", "ParentGlobalID")], 
                          y= ind.dat[,c("species","CC", "SS", "RR","Light", "Nitrogen", "Soil_disturbance")],
                          by.x="Species", by.y="species", all.x=T)
summary(GRUK.species.ind)
# checking which species didn't find a match
unique(GRUK.species.ind[is.na(GRUK.species.ind$Light & 
                                is.na(GRUK.species.ind$RR)),'Species'])

### make GRUK.variables into spatial object
names(GRUK.variables)
GRUK.variables <- st_as_sf(GRUK.variables, coords = c("x","y"),remove=F)
# add CRS
GRUK.variables <- st_set_crs(GRUK.variables,4326)
# transform CRS to match ANO
GRUK.variables <- GRUK.variables %>%
  st_as_sf() %>%
  st_transform(crs = st_crs(ANO.geo))

### merge GRUK.variables and GRUK2021.condition
#GRUK.variables <- as.data.frame(GRUK.variables)
GRUK2021.condition <- as.data.frame(GRUK2021.condition)

head(GRUK.variables)
head(GRUK2021.condition)

## fixing site-ID in the condition-data
GRUK2021.condition$Flate_ID <- GRUK2021.condition$områdenavn
unique(as.factor(GRUK2021.condition$Flate_ID))
GRUK2021.condition$Flate_ID <- gsub("_", " ", GRUK2021.condition$Flate_ID) # replace underscore with space
unique(as.factor(GRUK2021.condition$Flate_ID))
GRUK2021.condition$Flate_ID <- sub(" .*", "", GRUK2021.condition$Flate_ID) # remove everything after space
unique(as.factor(GRUK2021.condition$Flate_ID))

#GRUK2021.condition$Flate_ID <- gsub(" ", "-", GRUK2021.condition$Flate_ID) # replace space with hyphon
#GRUK2021.condition$Flate_ID <- gsub("[^0-9\\-]", "", GRUK2021.condition$Flate_ID) # remove all non-numerics except '-'
#GRUK2021.condition$Flate_ID <- gsub("--", "-", GRUK2021.condition$Flate_ID) # remove double hyphons
#GRUK2021.condition$Flate_ID <- substr(GRUK2021.condition$Flate_ID, 1, nchar(GRUK2021.condition$Flate_ID)-1) # remove the last character ('-' for all)

colnames(GRUK.variables)
colnames(GRUK2021.condition)

GRUK.variables <- merge(x=GRUK.variables, 
                        y=GRUK2021.condition[,c("tilstandsvurdering","tilstandsgrunn","artsmangfoldvurdering","lokalitetskvalitet","Flate_ID")], 
                        by.x="Flate_ID", by.y="Flate_ID", all.x=T)

## adding information on ecosystem and condition variables to species data
GRUK.species.ind <- merge(x=GRUK.species.ind, 
                          y=GRUK.variables[,c("GlobalID","year","Flate_ID","Punkt_ID",
                                              "Total dekning % av karplanter registert","Dekning % av karplanter i feltsjikt","Dekning % av moser",
                                              "Dekning % av lav","Dekning % av strø","Dekning % av bar jord/grus/stein/berg",
                                              "Kartleggingsenhet",
                                              "Spor etter ferdsel med tunge kjøretøy (%)","Spor etter slitasje og slitasjebetinget erosjon (%)","Dekning % av nakent berg","Menneskeskapte objekter i sirkelen?",
                                              "Total dekning % av vedplanter i feltsjikt","Dekning % av busker i busksjikt","Dekning % av tresjikt","Dekning % av problemarter","Total dekning % av fremmede arter",
                                              "x","y")], 
                          by.x="ParentGlobalID", by.y="GlobalID", all.x=T)
# trimming away the points without information on NiN, species or cover
GRUK.species.ind <- GRUK.species.ind[!is.na(GRUK.species.ind$Species),]
GRUK.species.ind <- GRUK.species.ind[!is.na(GRUK.species.ind$art_dekning),]

#rm(GRUK.species)
#rm(GRUK.variables)

summary(GRUK.species.ind)
head(GRUK.species.ind)

### reference data - data handling

head(natopen_NiN_ref)
head(natopen_NiN_ref_spInfo)

colnames(natopen_NiN_ref)[1] <- "sp"
head(natopen_NiN_ref)

natopen_NiN_ref <- merge(natopen_NiN_ref,natopen_NiN_ref_spInfo[,c(1,4)], by.x="sp", by.y="ScientificName", all.x=T)
unique(natopen_NiN_ref[is.na(natopen_NiN_ref$Phylum),'sp']) # Pucinella does not exist in ind.dat, so we don't care
# we're only interested in vascular plants and ferns, which we have indicators on
unique(natopen_NiN_ref$Phylum)
natopen_NiN_ref <- natopen_NiN_ref %>%
  filter(Phylum %in% c("Magnoliophyta","Pteridophyta"))
unique(natopen_NiN_ref$Phylum)

natopen_NiN_ref$sp
# only genus and species name
natopen_NiN_ref$sp.orig <- natopen_NiN_ref$sp
natopen_NiN_ref$sp <- word(natopen_NiN_ref$sp, 1,2)
natopen_NiN_ref <- natopen_NiN_ref[!is.na(natopen_NiN_ref$sp),]
# merging with indicator values
NiN.natopen <- merge(natopen_NiN_ref,ind.dat[,c(1,3:5,20,23,27)], by.x="sp", by.y="species", all.x=T)
head(NiN.natopen)
summary(NiN.natopen)

NiN.natopen
unique(NiN.natopen$sp)
#NiN.sp$spgr <- as.factor(as.vector(Eco_State$Concept_Data$Species$Species_List$art.code))

# checking which species didn't find a match
unique(NiN.natopen[is.na(NiN.natopen$Light) & is.na(NiN.natopen$Nitrogen) & is.na(NiN.natopen$RR),'sp'])

# fix species name issues
natopen_NiN_ref <- natopen_NiN_ref %>% 
  mutate(sp=str_replace(sp,"Cirsium acaulon", "Cirsium acaule")) %>%
  mutate(sp=str_replace(sp,"Hylotelephium maximum", "Hylotelephium telephium"))

ind.dat[2556,'species'] <- "Saxifraga osloënsis"
ind.dat[17,'species'] <- "Hierochloë odorata"
ind.dat[9,'species'] <- "Hippophaë rhamnoides"

# merging with indicator values
NiN.natopen <- merge(natopen_NiN_ref,ind.dat[,c(1,3:5,20,23,27)], by.x="sp", by.y="species", all.x=T)
# checking which species didn't find a match
unique(NiN.natopen[is.na(NiN.natopen$Light) & is.na(NiN.natopen$Nitrogen) & is.na(NiN.natopen$RR),'sp'])
# ok now

# translating the abundance classes into %-cover
coverscale <- data.frame(orig=0:6,
                         cov=c(0, 1/32 ,1/8, 3/8, 0.6, 4/5, 1)
)

NiN.natopen.cov <- NiN.natopen
colnames(NiN.natopen.cov)
for (i in 2:71) {
  NiN.natopen.cov[,i] <- coverscale[,2][ match(NiN.natopen[,i], 0:6 ) ]
}

summary(NiN.natopen)
summary(NiN.natopen.cov)

```
leaving us with the monitoring data including plant indicators (ANO.sp.ind, GRUK.species.ind) and the reference data including plant indicators (NiN.natopen.cov)

```{r}
head(ANO.sp.ind)
head(GRUK.species.ind)
head(NiN.natopen.cov)
```
For each ecosystem type with a NiN species list, we can calculate a community weighted mean (CWM) for the relevant functional plant indicators.
For semi-natural ecosystems, we are testing "Light", "Moisture", "Soil_reaction_pH", "Nitrogen", "Phosphorus", "Grazing_mowing", and "Soil_disturbance". In order to get distributions of CWMs rather than one single value (for comparison with the empirical testing data) the NiN lists can be bootstrapped.

##### bootstrap function for frequency abundance
- function to calculate community weighted means of selected indicator values (ind)
- for species lists (sp) with given abundances in percent (or on a scale from 0 to 1) in one or more 'sites' (abun)
- with a given number of iterations (iter),
- with species given a certain minimum abundance occurring in all bootstraps (obl), and
- with a given re-sampling ratio of the original species list (rat)
- in every bootstrap iteration the abundance of the sampled species can be randomly changed by a limited amount if wished by introducing a re-sampling of abundance values from adjacent abundance steps with a certain probability (var.abun)

```{r, echo = F}
indBoot.freq <- function(sp,abun,ind,iter,obl,rat=2/3,var.abun=F) {

  ind.b <- matrix(nrow=iter,ncol=length(colnames(abun)))
  colnames(ind.b) <- colnames(abun)
  ind.b <- as.data.frame(ind.b)

  ind <- as.data.frame(ind)
  ind.list <- as.list(1:length(colnames(ind)))
  names(ind.list) <- colnames(ind)

  for (k in 1:length(colnames(ind)) ) {
    ind.list[[k]] <- ind.b }

  for (j in 1:length(colnames(abun)) ) {

    dat <- cbind(sp,abun[,j],ind)
    dat <- dat[dat[,2]>0,]            # only species that are present in the ecosystem
    dat <- dat[!is.na(dat[,3]),]      # only species that have indicator values

    for (i in 1:iter) {

      speciesSample <- sample(dat$sp[dat[,2] < obl], size=round( (length(dat$sp)-length(dat$sp[dat[,2]>=obl])) *rat,0), replace=F)  
      dat.b <- rbind(dat[dat[,2] >= obl,],
                     dat[match(speciesSample,dat$sp),]
      )

      if (var.abun==T) {
        for (m in 1:nrow(coverscale[-1,]) ) {
          xxx <- dat.b[dat.b[,2]==coverscale[-1,][m,2],2]
          if ( m==1 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.5, 0.5, 0.0, 0.0, 0.0, 0.0, 0.0) ,size=length(xxx),replace=T) }
          if ( m==2 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.2, 0.3, 0.5, 0.0, 0.0, 0.0, 0.0) ,size=length(xxx),replace=T) }
          if ( m==3 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.0, 0.2, 0.3, 0.5, 0.0, 0.0, 0.0) ,size=length(xxx),replace=T) }
          if ( m==4 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.0, 0.0, 0.2, 0.3, 0.5, 0.0, 0.0) ,size=length(xxx),replace=T) }
          if ( m==5 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.0, 0.0, 0.0, 0.2, 0.3, 0.5, 0.0) ,size=length(xxx),replace=T) }
          if ( m==6 ) { dat.b[dat.b[,2]==coverscale[-1,][m,2],2] <- sample( c(0.01,coverscale[2:7,2]), prob = c(0.0, 0.0, 0.0, 0.0, 0.2, 0.3, 0.5) ,size=length(xxx),replace=T) }
        }
        dat.b[!is.na(dat.b[,2]) & dat.b[,2]<=(0),2] <- 0.01
        dat.b[!is.na(dat.b[,2]) & dat.b[,2]>1,2] <- 1
      }

      for (k in 1:length(colnames(ind))) {

        if ( nrow(dat.b)>2 ) {

          ind.b <- sum(dat.b[!is.na(dat.b[,2+k]),2] * dat.b[!is.na(dat.b[,2+k]),2+k] , na.rm=T) / sum(dat.b[!is.na(dat.b[,2+k]),2],na.rm=T)
          ind.list[[k]][i,j] <- ind.b

        } else {ind.list[[k]][i,j] <- NA}

      }

#      print(paste(i,"",j)) 
    }

  }
  return(ind.list)
}
```

Running the bootstraps

```{r, echo = T, eval = F}
colnames(NiN.natopen)
# 1st column is the species
# 6th-71st column is the abundances of sp in different ecosystem types
# 74th-79th column is the indicator values of the respective species
# we choose 1000 iterations
# species with abundance 1 (i.e. a max of 100%, must be included in each sample)
# each sample re-samples 1/3 of the number of species
# the abundance of the re-sampled species may vary (see bootstrap function for details)
natopen.ref.cov <- indBoot.freq(sp=NiN.natopen.cov[,1],abun=NiN.natopen.cov[,6:71],ind=NiN.natopen.cov[,74:79],iter=1000,obl=1,rat=1/3,var.abun=T)

### fixing NaN's
for (i in 1:length(natopen.ref.cov) ) {
  for (j in 1:ncol(natopen.ref.cov[[i]]) ) {
    v <- natopen.ref.cov[[i]][,j]
    v[is.nan(v)] <- NA
    natopen.ref.cov[[i]][,j] <- v
  }
}
```

```{r, include = F}
# Data from cache
natopen.ref.cov<-readRDS(paste0(here::here(), "/data/cache/natopen.ref.cov.RDS"))
```

```{r}
head(natopen.ref.cov[[1]])
```

This results in an R-list with a slot for every selected indicator, and in every slot there's a data frame with as many columns as there are NiN species lists and as many rows as there were iterations in the bootstrap.
Next, we need to derive scaling values from these bootstrap-lists (the columns) for every mapping unit in NiN. Here, we define things in the following way:

- Median = reference values
- 0.025 and 0.975 quantiles = lower and upper limit values
- min and max of the respective indicator's scale = min/max values

```{r, echo = F}

# NiN-types where each type is represented by one species list (including when one species list represents two NiN-types)
names(natopen.ref.cov[["Light"]])
x <- 1:66

# checking the actual NiN-types in the natopen lists
natopen.NiNtypes <- colnames(natopen.ref.cov[["Light"]])
natopen.NiNtypes[-x] <- substr(natopen.NiNtypes[-x], 1, nchar(natopen.NiNtypes[-x])-1)
natopen.NiNtypes

# 7 indicator-value indicators: Grimes C, S and R, Tyler's Light, Nitrogen, and Soil_disturbance
indEll.n=6
# creating a table to hold:
# Tyler: the 0.5 quantile (median), 0.05 quantile and  0.95 quantile for each NiN-type
# for every nature type (nrows)
tab <- matrix(ncol=3*indEll.n, nrow=length(unique(natopen.NiNtypes)) ) # 66 basic ecosystem types
# coercing the values into the table

for (i in 1:length(x) ) {
  tab[i,1:3] <- quantile(as.matrix(natopen.ref.cov[["CC"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[i,4:6] <- quantile(as.matrix(natopen.ref.cov[["SS"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[i,7:9] <- quantile(as.matrix(natopen.ref.cov[["RR"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[i,10:12] <- quantile(as.matrix(natopen.ref.cov[["Light"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[i,13:15] <- quantile(as.matrix(natopen.ref.cov[["Nitrogen"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[i,16:18] <- quantile(as.matrix(natopen.ref.cov[["Soil_disturbance"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)

}

tab <- as.data.frame(tab)
tab$NiN <- NA
tab$NiN[1:length(x)] <- names(natopen.ref.cov[[1]])[x]
tab

# making it a proper data frame
dim(tab)
round(tab[,1:18],digits=2)

colnames(tab) <- c("CC_q2.5","CC_q50","CC_q97.5",
                   "SS_q2.5","SS_q50","SS_q97.5",
                   "RR_q2.5","RR_q50","RR_q97.5",
                   "Light_q2.5","Light_q50","Light_q97.5",
                   "Nitrogen_q2.5","Nitrogen_q50","Nitrogen_q97.5",
                   "Soil_disturbance_q2.5","Soil_disturbance_q50","Soil_disturbance_q97.5",
                   "NiN")
summary(tab)
tab$NiN
#tab$NiN <- gsub("C", "C-", tab$NiN) # add extra hyphen after C for NiN-types
tab

# restructuring into separate indicators for lower (q2.5) and higher (q97.5) than reference value (=median, q50)
y.Light <- numeric(length=nrow(tab)*2)
y.Light[((1:dim(tab)[1])*2)-1] <- tab$Light_q2.5 
y.Light[((1:dim(tab)[1])*2)] <- tab$Light_q97.5

y.CC <- numeric(length=nrow(tab)*2)
y.CC[((1:dim(tab)[1])*2)-1] <- tab$CC_q2.5 
y.CC[((1:dim(tab)[1])*2)] <- tab$CC_q97.5

y.SS <- numeric(length=nrow(tab)*2)
y.SS[((1:dim(tab)[1])*2)-1] <- tab$SS_q2.5 
y.SS[((1:dim(tab)[1])*2)] <- tab$SS_q97.5

y.Nitrogen <- numeric(length=nrow(tab)*2)
y.Nitrogen[((1:dim(tab)[1])*2)-1] <- tab$Nitrogen_q2.5 
y.Nitrogen[((1:dim(tab)[1])*2)] <- tab$Nitrogen_q97.5

y.RR <- numeric(length=nrow(tab)*2)
y.RR[((1:dim(tab)[1])*2)-1] <- tab$RR_q2.5 
y.RR[((1:dim(tab)[1])*2)] <- tab$RR_q97.5

y.Soil_disturbance <- numeric(length=nrow(tab)*2)
y.Soil_disturbance[((1:dim(tab)[1])*2)-1] <- tab$Soil_disturbance_q2.5 
y.Soil_disturbance[((1:dim(tab)[1])*2)] <- tab$Soil_disturbance_q97.5

# creating final objects holding the reference and limit values for all indicators

# ref object for indicators
natopen.ref.cov.val <- data.frame(N1=rep('natopen',(nrow(tab)*2*indEll.n)),
                              hoved=c(rep('NA',(nrow(tab)*2*indEll.n))),
                              grunn=c(rep(rep(tab$NiN,each=2),indEll.n)),
                              county=rep('all',(nrow(tab)*2*indEll.n)),
                              region=rep('all',(nrow(tab)*2*indEll.n)),
                              Ind=c(rep(c('CC1','CC2'),nrow(tab)),
                                    rep(c('SS1','SS2'),nrow(tab)),
                                    rep(c('RR1','RR2'),nrow(tab)),
                                    rep(c('Light1','Light2'),nrow(tab)),
                                    rep(c('Nitrogen1','Nitrogen2'),nrow(tab)),
                                    rep(c('Soil_disturbance1','Soil_disturbance2'),nrow(tab))
                              ),
                              Rv=c(rep(tab$CC_q50,each=2),
                                   rep(tab$SS_q50,each=2),
                                   rep(tab$RR_q50,each=2),
                                   rep(tab$Light_q50,each=2),
                                   rep(tab$Nitrogen_q50,each=2),
                                   rep(tab$Soil_disturbance_q50,each=2)
                              ),
                              Gv=c(y.CC,y.SS,y.RR,y.Light,y.Nitrogen,y.Soil_disturbance),
                              maxmin=c(rep(c(0,1),nrow(tab)), # CC
                                       rep(c(0,1),nrow(tab)), # SS
                                       rep(c(0,1),nrow(tab)), # RR
                                       rep(c(1,7),nrow(tab)),  # 7 levels of Light
                                       rep(c(1,9),nrow(tab)),  # 9 levels of Nitrogen
                                       rep(c(1,9),nrow(tab))  # 9 levels of Soil_disturbance
                              )
)

natopen.ref.cov.val
natopen.ref.cov.val$grunn <- as.factor(natopen.ref.cov.val$grunn)
natopen.ref.cov.val$Ind <- as.factor(natopen.ref.cov.val$Ind)
summary(natopen.ref.cov.val)

```

```{r}
head(natopen.ref.cov.val)
```

Once test data (ANO, GRUK) and the scaling values from the reference data are in place, we can calculate community-weighted means (CWM) of the selected indicators for the ANO and GRUK community data and scale them against the scaling values from the reference distribution. Note that we scale each ANO/GRUK plot's CWM against either the lower threshold value and the min value OR the upper threshold value and the max value based on whether the CWM is smaller or higher than the reference value. Since the scaled values for both sides range between 0 and 1, we generate separate lower and upper condition indicators for each  functional plant indicator. An ANO/GRUK plot can only have a scaled value in either the lower or the upper indicator (the other one will be 'NA'), except for the unlikely event that the CWM exactly matches the reference value, in which case both lower and upper indicator will receive a scaled indicator value of 1.

Here is the scaling function
```{r}

#### scaled values ####
r.s <- 1    # reference value
l.s <- 0.6  # limit value
a.s <- 0    # abscence of indicator, or indicator at maximum

#### function for calculating scaled values for measured value ####

## scaling function including truncation
scal <- function() {
  # place to hold the result
   x <- numeric()
  if (maxmin < ref) {
    # values >= the reference value equal 1
    if (val >= ref) {x <- 1}
    # values < the reference value and >= the limit value can be deducted from the linear relationship between these two
    if (val < ref & val >= lim) {x <- (l.s + (val-lim) * ( (r.s-l.s) / (ref-lim) ) )}
    # values < the limit value and > maxmin can be deducted from the linear relationship between these two
    if (val < lim & val > maxmin) {x <- (a.s + (val-maxmin) * ( (l.s-a.s) / (lim-maxmin) ) )}
    # value equals or lower than maxmin
    if (val <= maxmin) {x <-0}
  } else {
    # values <= the reference value equal 1
    if (val <= ref) {x <- 1}
    # values > the reference value and <= the limit value can be deducted from the linear relationship between these two
    if (val > ref & val <= lim) {x <- ( r.s - ( (r.s - l.s) * (val - ref) / (lim - ref) ) )}
    # values > the limit value and < maxmin can be deducted from the linear relationship between these two
    if (val > lim) {x <- ( l.s - (l.s * (val - lim) / (maxmin - lim) ) )}
    # value equals or larger than maxmin
    if (val >= maxmin) {x <-0}
  }
  return(x)

}

```

We then can prepare a list of data frames to hold the results and perform the scaling according to the principles described in NINA report 1967 (Töpper and Jakobsson 2021)
This is done separately for ANO and ASO. First for ANO:
```{r, include=FALSE}

unique(ANO.geo$hovedtype_rute) # NiN types in data
unique(substr(natopen.ref.cov.val$grunn,1,3)) # NiN types in reference
#### creating dataframe to hold the results for natopens ####
# all ANO points
nrow(ANO.geo)
# all natopen ANO points
nrow(ANO.geo[ANO.geo$hovedtype_rute %in% list("T2","T8","T11","T12","T13","T15","T16","T18","T21","T24","T29") & ANO.geo$hovedoekosystem_punkt!='fjell',])
nrow(ANO.geo[ANO.geo$hovedtype_rute %in% list("T2") & ANO.geo$hovedoekosystem_punkt!='fjell',])
nrow(ANO.geo[ANO.geo$hovedtype_rute %in% list("T13") & ANO.geo$hovedoekosystem_punkt!='fjell',])
nrow(ANO.geo[ANO.geo$hovedtype_rute %in% list("T16") & ANO.geo$hovedoekosystem_punkt!='fjell',])
# all natopen ANO points with a NiN-type represented in the reference
nrow(ANO.geo[ANO.geo$hovedtype_rute %in% unique(gsub("-", "",substr(natopen.ref.cov.val$grunn,1,3))) & ANO.geo$hovedoekosystem_punkt!='fjell',])
# leaving us with a total of 272 ANO-points with naturally open ecosystems that are not in the mountains
ANO.natopen <- ANO.geo[ANO.geo$hovedtype_rute %in% unique(gsub("-", "",substr(natopen.ref.cov.val$grunn,1,3))) & ANO.geo$hovedoekosystem_punkt!='fjell',]
ANO.natopen <- ANO.natopen[!is.na(ANO.natopen$kartleggingsenhet_1m2),]
head(ANO.natopen)
# update row-numbers
row.names(ANO.natopen) <- 1:nrow(ANO.natopen)
head(ANO.natopen)
dim(ANO.natopen)
colnames(ANO.natopen)

length(levels(as.factor(ANO.natopen$ano_flate_id)))
length(levels(as.factor(ANO.natopen$ano_punkt_id)))
summary(as.factor(ANO.natopen$ano_punkt_id))
# one point is double
ANO.natopen[ANO.natopen$ano_punkt_id=="ANO0084_51",] # double registration of # 51
ANO.geo[ANO.geo$ano_punkt_id=="ANO0084_51",] # double registration of # 51

ANO.geo[ANO.geo$ano_flate_id=="ANO0084","ano_punkt_id"] # also # 62 is registered double
ANO.geo[ANO.geo$ano_punkt_id=="ANO0084_62",] # but this is not a natopen-type
# seems to be a legit registration of a naturally open type on both ANO-points # 51.
# Both points # 22, 35, 44, and 53 are missing, so we simply assign the 2nd # 51 to # 53 for this analysis
ANO.natopen[ANO.natopen$ano_punkt_id=="ANO0084_51" & !is.na(ANO.natopen$ano_punkt_id),][2,"ano_punkt_id"] <- "ANO0084_53"
ANO.natopen[ANO.natopen$ano_flate_id=="ANO0084","ano_punkt_id"]
summary(as.factor(ANO.natopen$ano_punkt_id))

#ANO.natopen <- ANO.natopen[-206,]
#row.names(ANO.natopen) <- 1:nrow(ANO.natopen) # update row-numbers

unique(ANO.natopen$hovedoekosystem_punkt)
unique(ANO.natopen$hovedtype_rute)
unique(ANO.natopen$kartleggingsenhet_1m2)
ANO.natopen$hovedtype_rute <- as.factor(ANO.natopen$hovedtype_rute)
ANO.natopen$kartleggingsenhet_1m2 <- as.factor(ANO.natopen$kartleggingsenhet_1m2)
summary(ANO.natopen$hovedtype_rute)
summary(ANO.natopen$kartleggingsenhet_1m2)
# we have no reference for overall T2 registrations, these 24 observations won't be evaluated
ANO.natopen[ANO.natopen$kartleggingsenhet_1m2=="T2",] # were largely inaccessible and thus mapped from arial pictures

results.natopen.ANO <- list()
ind <- unique(natopen.ref.cov.val$Ind)
# choose columns for site description
colnames(ANO.natopen)
results.natopen.ANO[['original']] <- ANO.natopen
# drop geometry
st_geometry(results.natopen.ANO[['original']]) <- NULL
results.natopen.ANO[['original']]

# add columns for indicators
nvar.site <- ncol(results.natopen.ANO[['original']])
for (i in 1:length(ind) ) {results.natopen.ANO[['original']][,i+nvar.site] <- NA}
colnames(results.natopen.ANO[['original']])[(nvar.site+1):(length(ind)+nvar.site)] <- paste(ind)
for (i in (nvar.site+1):(length(ind)+nvar.site) ) {results.natopen.ANO[['original']][,i] <- as.numeric(results.natopen.ANO[['original']][,i])}
summary(results.natopen.ANO[['original']])
#results.natopen.ANO[['original']]$Region <- as.factor(results.natopen.ANO[['original']]$Region)
results.natopen.ANO[['original']]$GlobalID <- as.factor(results.natopen.ANO[['original']]$GlobalID)
results.natopen.ANO[['original']]$ano_flate_id <- as.factor(results.natopen.ANO[['original']]$ano_flate_id)
results.natopen.ANO[['original']]$ano_punkt_id <- as.factor(results.natopen.ANO[['original']]$ano_punkt_id)
results.natopen.ANO[['original']]$hovedoekosystem_punkt <- as.factor(results.natopen.ANO[['original']]$hovedoekosystem_punkt)
#results.natopen.ANO[['original']]$Hovedoekosystem_rute  <- as.factor(results.natopen.ANO[['original']]$Hovedoekosystem_rute )
results.natopen.ANO[['original']]$kartleggingsenhet_1m2 <- as.factor(results.natopen.ANO[['original']]$kartleggingsenhet_1m2)
results.natopen.ANO[['original']]$hovedtype_rute    <- as.factor(results.natopen.ANO[['original']]$hovedtype_rute)

# roll out
results.natopen.ANO[['scaled']] <- results.natopen.ANO[['non-truncated']] <- results.natopen.ANO[['original']]

```

```{r, echo = T, eval = F}

#### calculating scaled and non-truncated values for the indicators based on the dataset ####
for (i in 1:nrow(ANO.natopen) ) {  #
  tryCatch({
    print(i)
    print(paste(ANO.natopen$ano_flate_id[i]))
    print(paste(ANO.natopen$ano_punkt_id[i]))
#    ANO.natopen$Hovedoekosystem_sirkel[i]
#    ANO.natopen$Hovedoekosystem_rute[i]

    # if the ANO.hovedtype exists in the reference
    if (ANO.natopen$hovedtype_rute[i] %in% gsub("-","",unique(substr(natopen.ref.cov.val$grunn,1,3))) ) {

      # if there is any species present in current ANO point  
      if ( length(ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.natopen$GlobalID[i]),'Species']) > 0 ) {

        # Grime's C

        dat <- ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.natopen$GlobalID[i]),c('art_dekning','CC')]
        results.natopen.ANO[['original']][i,'richness'] <- nrow(dat)
        dat <- dat[!is.na(dat$CC),]

        if ( nrow(dat)>0 ) {

          val <- sum(dat[,'art_dekning'] * dat[,'CC'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
          # lower part of distribution
          if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
          } else {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
          }
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
          # coercing x into results.natopen.ANO dataframe
          results.natopen.ANO[['scaled']][i,'CC1'] <- scal() 
          results.natopen.ANO[['non-truncated']][i,'CC1'] <- scal.2() 
          results.natopen.ANO[['original']][i,'CC1'] <- val

          # upper part of distribution
          if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
          } else {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
          }          
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
          # coercing x into results.natopen.ANO dataframe
          results.natopen.ANO[['scaled']][i,'CC2'] <- scal() 
          results.natopen.ANO[['non-truncated']][i,'CC2'] <- scal.2() 
          results.natopen.ANO[['original']][i,'CC2'] <- val
        }

        # Grime's S
        dat <- ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.natopen$GlobalID[i]),c('art_dekning','SS')]
        results.natopen.ANO[['original']][i,'richness'] <- nrow(dat)
        dat <- dat[!is.na(dat$SS),]

        if ( nrow(dat)>0 ) {

          val <- sum(dat[,'art_dekning'] * dat[,'SS'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
          # lower part of distribution
          if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
          } else {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
          }          
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
          # coercing x into results.natopen.ANO dataframe
          results.natopen.ANO[['scaled']][i,'SS1'] <- scal() 
          results.natopen.ANO[['non-truncated']][i,'SS1'] <- scal.2() 
          results.natopen.ANO[['original']][i,'SS1'] <- val

          # upper part of distribution
          if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
          } else {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
          }          
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
          # coercing x into results.natopen.ANO dataframe
          results.natopen.ANO[['scaled']][i,'SS2'] <- scal() 
          results.natopen.ANO[['non-truncated']][i,'SS2'] <- scal.2() 
          results.natopen.ANO[['original']][i,'SS2'] <- val
        }

        # Grime's R
        dat <- ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.natopen$GlobalID[i]),c('art_dekning','RR')]
        results.natopen.ANO[['original']][i,'richness'] <- nrow(dat)
        dat <- dat[!is.na(dat$RR),]

        if ( nrow(dat)>0 ) {

          val <- sum(dat[,'art_dekning'] * dat[,'RR'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
          # lower part of distribution
          if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
          } else {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
          }          
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
          # coercing x into results.natopen.ANO dataframe
          results.natopen.ANO[['scaled']][i,'RR1'] <- scal() 
          results.natopen.ANO[['non-truncated']][i,'RR1'] <- scal.2() 
          results.natopen.ANO[['original']][i,'RR1'] <- val

          # upper part of distribution
          if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
          } else {
            ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
            lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
          }          
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
          # coercing x into results.natopen.ANO dataframe
          results.natopen.ANO[['scaled']][i,'RR2'] <- scal() 
          results.natopen.ANO[['non-truncated']][i,'RR2'] <- scal.2() 
          results.natopen.ANO[['original']][i,'RR2'] <- val
        }

          # Light
          dat <- ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.natopen$GlobalID[i]),c('art_dekning','Light')]
          results.natopen.ANO[['original']][i,'richness'] <- nrow(dat)
          dat <- dat[!is.na(dat$Light),]

          if ( nrow(dat)>0 ) {

            val <- sum(dat[,'art_dekning'] * dat[,'Light'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
            # lower part of distribution
            if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
            } else {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            }            
            maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.natopen.ANO dataframe
            results.natopen.ANO[['scaled']][i,'Light1'] <- scal() 
            results.natopen.ANO[['non-truncated']][i,'Light1'] <- scal.2() 
            results.natopen.ANO[['original']][i,'Light1'] <- val

            # upper part of distribution
            if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
            } else {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            }            
            maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.natopen.ANO dataframe
            results.natopen.ANO[['scaled']][i,'Light2'] <- scal() 
            results.natopen.ANO[['non-truncated']][i,'Light2'] <- scal.2() 
            results.natopen.ANO[['original']][i,'Light2'] <- val
          }

          # Nitrogen
          dat <- ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.natopen$GlobalID[i]),c('art_dekning','Nitrogen')]
          results.natopen.ANO[['original']][i,'richness'] <- nrow(dat)
          dat <- dat[!is.na(dat$Nitrogen),]

          if ( nrow(dat)>0 ) {

            val <- sum(dat[,'art_dekning'] * dat[,'Nitrogen'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
            # lower part of distribution
            if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
            } else {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            }            
            maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.natopen.ANO dataframe
            results.natopen.ANO[['scaled']][i,'Nitrogen1'] <- scal() 
            results.natopen.ANO[['non-truncated']][i,'Nitrogen1'] <- scal.2() 
            results.natopen.ANO[['original']][i,'Nitrogen1'] <- val

            # upper part of distribution
            if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
            } else {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            }            
            maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.natopen.ANO dataframe
            results.natopen.ANO[['scaled']][i,'Nitrogen2'] <- scal() 
            results.natopen.ANO[['non-truncated']][i,'Nitrogen2'] <- scal.2() 
            results.natopen.ANO[['original']][i,'Nitrogen2'] <- val
          }

          # Soil_disturbance
          dat <- ANO.sp.ind[ANO.sp.ind$ParentGlobalID==as.character(ANO.natopen$GlobalID[i]),c('art_dekning','Soil_disturbance')]
          results.natopen.ANO[['original']][i,'richness'] <- nrow(dat)
          dat <- dat[!is.na(dat$Soil_disturbance),]

          if ( nrow(dat)>0 ) {

            val <- sum(dat[,'art_dekning'] * dat[,'Soil_disturbance'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
            # lower part of distribution
            if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
            } else {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            }           
            maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance1' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.natopen.ANO dataframe
            results.natopen.ANO[['scaled']][i,'Soil_disturbance1'] <- scal() 
            results.natopen.ANO[['non-truncated']][i,'Soil_disturbance1'] <- scal.2() 
            results.natopen.ANO[['original']][i,'Soil_disturbance1'] <- val

            # upper part of distribution
            if( ANO.natopen$kartleggingsenhet_1m2[i] %in% c("T2-C-7","T2-C-8") ) {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),"_BN",sep=""),'Gv']
            } else {
              ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Rv']
              lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'Gv']
            }
            maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance2' & natopen.ref.cov.val$grunn==as.character(results.natopen.ANO[['original']][i,"kartleggingsenhet_1m2"]),'maxmin']
            # coercing x into results.natopen.ANO dataframe
            results.natopen.ANO[['scaled']][i,'Soil_disturbance2'] <- scal() 
            results.natopen.ANO[['non-truncated']][i,'Soil_disturbance2'] <- scal.2() 
            results.natopen.ANO[['original']][i,'Soil_disturbance2'] <- val
          }

        }
      }

  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

# for using both sides of the plant indicators
results.natopen.ANO[['2-sided']] <- results.natopen.ANO[['non-truncated']]

# remove values >1 for 2-sided indicators
results.natopen.ANO[['2-sided']]$CC1[results.natopen.ANO[['2-sided']]$CC1>1] <- NA
results.natopen.ANO[['2-sided']]$CC2[results.natopen.ANO[['2-sided']]$CC2>1] <- NA

results.natopen.ANO[['2-sided']]$SS1[results.natopen.ANO[['2-sided']]$SS1>1] <- NA
results.natopen.ANO[['2-sided']]$SS2[results.natopen.ANO[['2-sided']]$SS2>1] <- NA

results.natopen.ANO[['2-sided']]$RR1[results.natopen.ANO[['2-sided']]$RR1>1] <- NA
results.natopen.ANO[['2-sided']]$RR2[results.natopen.ANO[['2-sided']]$RR2>1] <- NA

results.natopen.ANO[['2-sided']]$Light1[results.natopen.ANO[['2-sided']]$Light1>1] <- NA
results.natopen.ANO[['2-sided']]$Light2[results.natopen.ANO[['2-sided']]$Light2>1] <- NA

results.natopen.ANO[['2-sided']]$Nitrogen1[results.natopen.ANO[['2-sided']]$Nitrogen1>1] <- NA
results.natopen.ANO[['2-sided']]$Nitrogen2[results.natopen.ANO[['2-sided']]$Nitrogen2>1] <- NA

results.natopen.ANO[['2-sided']]$Soil_disturbance1[results.natopen.ANO[['2-sided']]$Soil_disturbance1>1] <- NA
results.natopen.ANO[['2-sided']]$Soil_disturbance2[results.natopen.ANO[['2-sided']]$Soil_disturbance2>1] <- NA

```

and for GRUK:
```{r, include=FALSE}

colnames(GRUK.variables)
levels(as.factor(GRUK.variables$Kartleggingsenhet)) # NiN types in data
levels(natopen.ref.cov.val$grunn) # NiN types in reference
#### creating dataframe to hold the results for natopens ####
# all GRUK points
nrow(GRUK.variables)
GRUK.natopen <- GRUK.variables

head(GRUK.natopen)
# update row-numbers
row.names(GRUK.natopen) <- 1:nrow(GRUK.natopen)
head(GRUK.natopen)
dim(GRUK.natopen)
colnames(GRUK.natopen)

length(levels(as.factor(GRUK.natopen$Flate_ID)))
length(levels(as.factor(GRUK.natopen$uPunkt_ID)))
summary(as.factor(GRUK.natopen$uPunkt_ID))
# none are double

unique(GRUK.natopen$Kartleggingsenhet)
GRUK.natopen$Kartleggingsenhet <- as.factor(GRUK.natopen$Kartleggingsenhet)
summary(GRUK.natopen$Kartleggingsenhet)

results.natopen.GRUK <- list()
ind <- unique(natopen.ref.cov.val$Ind)
# choose columns for site description
colnames(GRUK.natopen)
results.natopen.GRUK[['original']] <- GRUK.natopen
# drop geometry
st_geometry(results.natopen.GRUK[['original']]) <- NULL
results.natopen.GRUK[['original']] <- as.data.frame(results.natopen.GRUK[['original']])

# add columns for indicators
nvar.site <- ncol(results.natopen.GRUK[['original']])
for (i in 1:length(ind) ) {results.natopen.GRUK[['original']][,i+nvar.site] <- NA}
colnames(results.natopen.GRUK[['original']])[(nvar.site+1):(length(ind)+nvar.site)] <- paste(ind)
for (i in (nvar.site+1):(length(ind)+nvar.site) ) {results.natopen.GRUK[['original']][,i] <- as.numeric(results.natopen.GRUK[['original']][,i])}
summary(results.natopen.GRUK[['original']])
#results.natopen.GRUK[['original']]$Region <- as.factor(results.natopen.GRUK[['original']]$Region)
results.natopen.GRUK[['original']]$GlobalID <- as.factor(results.natopen.GRUK[['original']]$GlobalID)
results.natopen.GRUK[['original']]$Flate_ID <- as.factor(results.natopen.GRUK[['original']]$Flate_ID)
results.natopen.GRUK[['original']]$uPunkt_ID <- as.factor(results.natopen.GRUK[['original']]$uPunkt_ID)
results.natopen.GRUK[['original']]$Kartleggingsenhet <- as.factor(results.natopen.GRUK[['original']]$Kartleggingsenhet)

# roll out
results.natopen.GRUK[['scaled']] <- results.natopen.GRUK[['non-truncated']] <- results.natopen.GRUK[['original']]

```

```{r, echo = T, eval = F}

#### calculating scaled and non-truncated values for the indicators based on the dataset ####
for (i in 1:nrow(GRUK.natopen) ) {  #
  tryCatch({
    print(i)
    print(paste(GRUK.natopen$Flate_ID[i]))
    print(paste(GRUK.natopen$uPunkt_ID[i]))
    #    GRUK.natopen$Hovedoekosystem_sirkel[i]
    #    GRUK.natopen$Hovedoekosystem_rute[i]

    # if the GRUK.hovedtype exists in the reference
#    if (GRUK.natopen$hovedtype_rute[i] %in% unique(substr(natopen.ref.cov.val$grunn,1,2)) ) {

      # if there is any species present in current GRUK point  
      if ( length(GRUK.species.ind[GRUK.species.ind$ParentGlobalID==as.character(GRUK.natopen$GlobalID[i]),'Species']) > 0 ) {

        # Grime's C

        dat <- GRUK.species.ind[GRUK.species.ind$ParentGlobalID==as.character(GRUK.natopen$GlobalID[i]),c('art_dekning','CC')]
        results.natopen.GRUK[['original']][i,'richness'] <- nrow(dat)
        dat <- dat[!is.na(dat$CC),]

        if ( nrow(dat)>0 ) {

          val <- sum(dat[,'art_dekning'] * dat[,'CC'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
          # lower part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC1' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'CC1'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'CC1'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'CC1'] <- val

          # upper part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='CC2' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'CC2'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'CC2'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'CC2'] <- val

        }

        # Grime's S
        dat <- GRUK.species.ind[GRUK.species.ind$ParentGlobalID==as.character(GRUK.natopen$GlobalID[i]),c('art_dekning','SS')]
        results.natopen.GRUK[['original']][i,'richness'] <- nrow(dat)
        dat <- dat[!is.na(dat$SS),]

        if ( nrow(dat)>0 ) {

          val <- sum(dat[,'art_dekning'] * dat[,'SS'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
          # lower part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS1' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'SS1'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'SS1'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'SS1'] <- val

          # upper part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='SS2' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'SS2'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'SS2'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'SS2'] <- val

        }

        # Grime's R
        dat <- GRUK.species.ind[GRUK.species.ind$ParentGlobalID==as.character(GRUK.natopen$GlobalID[i]),c('art_dekning','RR')]
        results.natopen.GRUK[['original']][i,'richness'] <- nrow(dat)
        dat <- dat[!is.na(dat$RR),]

        if ( nrow(dat)>0 ) {

          val <- sum(dat[,'art_dekning'] * dat[,'RR'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
          # lower part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR1' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'RR1'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'RR1'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'RR1'] <- val

          # upper part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='RR2' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'RR2'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'RR2'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'RR2'] <- val

        }

        # Light
        dat <- GRUK.species.ind[GRUK.species.ind$ParentGlobalID==as.character(GRUK.natopen$GlobalID[i]),c('art_dekning','Light')]
        results.natopen.GRUK[['original']][i,'richness'] <- nrow(dat)
        dat <- dat[!is.na(dat$Light),]

        if ( nrow(dat)>0 ) {

          val <- sum(dat[,'art_dekning'] * dat[,'Light'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
          # lower part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light1' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'Light1'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'Light1'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'Light1'] <- val

          # upper part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Light2' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'Light2'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'Light2'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'Light2'] <- val

        }

        # Nitrogen
        dat <- GRUK.species.ind[GRUK.species.ind$ParentGlobalID==as.character(GRUK.natopen$GlobalID[i]),c('art_dekning','Nitrogen')]
        results.natopen.GRUK[['original']][i,'richness'] <- nrow(dat)
        dat <- dat[!is.na(dat$Nitrogen),]

        if ( nrow(dat)>0 ) {

          val <- sum(dat[,'art_dekning'] * dat[,'Nitrogen'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
          # lower part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen1' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'Nitrogen1'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'Nitrogen1'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'Nitrogen1'] <- val

          # upper part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Nitrogen2' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'Nitrogen2'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'Nitrogen2'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'Nitrogen2'] <- val

        }

        # Soil_disturbance
        dat <- GRUK.species.ind[GRUK.species.ind$ParentGlobalID==as.character(GRUK.natopen$GlobalID[i]),c('art_dekning','Soil_disturbance')]
        results.natopen.GRUK[['original']][i,'richness'] <- nrow(dat)
        dat <- dat[!is.na(dat$Soil_disturbance),]

        if ( nrow(dat)>0 ) {

          val <- sum(dat[,'art_dekning'] * dat[,'Soil_disturbance'],na.rm=T) / sum(dat[,'art_dekning'],na.rm=T)
          # lower part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance1' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance1' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'Soil_disturbance1'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'Soil_disturbance1'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'Soil_disturbance1'] <- val

          # upper part of distribution
          ref <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Rv']
          lim <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance2' & natopen.ref.cov.val$grunn==paste(as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),"_BN",sep=""),'Gv']
          maxmin <- natopen.ref.cov.val[natopen.ref.cov.val$Ind=='Soil_disturbance2' & natopen.ref.cov.val$grunn==as.character(results.natopen.GRUK[['original']][i,"Kartleggingsenhet"]),'maxmin']
          # coercing x into results.natopen.GRUK dataframe
          results.natopen.GRUK[['scaled']][i,'Soil_disturbance2'] <- scal() 
          results.natopen.GRUK[['non-truncated']][i,'Soil_disturbance2'] <- scal.2() 
          results.natopen.GRUK[['original']][i,'Soil_disturbance2'] <- val

        }

      }
#    }

  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

# for using both sides of the plant indicators
results.natopen.GRUK[['2-sided']] <- results.natopen.GRUK[['non-truncated']]

# remove values >1 for 2-sided indicators
results.natopen.GRUK[['2-sided']]$CC1[results.natopen.GRUK[['2-sided']]$CC1>1] <- NA
results.natopen.GRUK[['2-sided']]$CC2[results.natopen.GRUK[['2-sided']]$CC2>1] <- NA

results.natopen.GRUK[['2-sided']]$SS1[results.natopen.GRUK[['2-sided']]$SS1>1] <- NA
results.natopen.GRUK[['2-sided']]$SS2[results.natopen.GRUK[['2-sided']]$SS2>1] <- NA

results.natopen.GRUK[['2-sided']]$RR1[results.natopen.GRUK[['2-sided']]$RR1>1] <- NA
results.natopen.GRUK[['2-sided']]$RR2[results.natopen.GRUK[['2-sided']]$RR2>1] <- NA

results.natopen.GRUK[['2-sided']]$Light1[results.natopen.GRUK[['2-sided']]$Light1>1] <- NA
results.natopen.GRUK[['2-sided']]$Light2[results.natopen.GRUK[['2-sided']]$Light2>1] <- NA

results.natopen.GRUK[['2-sided']]$Nitrogen1[results.natopen.GRUK[['2-sided']]$Nitrogen1>1] <- NA
results.natopen.GRUK[['2-sided']]$Nitrogen2[results.natopen.GRUK[['2-sided']]$Nitrogen2>1] <- NA

results.natopen.GRUK[['2-sided']]$Soil_disturbance1[results.natopen.GRUK[['2-sided']]$Soil_disturbance1>1] <- NA
results.natopen.GRUK[['2-sided']]$Soil_disturbance2[results.natopen.GRUK[['2-sided']]$Soil_disturbance2>1] <- NA

```

```{r, include = F}

# Use cache data
results.natopen.ANO<-readRDS(paste0(here::here(),"/data/cache/results.natopen.ANO.RDS"))
results.natopen.GRUK<-readRDS(paste0(here::here(),"/data/cache/results.natopen.GRUK.RDS"))
```

```{r}
head(results.natopen.ANO[['2-sided']])
head(results.natopen.GRUK[['2-sided']])
```
#### Scaled value analyses

In order to visualize the results we need to rearrange the results-objects from wide to long format (note that there is both a lower and an upper condition indicator for each of the functional plant indicators).

```{r, message = FALSE}
#### plotting scaled values by main ecosystem type ####
## continuing with 2-sided
res.natopen.ANO <- results.natopen.ANO[['2-sided']]

# make long version of the scaled value part
res.natopen.ANO <-
  res.natopen.ANO %>% 
  pivot_longer(
    cols = c("CC1","CC2",
             "SS1","SS2",
             "RR1","RR2",
             "Light1","Light2",
             "Nitrogen1","Nitrogen2",
             "Soil_disturbance1","Soil_disturbance2"),
    names_to = "fp_ind",
    values_to = "scaled_value",
    values_drop_na = FALSE
  )

# add original values as well
res.natopen.ANO <- 
  res.natopen.ANO %>% add_column(original = results.natopen.ANO[['original']] %>% 
                               pivot_longer(
                                 cols = c("CC1","CC2",
                                          "SS1","SS2",
                                          "RR1","RR2",
                                          "Light1","Light2",
                                          "Nitrogen1","Nitrogen2",
                                          "Soil_disturbance1","Soil_disturbance2"),
                                 names_to = NULL,
                                 values_to = "original",
                                 values_drop_na = FALSE
                               ) %>%
                               pull(original)
  )

# similarly for GRUK
res.natopen.GRUK <- results.natopen.GRUK[['2-sided']]

# make long version of the scaled value part
res.natopen.GRUK <-
  res.natopen.GRUK %>% 
  pivot_longer(
    cols = c("CC1","CC2",
             "SS1","SS2",
             "RR1","RR2",
             "Light1","Light2",
             "Nitrogen1","Nitrogen2",
             "Soil_disturbance1","Soil_disturbance2"),
    names_to = "fp_ind",
    values_to = "scaled_value",
    values_drop_na = FALSE
  )

# add original values as well
res.natopen.GRUK <- 
  res.natopen.GRUK %>% add_column(original = results.natopen.GRUK[['original']] %>% 
                                   pivot_longer(
                                     cols = c("CC1","CC2",
                                              "SS1","SS2",
                                              "RR1","RR2",
                                              "Light1","Light2",
                                              "Nitrogen1","Nitrogen2",
                                              "Soil_disturbance1","Soil_disturbance2"),
                                     names_to = NULL,
                                     values_to = "original",
                                     values_drop_na = FALSE
                                   ) %>%
                                   pull(original)
  )

```

### Ecosystem sub-types
And we can show the resulting scaled values as Violin plots for each indicator and main ecosystem type 
```{r,echo=FALSE,message=FALSE, warning=FALSE}

# making the plot, ANO, CSR-indicators
ggplot(data=subset(res.natopen.ANO,!is.na(scaled_value) & fp_ind %in% c("CC1","SS1","RR1",
                                                                        "CC2","SS2","RR2")), 
       aes(x=factor(kartleggingsenhet_1m2), y=scaled_value, fill=fp_ind)) + 
  geom_hline(yintercept=0.6, linetype="dashed") + 
  geom_violin(color=NA) +
  #  geom_boxplot(width=0.2, color="grey") +
  geom_point(size=1, shape=16, color="black") +
  facet_wrap(~factor(fp_ind,levels=c("CC1","SS1","RR1",
                                     "CC2","SS2","RR2"),
                     labels=c("CSR-C1","CSR-S1","CSR-R1",
                                     "CSR-C2","CSR-S2","CSR-R2")
                     ), ncol = 3) + 
  xlab("Basic ecosystem type") + 
  ylab("Scaled indicator value (ANO data, CSR indicators)") +
  scale_fill_discrete(labels=c("CSR-C1","CSR-S1","CSR-R1",
                                     "CSR-C2","CSR-S2","CSR-R2")) + 
  theme(legend.position="none") +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust=0.2))

# making the plot, ANO, Tyler-indicators
ggplot(data=subset(res.natopen.ANO,!is.na(scaled_value) & fp_ind %in% c("Light1","Nitrogen1","Soil_disturbance1",
                                                                        "Light2","Nitrogen2","Soil_disturbance2")), 
       aes(x=factor(kartleggingsenhet_1m2), y=scaled_value, fill=fp_ind)) + 
  geom_hline(yintercept=0.6, linetype="dashed") + 
  geom_violin(color=NA) +
#  geom_boxplot(width=0.2, color="grey") +
  geom_point(size=1, shape=16, color="black") +
  facet_wrap(~factor(fp_ind,levels=c("Light1","Nitrogen1","Soil_disturbance1",
                                     "Light2","Nitrogen2","Soil_disturbance2")), ncol = 3) + 
  xlab("Basic ecosystem type") + 
  ylab("Scaled indicator value (ANO data, Tyler indicators)") + 
  theme(legend.position="none") +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust=0.2))

# making the plot, GRUK
ggplot(res.natopen.GRUK, aes(x=factor(Kartleggingsenhet), y=scaled_value, fill=fp_ind)) + 
  geom_hline(yintercept=0.6, linetype="dashed") + 
  geom_violin(color = NA) +
  #  geom_boxplot(width=0.2, color="grey") +
  geom_point(size=1, shape=16, color="black") +
  facet_wrap(~factor(fp_ind,levels=c("CC1","SS1","RR1","Light1","Nitrogen1","Soil_disturbance1",
                                     "CC2","SS2","RR2","Light2","Nitrogen2","Soil_disturbance2"),
                     labels=c("CSR-C1","CSR-S1","CSR-R1","Light1","Nitrogen1","Soil_disturbance1",
                                     "CSR-C2","CSR-S2","CSR-R2","Light2","Nitrogen2","Soil_disturbance2")), ncol = 6) + 
  xlab("Basic ecosystem type") + 
  ylab("Scaled indicator value (GRUK data, all indicators)") + 
  theme(legend.position="none") +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust=0.2))

```
The ANO results show a high frequency of too low CSR-R values (CSR-R1) and Soil disturbance while also showing signs of higher than expected competition (CSR-C2), which  indicates that a number of naturally open areas are changing in their dominance structure and physical stability. The GRUK data show a generally large spread of scaled values, but the violin plots (the shapes indicate the relative amount of observations across the y-axis) suggest that there is a number of sites with too high nitrogen (Nitrogen2), and some sites with too much disturbance (CSR-R2) and light (Light2). Many GRUK sites are popular outdoor locations with the local population, which could increase disturbance and thus cause deviations towards high CSR-R values (CSR-R2).
Increased nitrogen may be an effect of alien species, which the GRUK data records as a major pressure and problem for many monitoring sites.

```{r}
ggplot(res.natopen.GRUK[res.natopen.GRUK$fp_ind=="Nitrogen2",], aes(x=fremmedartsdekning, y=scaled_value)) +
  geom_point() +
  xlab("Alien species cover (%)") + ylab("Scaled Nitrogen2 value (GRUK data)")
```
And indeed, we can see that very high alien species cover predicts poor scores in the Nitrogen2 indicator.

### Indicator index maps
We can also show the results as a map, for instance for CSR-R1 (the lower ruderal/disturbance indicator) for ANO and Nitrogen2 (the upper nitrogen indicator) for GRUK, either by directly plotting the data onto the map...
```{r, echo=FALSE, warning=FALSE,message = FALSE}
#### scaled value maps ####
# keep wide format and add geometry again
res.natopen.ANO2 <- results.natopen.ANO[['2-sided']]
st_geometry(res.natopen.ANO2) <- st_geometry(ANO.natopen)

res.natopen.GRUK2 <- results.natopen.GRUK[['2-sided']]
st_geometry(res.natopen.GRUK2) <- st_geometry(GRUK.natopen)

nor <- st_read("data/outlineOfNorway_EPSG25833.shp")%>%
  st_as_sf() %>%
  st_transform(crs = st_crs(ANO.geo))

reg <- st_read("data/regions.shp")%>%
  st_as_sf() %>%
  st_transform(crs = st_crs(ANO.geo))

# change region names to something R-friendly
#reg$region
reg$region <- c("Northern Norway","Central Norway","Eastern Norway","Western Norway","Southern Norway")

regnor <- st_intersection(reg,nor)

## scaled value maps
# ANO, RR1 (lower indicator)
tm_shape(regnor) +
  tm_fill('GID_0', labels="", title="", legend.show = FALSE) + 
  tm_borders() +
  tm_shape(res.natopen.ANO2) +
  tm_dots('RR1',midpoint=NA, palette=tmaptools::get_brewer_pal("YlOrRd", 7, plot = FALSE), scale=2, legend.show = FALSE) + # 
  tm_layout(main.title = "Ruderal index (lower), natopen ANO",legend.position = c("right", "bottom"), main.title.size=1.2) + 
  tm_add_legend(type = "fill", 
                col = c(tmaptools::get_brewer_pal("YlOrRd", 6, plot = FALSE),'grey'),
                labels = c("0.4 - 0.5", "0.5 - 0.6", "0.6 - 0.7", 
                           "0.7 - 0.8", "0.8 - 0.9", "0.9 - 1.0", "NA"),
                title = "index values")

# GRUK, Nitrogen2 (upper indicator)
tm_shape(regnor) +
  tm_fill('GID_0', labels="", title="", legend.show = FALSE) + 
  tm_borders() +
  tm_shape(res.natopen.GRUK2) +
  tm_dots('Nitrogen2',midpoint=NA, palette=tmaptools::get_brewer_pal("YlOrRd", 7, plot = FALSE), scale=2, legend.show = FALSE) + # 
  tm_layout(main.title = "Nitrogen index (upper), natopen GRUK",legend.position = c("right", "bottom"), main.title.size=1.2) + 
  tm_add_legend(type = "fill", 
                col = c(tmaptools::get_brewer_pal("YlOrRd", 7, plot = FALSE),'grey'),
                labels = c("0.3 - 0.4", "0.4 - 0.5", "0.5 - 0.6", "0.6 - 0.7", 
                           "0.7 - 0.8", "0.8 - 0.9", "0.9 - 1.0", "NA"),
                title = "index values")

```
For GRUK we can zoom in on the area around the Oslofjord...

```{r}
boks <- st_bbox(c(xmin = 10.3, xmax = 10.8, ymax = 59.95, ymin = 59.4), crs = st_crs(4326))

tm_shape(regnor,bbox=boks) +
  tm_fill('GID_0', labels="", title="", legend.show = FALSE) + 
  tm_borders() +
  tm_shape(res.natopen.GRUK2) +
  tm_dots('Nitrogen2',midpoint=NA, palette=tmaptools::get_brewer_pal("YlOrRd", 7, plot = FALSE), scale=3, legend.show = FALSE) + # 
  tm_layout(main.title = "Nitrogen index (upper), natopen GRUK",legend.position = c("right", "bottom"), main.title.size=1.2) + 
  tm_add_legend(type = "fill", 
                col = c(tmaptools::get_brewer_pal("YlOrRd", 7, plot = FALSE),'grey'),
                labels = c("0.3 - 0.4", "0.4 - 0.5", "0.5 - 0.6", "0.6 - 0.7", 
                           "0.7 - 0.8", "0.8 - 0.9", "0.9 - 1.0", "NA"),
                title = "index values")
```
...but for ANO the colors and values of the data points on the map are hard to make out.

#### continue here ####

### Regions - maps and statistics
Alternatively we can calculate and show the region-wise means and their related standard errors. But note that calculating a simple mean would be inappropriate for these data. This is because:
(i) the scaled data are bound between 0 and 1, and thus follow a beta-distribution rather than a Gaussian one
(ii) both the ANO and the ASO datasets have a nested structure

Therefore, we need to (i) use a beta-model, that (ii) can account for the nested structure of the data.
Here, we apply the following function using either a glmmTMB null-model with a beta-distribution, logit link, and the nesting as a random intercept, or a simple betareg null-model with logit link if the nesting is not extensive enough for a mixed model.
```{r}
library(betareg)
library(glmmTMB)

expit <- function(L) exp(L) / (1+exp(L)) # since the beta-models use a logit link, we need to calculate the estimates back to the identity scale

# the function performs a glmmTMB if there's >= 5 random levels in the nesting structure
# if that is not the case, then the function performs a betareg if theres >= 2 observations
# if that is not the case either, then the function returns the value of the single observation with NA standard error

indmean.beta <- function(df) {

  st_geometry(df) <- NULL
  colnames(df) <- c("y","ran")

  if ( nrow(df[!is.na(df[,1]),]) >= 2 ) {

    if ( length(unique(df[!is.na(df[,1]),2])) >=5 ) {

      mod1 <- glmmTMB(y ~ 1 +(1|ran), family=beta_family(), data=df)

      return(c(
        expit(summary( mod1 )$coefficients$cond[1]),

        expit( summary( mod1 )$coefficients$cond[1] + 
                 summary( mod1 )$coefficients$cond[2] )-
          expit( summary( mod1 )$coefficients$cond[1] ),

        nrow(df[!is.na(df$y),]),
        summary( mod1 )$coefficients$cond[1],
        summary( mod1 )$coefficients$cond[2]
      ))

    } else {

      mod2 <- betareg(y ~ 1, data=df)

      return(c(
        expit(summary( mod2 )$coefficients$mean[1]),
        expit( summary( mod2 )$coefficients$mean[1] + 
                 summary( mod2 )$coefficients$mean[2] )-
          expit( summary( mod2 )$coefficients$mean[1] ),
        nrow(df[!is.na(df$y),]),
        summary( mod2 )$coefficients$mean[1],
        summary( mod2 )$coefficients$mean[2]
      ))

    }

  } else {

    return(c(df$y,NA,1,NA,NA))

  }

}

```

```{r}

res.seminat.ANO2 = st_join(res.seminat.ANO2, regnor, left = TRUE)
res.seminat.ASO2 = st_join(res.seminat.ASO2, regnor, left = TRUE)

regnor <- regnor %>%
  mutate(
    Grazing_mowing1.ANO.reg.mean = c(indmean.beta(df=res.seminat.ANO2[res.seminat.ANO2$region=="Northern Norway",c("Grazing_mowing1","ano_flate_id")])[1],
                                 indmean.beta(df=res.seminat.ANO2[res.seminat.ANO2$region=="Central Norway",c("Grazing_mowing1","ano_flate_id")])[1],
                                 indmean.beta(df=res.seminat.ANO2[res.seminat.ANO2$region=="Eastern Norway",c("Grazing_mowing1","ano_flate_id")])[1],
                                 indmean.beta(df=res.seminat.ANO2[res.seminat.ANO2$region=="Western Norway",c("Grazing_mowing1","ano_flate_id")])[1],
                                 indmean.beta(df=res.seminat.ANO2[res.seminat.ANO2$region=="Southern Norway",c("Grazing_mowing1","ano_flate_id")])[1]
                                 ),
    Grazing_mowing1.ANO.reg.se = c(indmean.beta(df=res.seminat.ANO2[res.seminat.ANO2$region=="Northern Norway",c("Grazing_mowing1","ano_flate_id")])[2],
                               indmean.beta(df=res.seminat.ANO2[res.seminat.ANO2$region=="Central Norway",c("Grazing_mowing1","ano_flate_id")])[2],
                               indmean.beta(df=res.seminat.ANO2[res.seminat.ANO2$region=="Eastern Norway",c("Grazing_mowing1","ano_flate_id")])[2],
                               indmean.beta(df=res.seminat.ANO2[res.seminat.ANO2$region=="Western Norway",c("Grazing_mowing1","ano_flate_id")])[2],
                               indmean.beta(df=res.seminat.ANO2[res.seminat.ANO2$region=="Southern Norway",c("Grazing_mowing1","ano_flate_id")])[2]
                               ),
    Grazing_mowing1.ANO.reg.n = c(nrow(res.seminat.ANO2[res.seminat.ANO2$region=="Northern Norway" & !is.na(res.seminat.ANO2$Grazing_mowing1),]),
                              nrow(res.seminat.ANO2[res.seminat.ANO2$region=="Central Norway" & !is.na(res.seminat.ANO2$Grazing_mowing1),]),
                              nrow(res.seminat.ANO2[res.seminat.ANO2$region=="Eastern Norway" & !is.na(res.seminat.ANO2$Grazing_mowing1),]),
                              nrow(res.seminat.ANO2[res.seminat.ANO2$region=="Western Norway" & !is.na(res.seminat.ANO2$Grazing_mowing1),]),
                              nrow(res.seminat.ANO2[res.seminat.ANO2$region=="Southern Norway" & !is.na(res.seminat.ANO2$Grazing_mowing1),])
                              )
  )

# and adding the values for ASO
regnor <- regnor %>%
  mutate(
    Nitrogen2.ASO.reg.mean = c(indmean.beta(df=res.seminat.ASO2[res.seminat.ASO2$region=="Northern Norway",c("Nitrogen2","Omradenummer_flatenummer")])[1],
                               indmean.beta(df=res.seminat.ASO2[res.seminat.ASO2$region=="Central Norway",c("Nitrogen2","Omradenummer_flatenummer")])[1],
                               indmean.beta(df=res.seminat.ASO2[res.seminat.ASO2$region=="Eastern Norway",c("Nitrogen2","Omradenummer_flatenummer")])[1],
                               indmean.beta(df=res.seminat.ASO2[res.seminat.ASO2$region=="Western Norway",c("Nitrogen2","Omradenummer_flatenummer")])[1],
                               indmean.beta(df=res.seminat.ASO2[res.seminat.ASO2$region=="Southern Norway",c("Nitrogen2","Omradenummer_flatenummer")])[1]
                               ),
    Nitrogen2.ASO.reg.se = c(indmean.beta(df=res.seminat.ASO2[res.seminat.ASO2$region=="Northern Norway",c("Nitrogen2","Omradenummer_flatenummer")])[2],
                             indmean.beta(df=res.seminat.ASO2[res.seminat.ASO2$region=="Central Norway",c("Nitrogen2","Omradenummer_flatenummer")])[2],
                             indmean.beta(df=res.seminat.ASO2[res.seminat.ASO2$region=="Eastern Norway",c("Nitrogen2","Omradenummer_flatenummer")])[2],
                             indmean.beta(df=res.seminat.ASO2[res.seminat.ASO2$region=="Western Norway",c("Nitrogen2","Omradenummer_flatenummer")])[2],
                             indmean.beta(df=res.seminat.ASO2[res.seminat.ASO2$region=="Southern Norway",c("Nitrogen2","Omradenummer_flatenummer")])[2]
                             ),
    Nitrogen2.ASO.reg.n = c(nrow(res.seminat.ASO2[res.seminat.ASO2$region=="Northern Norway" & !is.na(res.seminat.ASO2$Nitrogen2),]),
                            nrow(res.seminat.ASO2[res.seminat.ASO2$region=="Central Norway" & !is.na(res.seminat.ASO2$Nitrogen2),]),
                            nrow(res.seminat.ASO2[res.seminat.ASO2$region=="Eastern Norway" & !is.na(res.seminat.ASO2$Nitrogen2),]),
                            nrow(res.seminat.ASO2[res.seminat.ASO2$region=="Western Norway" & !is.na(res.seminat.ASO2$Nitrogen2),]),
                            nrow(res.seminat.ASO2[res.seminat.ASO2$region=="Southern Norway" & !is.na(res.seminat.ASO2$Nitrogen2),]))
  )

## scaled value maps for Grazing_mowing1 (lower indicator), ANO
# mean
tm_shape(regnor) +
  tm_polygons(col="Grazing_mowing1.ANO.reg.mean", title="Grazing_mowing (lower), mean", style="quantile", palette=rev(get_brewer_pal(palette="OrRd", n=5, plot=FALSE))) +
  tm_text("Grazing_mowing1.ANO.reg.n",col="black",bg.color="grey")
```
Mean index value by region for the upper Grazing_mowing indicator (i.e. index shows deviations towards less grazing or mowing) from the ANO monitoring data. Numbers in grey fields show the number of observations in the respective region.

```{r}
# se
tm_shape(regnor) +
  tm_polygons(col="Grazing_mowing1.ANO.reg.se", title="Grazing_mowing (lower), SE", style="quantile", palette=get_brewer_pal(palette="OrRd", n=5, plot=FALSE)) +
  tm_text("Grazing_mowing1.ANO.reg.n",col="black",bg.color="grey")

```
Standard error to the mean index value by region for the lower Grazing_mowing indicator from the ANO monitoring data. Numbers in grey fields show the number of observations in the respective region.

And here are the corresponding maps for the upper Nitrogen indicator in the ASO data:
```{r}
## scaled value maps for Nitrogen2 (upper indicator), ASO
# mean
tm_shape(regnor) +
  tm_polygons(col="Nitrogen2.ASO.reg.mean", title="Nitrogen (upper), mean", style="quantile", palette=rev(get_brewer_pal(palette="OrRd", n=5, plot=FALSE))) +
  tm_text("Nitrogen2.ASO.reg.n",col="black",bg.color="grey")
```
Mean index value by region for the upper Nitrogen indicator (i.e. index shows deviations towards increased Nitrogen availability) from the ASO monitoring data. Numbers in grey fields show the number of observations in the respective region.

```{r}
# se
tm_shape(regnor) +
  tm_polygons(col="Nitrogen2.ASO.reg.se", title="Nitrogen (upper), SE", style="quantile", palette=get_brewer_pal(palette="OrRd", n=5, plot=FALSE)) +
  tm_text("Nitrogen2.ASO.reg.n",col="black",bg.color="grey")

```
Standard error to the mean index value by region for the upper Nitrogen indicator from the ASO monitoring data. Numbers in grey fields show the number of observations in the respective region.

#### Unscaled values vs. reference
We can also compare the unscaled values to the reference distribution in order to identify ecosystem types and functional plant indicators showing a deviation from the expectation. Since Grazing_mowing for ANO and nitrogen for ASO show some deviations  we exemplify this with these indicators for unscaled values.

Grazing_mowing, ANO:
```{r}
#summary(res.seminat$kartleggingsenhet_1m2)
#length(unique(res.seminat$kartleggingsenhet_1m2))
# 19 NiN-types, of which T41-C-1 has 0 observations, and T32 & T34 as main types don't have a reference -> so, 16 NiN-Types to plot
#colnames(seminat.ref.cov[['Light']])

### Grazing_mowing
par(mfrow=c(2,4))
## T32s
# T32-C-1
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,"T32-C1C2"]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C1',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-1",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-1",]$original)),
       col="red")

# T32-C-2
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,"T32-C1C2"]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C2',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-2",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-2",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-2",]$original)),
       col="red")

# T32-C-3
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,"T32-C3C4"]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C3',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-3",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-3",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-3",]$original)),
       col="red")

# T32-C-4
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,"T32-C3C4"]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C4',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-4",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-4",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-4",]$original)),
       col="red")

# T32-C-5
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,c("T32-C5C20a","T32-C5C20b")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C5',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-5",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-5",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-5",]$original)),
       col="red")

# T32-C-6
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,c("T32-C21C6a","T32-C21C6b")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C6',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-6",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-6",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-6",]$original)),
       col="red")

# T32-C-9
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,c("T32-C9a","T32-C9b")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C9',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-9",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-9",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-9",]$original)),
       col="red")

# T32-C-10
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,c("T32-C10a","T32-C10b")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C10',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-10",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-10",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-10",]$original)),
       col="red")

par(mfrow=c(2,4))
# T32-C-15
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,c("T32-C15")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C15',xlab='Grazing_mowing value')
#points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-15",]$original,na.rm=T),
#       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-15",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-15",]$original)),
       col="red")

# T32-C-20
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,c("T32-C5C20a","T32-C5C20b")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C20',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-20",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-20",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T32-C-20",]$original)),
       col="red")

## V10s
# V10-C-1
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,"V10-C1C2"]) ,na.rm=T),
     xlim=c(1,8), type="l", main='V10-C1',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="V10-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="V10-C-1",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="V10-C-1",]$original)),
       col="red")

# V10-C-3
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,"V10-C3"]) ,na.rm=T),
     xlim=c(1,8), type="l", main='V10-C3',xlab='Grazing_mowing value')
#points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="V10-C-3",]$original,na.rm=T),
#       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="V10-C-3",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="V10-C-3",]$original)),
       col="red")

## T34s
# T34-C-1
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,"T34-C1"]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T34-C1',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T34-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T34-C-1",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T34-C-1",]$original)),
       col="red")

# T34-C-2
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,c("T34-C2a","T34-C2b","T34-C2c")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T34-C2',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T34-C-2",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T34-C-2",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T34-C-2",]$original)),
       col="red")

# T34-C-4
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,c("T34-C4a","T34-C4b","T34-C4c","T34-C4d")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T34-C4',xlab='Grazing_mowing value')
#points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T34-C-4",]$original,na.rm=T),
#       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T34-C-4",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T34-C-4",]$original)),
       col="red")

## T41
plot(density( as.matrix(seminat.ref.cov[['Grazing_mowing']][,c("T41a","T41b")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T41',xlab='Grazing_mowing value')
points(density(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T41",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T41",]$original,
       rep(0,length(res.seminat.ANO[res.seminat.ANO$fp_ind=="Grazing_mowing1" & res.seminat.ANO$kartleggingsenhet_1m2=="T41",]$original)),
       col="red")
legend("topleft", legend=c("reference","field data"), pch=c(NA,1), lty=1, col=c("black","red"))

```

Nitrogen, ASO
```{r}

#summary(res.seminat.ASO$NiN_grunntype)
#length(levels(res.seminat.ASO$NiN_grunntype))
# 12 NiN-types, of which C-13 and C-2 have 0 observations -> so, 10 NiN-Types to plot
#colnames(seminat.ref.cov[['Nitrogen']])

### Nitrogen
par(mfrow=c(3,4))
# T32-C-1
plot(density( as.matrix(seminat.ref.cov[['Nitrogen']][,"T32-C1C2"]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C1',xlab='Nitrogen value')
points(density(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-1",]$original,
       rep(0,length(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-1",]$original)),
       col="red")

# T32-C-3
plot(density( as.matrix(seminat.ref.cov[['Nitrogen']][,"T32-C3C4"]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C3',xlab='Nitrogen value')
points(density(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-3",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-3",]$original,
       rep(0,length(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-3",]$original)),
       col="red")

# T32-C-4
plot(density( as.matrix(seminat.ref.cov[['Nitrogen']][,"T32-C3C4"]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C4',xlab='Nitrogen value')
points(density(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-4",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-4",]$original,
       rep(0,length(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-4",]$original)),
       col="red")

# T32-C-5
plot(density( as.matrix(seminat.ref.cov[['Nitrogen']][,c("T32-C5C20a","T32-C5C20b")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C5',xlab='Nitrogen value')
points(density(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-5",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-5",]$original,
       rep(0,length(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-5",]$original)),
       col="red")

# T32-C-6
plot(density( as.matrix(seminat.ref.cov[['Nitrogen']][,c("T32-C21C6a","T32-C21C6b")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C6',xlab='Nitrogen value')
points(density(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-6",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-6",]$original,
       rep(0,length(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-6",]$original)),
       col="red")

# T32-C-7
plot(density( as.matrix(seminat.ref.cov[['Nitrogen']][,c("T32-C7C8")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C7',xlab='Nitrogen value')
#points(density(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-7",]$original,na.rm=T),
#       type="l", col="red")
points(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-7",]$original,
       rep(0,length(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-7",]$original)),
       col="red")

# T32-C-8
plot(density( as.matrix(seminat.ref.cov[['Nitrogen']][,c("T32-C7C8")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C8',xlab='Nitrogen value')
points(density(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-8",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-8",]$original,
       rep(0,length(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-8",]$original)),
       col="red")

# T32-C-9
plot(density( as.matrix(seminat.ref.cov[['Nitrogen']][,c("T32-C9a","T32-C9b")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C9',xlab='Nitrogen value')
#points(density(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-9",]$original,na.rm=T),
#       type="l", col="red")
points(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-9",]$original,
       rep(0,length(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-9",]$original)),
       col="red")

# T32-C-10
plot(density( as.matrix(seminat.ref.cov[['Nitrogen']][,c("T32-C10a","T32-C10b")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C10',xlab='Nitrogen value')
#points(density(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-10",]$original,na.rm=T),
#       type="l", col="red")
points(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-10",]$original,
       rep(0,length(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-10",]$original)),
       col="red")

# T32-C-20
plot(density( as.matrix(seminat.ref.cov[['Nitrogen']][,c("T32-C5C20a","T32-C5C20b")]) ,na.rm=T),
     xlim=c(1,8), type="l", main='T32-C20',xlab='Nitrogen value')
points(density(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-20",]$original,na.rm=T),
       type="l", col="red")
points(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-20",]$original,
       rep(0,length(res.seminat.ASO[res.seminat.ASO$fp_ind=="Nitrogen2" & res.seminat.ASO$NiN_grunntype=="T32-C-20",]$original)),
       col="red")

legend("topleft", legend=c("reference","field data"), pch=c(NA,1), lty=1, col=c("black","red"))

```
The figure shows that deviations mainly occur in V2-C1, which represents limestone-poor swamp forests. According to the functional signature from the plant community composition many of the occurrences of this ecosystem type are too acidic and may have too low availability of nitrogen.

### Eksport file (final product)
<!-- Export final file. Ideally a georeferenced shape or raster wit indicators values (raw and normalised), reference values and errors. -->