### 2023-09-13 DRAFT_NDVI_wetland.Rmd
ndvi wetland year regressions for good condition

# NDVI (greenness) indicator, wetland ecosystems  {#NDVI-indicator-wetland}

<br />

_Author and date:_
Joachim Töpper, Tessa Bargmann

June 2023

<br />

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(downloader)
library(data.table)
library(lubridate)
library(sf)
library(plyr)
library(stringr)
library(tidyverse)
library(RColorBrewer)
library("gridExtra") 
library(ggridges)
library(tmap)
library(tmaptools)
library(betareg)
library(StepBeta)
library(glmmTMB)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
Ecosystem <- "våtmark"
Egenskap  <- c("Primærproduksjon")
ECT       <- "Structural state characteristics"
Contact   <- "Joachim Töpper"
```

```{r, echo=FALSE}
metaData <- data.frame(Ecosystem,
                       "Økologisk egenskap" = Egenskap,
                       "ECT class" = ECT)
knitr::kable(metaData)
```

<!-- Don't remove these three html lines -->
<br />
<br />
<hr />

## Introduction
The Normalized difference vegetation index (NDVI) can be used to describe the greenness of an area, and may thus represent a proxy for the amount of chlorophyll and plant productivity in the given area. Different ecosystem types naturally have greenness signals of different intensity but also with a certain variance. Deviations beyond the natural variation in either direction may indicate a reduction in ecological condition of a given ecosystem. E.g. lower than normal NDVI may indicate excessive disturbance, while higher than normal NDVI may indicate regrowth and transitions towards more densely vegetated systems. In recent assessments of ecological condition for forest and mountain ecosystems, two different NDVI-based indicators were developed and applied. For forests, NDVI-values across the country were compared to a spatial reference NDVI-signal from protected areas. For mountains, the extent and consistence of the temporal change of NDVI-values over the last twenty years was assessed. In both cases, the NDVI-product from the MODIS satellite reflectance imagery was used based on available maps of spatial ecosystem extent in Norway (i.e. national maps for the forest and mountain ecosystems). Here, we explore the potential for developing an NDVI-indicator for ecological condition representing open wetland ecosystems (i.e. excluding forest swamps). For this task, two initial challenges can be identified early on:
1) We don't yet have a national wetland map, i.e. we don't know the exact locations of all wetlands in Norway, which is a prerequisite to extract satellite data for the right locations and ecosystems, and thus achieve a good and unbiased spatial representation for the assessment area.
2) The spatial extent of the single wetland ecosystem occurrences (hereafter 'polygons') includes many wetland areas which are way smaller than the pixel size in the MODIS imagery (250 x 250 m) which was used for forests and mountains and has a time series back to year 2000.

In this draft development for a wetland NDVI indicator for ecological condition we approach these challenges in the following ways:
1) We use ecosystem occurrence data from NiN-mapping to guide the extraction of satellite NDVI data for pixels representing the right ecosystem. However, the ecosystem occurrences in the NiN data are highly spatially biased and thus cannot provide a good and unbiased spatial representation for the assessment area. Our work therefore represents a concept and method development which may be applied in an assessment of ecological condition once spatially unbiased occurrence data or a national map of wetland ecosystems are available.
2) We include Sentinel-2 imagery for wetland ecosystems as the resolution for the bands underlying the NDVI product here is 10 x 10 m which is smaller than almost all wetland polygons in the NiN-data. However, since Sentinel-2 only goes back to 2015, we lose much of the time series aspect as compared to the MODIS data. In addition, we still explore NDVI data from the MODIS and Landsat satellites (see more details below).

## About the underlying data
In the NDVI project for wetlands, we use four sets of data for building indicators for ecological condition:

- NiN ecosystem data from 'Naturtypekartlegging etter Miljødirektoratets innstruks' to guide extraction of NDVI data for the right locations and to provide information on the field-assessed ecological condition of mapped polygons
- Sentinel-2 satellite NDVI data
- MODIS satellite NDVI data
- Landsat satellite NDVI data

(1) NiN ecosystem data:

(2) Sentinel-2 satellite NDVI data:

(3) MODIS satellite NDVI data:

(4) Landsat satellite NDVI data:

have to exclude all polygons smaller than the MODIS pixel size of 250 x 250 m, which reduces the number of available polygons greatly. To explore temporal trends even further back in time, we also include Landsat NDVI data, which span a time frame from 1984 to 2021, and have a resolution of 30 x 30 m, thus excluding only few polygons of insufficient size.

### Representativity in time and space
For wetlands, the NiN data contain 6503 wetland polygons of known area size across mainland Norway. However, NiN-mapping is not planned and performed in a spatially representative way and thus both the ecosystem occurrences and their spatial extent have to be treated as spatially biased. The 6503 wetland polygons are distributed across regions and main ecosystem types in the following way:

----build a matrix-table----

The NDVI data from either satellite follow the spatial distribution and bias status of the NiN-data. Due to the larger number of excluded polygons in the NDVI data from MODIS (because of the large pixel size in MODIS, see above), the MODIS NDVI data for the wetland occurrences in the NiN data likely are even more spatially biased than it is the case for the NDVI data from Sentinel-2 and Landsat.

### Temporal coverage
The currently available NiN data on wetland ecosystems span a time period from 2018-2021 and thus represent a contemporary sample of ecosystem occurrences. Note that none of the polygons in the NiN-data has been visited more than once during that period.

Available Sentinel-2 satellite NDVI data span the years 2015-2022.
Available MODIS satellite NDVI data span the years 2000-2022.
Available Landsat satellite NDVI data span the years 1984-2021.

## Collinearities with other indicators
The NDVI signal may likely be co-linear with the re-growth indicator based on LiDAr derived canopy height in drained wetland areas under shrub and tree encroachment.

## Reference state and values
### Reference state
This indicator project is mainly exploratory, mapping the possibilities of and lacking prerequisites for satellite derived NDVI data to inform ecological condition assessments. Therefore, we lack a concise strategy for how to define the reference state, especially prior to having access to an ecosystem map for wetland ecosystems. In this exploratory work, we will use the condition classification in the NiN-data to explore its applicability for defining a reference state. We will use statistical models to test whether or not NDVI varies systematically with condition and across time.

### Reference values, thresholds for defining _good ecological condition_, minimum and/or maximum values
We will explore (i) if the condition classes in the NiN data can be used to define scaling values, and (ii) the potential for taking a probabilistic approach to scaling original NDVI values to a condition index based on probability density curves derived from above named models.

## Uncertainties
Given a condition index can be achieved, we can calculate a mean indicator value (after scaling) for every region (or any other delimited area of interest) as well as its corresponding standard error and standard deviation as a measure of spatial uncertainty for a geographical area.

## References

----not referred to in text yet, but likely will :)----

Framstad, E., Kolstad, A. L., Nybø, S., Töpper, J. & Vandvik, V. 2022. The condition of forest and mountain ecosystems in Norway. Assessment by the IBECA method. NINA Report 2100. Norwegian Institute for Nature Research.

Halvorsen, R., Skarpaas, O., Bryn, A., Bratli, H., Erikstad, L., Simensen, T., & Lieungh, E. (2020). Towards a systematics of ecodiversity: The EcoSyst framework. Global Ecology and Biogeography, 29(11), 1887-1906. doi:10.1111/geb.13164

Jakobsson, S., Töpper, J.P., Evju, M., Framstad, E., Lyngstad, A., Pedersen, B., Sickel, H., Sverdrup-Thygeson, A., Vandvik. V., Velle, L.G., Aarrestad, P.A. & Nybø, S. 2020. Setting reference levels and limits for good ecological condition in terrestrial ecosystems. Insights from a case study based on the IBECA approach. Ecological Indicators 116: 106492.

Töpper, J. & Jakobsson, S. 2021. The Index-Based Ecological Condition Assessment (IBECA) - Technical protocol, version 1.0. NINA Report 1967. Norwegian Institute for Nature Research.

## Analyses
### Data sets
The analyses in this document make use of the following data sets:
-NiN data (Naturtyper etter Miljødirektoratets innstruks)
-Sentinel-2 NDVI data
-MODIS NDVI data
-Landsat NDVI data

#### NiN data and geometry for Norway and the five regions
We read the NiN data for this document from a cache at the NINA server...
```{r, echo = F}
# Add NiN data from cache
nin <- readRDS(paste0(here::here(),"/data/cache/nin.RDS"))
```

...but it can also be directly downloaded from Miljødirektoratets kartkatalog (https://kartkatalog.miljodirektoratet.no/Dataset)
```{r, echo = T, eval = F}

url <- "https://nedlasting.miljodirektoratet.no/Miljodata//Naturtyper_nin/FILEGDB/4326/Naturtyper_nin_0000_norge_4326_FILEGDB.zip"
download(url, dest="P:/41201785_okologisk_tilstand_2022_2023/data/Naturtyper_nin_0000_norge_4326_FILEGDB.zip", mode="wb") 
unzip ("P:/41201785_okologisk_tilstand_2022_2023/data/Naturtyper_nin_0000_norge_4326_FILEGDB.zip", 
       exdir = "P:/41201785_okologisk_tilstand_2022_2023/data/nin_data")
st_layers("P:/41201785_okologisk_tilstand_2022_2023/data/nin_data/Naturtyper_nin_0000_norge_4326_FILEGDB.gdb")
nin <- st_read("P:/41201785_okologisk_tilstand_2022_2023/data/nin_data/Naturtyper_nin_0000_norge_4326_FILEGDB.gdb",
                 layer="naturtyper_nin_omr")
```
And we upload maps for Norway and the five regions: Southern, Western, Eastern, Central, and Northern Norway
```{r}
# Import region- og Norgeskart, make sure they have the same geometry as the nin data
nor <- st_read("data/outlineOfNorway_EPSG25833.shp")%>%
  st_as_sf() %>%
  st_transform(crs = st_crs(nin))

reg <- st_read("data/regions.shp")%>%
  st_as_sf() %>%
  st_transform(crs = st_crs(nin))

# change region names to something R-friendly
reg$region
reg$region <- c("Northern Norway","Central Norway","Eastern Norway","Western Norway","Southern Norway")

# combine the Norway and the region maps
regnor <- st_intersection(reg,nor)
```

#### NDVI data
All NDVI data are calculated mean NDVI values for each respective NiN polygon from each available Sentinel-2, MODIS, or Landsat image in Google Earth Engine (GEE).

##### Sentinel-2 NDVI data
The GEE code can be seen here: https://code.earthengine.google.com/2ceb0c3e03adade9e6f6d0903184b8c4
The image collection contains Sentinel imagery from June, July and August 2015-2022
To not exceed GEE memory limits, the exported files had to be iterated over a grid which resulted in 42 separate csv files. This script merges them and then merges the dataframe to the NiN data.
```{r}
# Import Sentinel NDVI Data
df.s <- list.files("P:/41201785_okologisk_tilstand_2022_2023/data/NDVI_åpenlavland/NDVI_data_Sentinel/", pattern = "*.csv", full.names=TRUE) %>%
  map_df(~fread(.))
df.s
```

##### MODIS NDVI data
The GEE code can be seen here: https://code.earthengine.google.com/efb84013701f1d5f6e1e81345f389b84
The image collection contains MODIS imagery from June, July and August 2000-2022

```{r}
# Import MODIS NDVI Data
# MODIS NDVI is scaled by 0.0001. Mean must be divided by 10000.
df.m<- read.csv("P:\\41201785_okologisk_tilstand_2022_2023\\data\\NDVI_åpenlavland\\NDVI_data_MODIS\\modis_ndvi_ts_2000_2022.csv", )
df.m$mean<-df.m$mean/10000
df.m
```

##### Landsat NDVI data
The GEE code can be seen here: https://code.earthengine.google.com/da8a9279238ef26d14be08a43788b6b7
The image collection contains Landsat imagery from June, July and August 1984-2022
To not exceed GEE memory limits, the exported files had to be iterated over a grid which resulted in 42 separate csv files. This script merges them and then merges the dataframe to the NiN data.
```{r}
# Import Landsat NDVI Data
# Set up conditional file paths
dir <- substr(getwd(), 1,2)

path <- ifelse(dir == "C:", 
               "R:/GeoSpatialData/Habitats_biotopes/Norway_Miljodirektoratet_Naturtyper_nin/Original/versjon20221231/Natur_Naturtyper_nin_norge_med_svalbard_25833/Natur_Naturtyper_NiN_norge_med_svalbard_25833.gdb",
               "/data/R/GeoSpatialData/Habitats_biotopes/Norway_Miljodirektoratet_Naturtyper_nin/Original/versjon20221231/Natur_Naturtyper_nin_norge_med_svalbard_25833/Natur_Naturtyper_NiN_norge_med_svalbard_25833.gdb")

pData <- ifelse(dir == "C:", 
                "P:/41201785_okologisk_tilstand_2022_2023/data/NDVI_åpenlavland/NDVI_data_Landsat",
                "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/NDVI_åpenlavland/NDVI_data_Landsat")

# Fread doesn't like the weird path to the server version of the P drive
# hence this horrendous work around
files=list.files(pData, pattern = "*.csv", full.names = TRUE)

df_list<-list()# initialise a list of dataframes
# read in a dataframe in each slot of the df_list
for (i in files){
  name <- gsub("-",".",i)
  name <- gsub(".csv","",name)  
  i <- paste(i,sep="")
  df_list[[i]]=assign(name,read.csv(i, header=TRUE))
}

df.l<-bind_rows(df_list, .id = "column_label")
df.l

```

### Data handling
#### NiN data
```{r, echo = T, eval = F}
# fixing variable- and ecosystem-names with special characters
colnames(nin)
colnames(nin)[c(3,8,17,26,31,33,34)] <- c("hovedoekosystem","kartleggingsaar","noyaktighet",
                                          "omraadenavn","uk_naertruet","uk_sentraloekosystemfunksjon",
                                          "uk_spesieltdaarligkartlagt")
unique(nin$hovedoekosystem)

nin <- nin %>% mutate(hovedoekosystem = recode(hovedoekosystem, 
                                               "Våtmark" = 'Vaatmark',
                                               "Semi-naturlig mark" = 'Semi_naturlig',
                                               "Naturlig åpne områder i lavlandet" = 'Naturlig-aapent',
                                               "Naturlig åpne områder under skoggrensa" = 'Naturlig_aapent')) %>%
  mutate(validGeo = st_is_valid(SHAPE))

# checking how many polygons have multiple ecosystem types
unique(nin$ninkartleggingsenheter)

nrow(nin)
# 95469 polygons altogether
nrow(nin %>%
       filter(grepl(',', ninkartleggingsenheter))
)
# 21094 polygons have more than 1 ecosystem type (they are separated by commas in the ninkartleggingsenheter-variable)

nrow(nin %>%
       filter(!grepl(',', ninkartleggingsenheter))
)
# 74375 polygons should have only 1 ecosystem type

# there's no information on the proportion of ecosystem types in the polygons, so we got to omit all polygons with multiple ecosystem types :(
nin <- nin %>%
  filter(!grepl(',', ninkartleggingsenheter))

# fix the content in the ninkartleggingsenheter-variable
summary(as.factor(nin$ninkartleggingsenheter))
# get rid of the NA- in the beginning
nin <- nin %>% mutate(ninkartleggingsenheter = str_remove(ninkartleggingsenheter, 'NA_'))
# making a main ecosystem type variable
nin <- nin %>% mutate(hovedtype = substr(ninkartleggingsenheter, 1, 3),
                      hovedtype = str_remove(hovedtype, '-'))
# checking mapping unit against main ecosystem type
nin[,c("hovedoekosystem","hovedtype")]
summary(as.factor(nin$hovedtype[nin$hovedoekosystem=="Vaatmark"]))
summary(as.factor(nin$hovedtype[nin$hovedoekosystem=="Semi_naturlig"]))
summary(as.factor(nin$hovedtype[nin$hovedoekosystem=="Naturlig_aapne"]))
summary(as.factor(nin$hovedtype[nin$hovedoekosystem=="Skog"]))
summary(as.factor(nin$hovedtype[nin$hovedoekosystem=="Fjell"]))
# making a new variable for the overarching ecosystem types based on the main ecosystem types
nin$hovedoekosystem.orig <- nin$hovedoekosystem

nin <- nin %>%
  mutate(hovedoekosystem = case_when(hovedtype %in% paste("V",c(1,3:7,9:10),sep="") ~ 'Vaatmark',
                                     hovedtype %in% paste("V",11:13,sep="") ~ 'Vaatmark_sterkt_endret',
                                     hovedtype %in% paste("T",c(31:34,40:41),sep="") ~ 'Semi_naturlig',
                                     hovedtype %in% paste("T",c(2,12,18,20:21),sep="") ~ 'Naturlig_aapent',
                                     hovedtype %in% c(paste("T",c(4,30,38),sep=""),paste("V",c(2,8),sep="")) ~ 'Skog',
                                     hovedtype %in% c(paste("T",c(3,7,9,10,14,22,26),sep=""),paste("V",c(6,7),sep="")) ~ 'Fjell',
                                     TRUE ~ 'NA'))

summary(as.factor(nin$tilstand))

nin <- nin %>% mutate(tilstand = recode(tilstand,
                                        "Dårlig" = "Redusert",
                                        "Svært redusert" = "Svaert_redusert"))
nin <- nin %>% mutate(tilstand_num = recode(tilstand,
                                        "God" = 1, "Moderat" = 2, 
                                        "Redusert" = 3, "Svaert_redusert" = 4))

summary(as.factor(nin$tilstand))
summary(as.numeric(nin$tilstand_num))

## filter out only wetland data
nin.wetland <- nin %>% 
  filter(hovedoekosystem %in% c('Vaatmark')) %>%
  mutate(id = identifikasjon_lokalid) %>%
  filter(validGeo) %>%
  drop_na(tilstand) %>%
  dplyr::select(id, hovedoekosystem, hovedtype, ninkartleggingsenheter, lokalitetskvalitet, tilstand, tilstand_num, ninbeskrivelsesvariabler, kartleggingsaar)

# merge NiN-data with region
nin.wetland = st_join(nin.wetland, regnor, left = TRUE)
nin.wetland

colnames(nin.wetland)[c(1,9)] <- c("id","region_id")

# check that every nin-polygon still occurs only once
summary(as.factor(nin.wetland$id))
nin.wetland[nin.wetland$id=="NINFP2110008060",]
# this one was assigned to both Western and Central Norway, we drop the latter
nin.wetland <- nin.wetland[!row.names(nin.wetland) %in% '1144.1',]

nin.wetland <- nin.wetland %>% 
  mutate(area_meters_nin = st_area(nin.wetland)
  )
# check and edit the order of regions
levels(nin.wetland$region)
nin.wetland$region <- as.factor(nin.wetland$region)
levels(nin.wetland$region)
nin.wetland$region <- factor(nin.wetland$region, levels = c("Northern Norway","Central Norway","Eastern Norway","Western Norway","Southern Norway"))
levels(nin.wetland$region)

# splitting content in the ninbeskrivelsesvariabler-column into separate rows per variable and then separating that column into two for the variable name and value
nin.wetland2 <- nin.wetland %>% separate_rows(ninbeskrivelsesvariabler, sep=",") %>%
  separate(col = ninbeskrivelsesvariabler,
           into = c("NiN_variable_code", "NiN_variable_value"),
           sep = "_",
           remove=F) %>%
  mutate(NiN_variable_value = as.numeric(NiN_variable_value)) %>%
  mutate(NiN_variable_code = as.factor(NiN_variable_code)) %>%
  subset(select = -ninbeskrivelsesvariabler)
nin.wetland2 <- as.data.frame(nin.wetland2)

```

```{r, echo = F}

#load processed NiN-data from cache
nin.wetland <- readRDS(paste0(here::here(),"/data/cache/nin.wetland.RDS"))
nin.wetland2 <- readRDS(paste0(here::here(),"/data/cache/nin.wetland2.RDS"))

```

#### NDVI data
```{r}
## Sentinel-2
# join nin.wetland & Sentinel NDVI data
SentinelNDVI.wetland <- full_join(nin.wetland, df.s, by="id")
#summary(SentinelNDVI.wetland)
SentinelNDVI.wetland <- SentinelNDVI.wetland %>%
  mutate(hovedoekosystem = as.factor(hovedoekosystem),
         hovedtype = as.factor(hovedtype),
         ninkartleggingsenheter = as.factor(ninkartleggingsenheter), 
         lokalitetskvalitet = as.factor(lokalitetskvalitet),
         tilstand = as.factor(tilstand),
         area_meters = st_area(SentinelNDVI.wetland))
#summary(SentinelNDVI.wetland)
# get rid of NAs (i.e. NDVI cells that were not in wetland polygons)
SentinelNDVI.wetland <- SentinelNDVI.wetland %>% filter(!is.na(hovedtype))
SentinelNDVI.wetland <- SentinelNDVI.wetland %>% filter(!is.na(mean))
#summary(SentinelNDVI.wetland)
# get rid of any nin-polygons smaller than the Sentinel grid cell size (100 sqm)
dim(SentinelNDVI.wetland)
SentinelNDVI.wetland <- SentinelNDVI.wetland %>% filter(as.numeric(area_meters) >= 100)
dim(SentinelNDVI.wetland)
# split date into year, month & day
SentinelNDVI.wetland <- SentinelNDVI.wetland %>%
  dplyr::mutate(year = lubridate::year(date), 
                month = lubridate::month(date), 
                day = lubridate::day(date))
# add column for sub-ecosystem types
SentinelNDVI.wetland <- SentinelNDVI.wetland %>% mutate(subtype = substring(ninkartleggingsenheter, 4),
                                                        subtype = str_remove(subtype, '-'),
                                                        subtype = str_remove(subtype, '-'))
# we are using max NDVI per year in every NiN polygon
SentinelNDVI.wetland <- SentinelNDVI.wetland %>%
  group_by(id, year) %>%
  filter(mean == max(mean, na.rm=TRUE))

summary(SentinelNDVI.wetland)

## MODIS
# join nin.wetland and Modis NDVI data
ModisNDVI.wetland <- full_join(nin.wetland, df.m, by="id")

#summary(ModisNDVI.wetland)
ModisNDVI.wetland <- ModisNDVI.wetland %>%
  mutate(hovedoekosystem = as.factor(hovedoekosystem),
         hovedtype = as.factor(hovedtype),
         ninkartleggingsenheter = as.factor(ninkartleggingsenheter), 
         lokalitetskvalitet = as.factor(lokalitetskvalitet),
         tilstand = as.factor(tilstand),
         area_meters = st_area(ModisNDVI.wetland))
#summary(ModisNDVI.wetland)
# get rid of NAs (i.e. NDVI cells that were not in wetland polygons)
ModisNDVI.wetland <- ModisNDVI.wetland %>% filter(!is.na(hovedtype))
ModisNDVI.wetland <- ModisNDVI.wetland %>% filter(!is.na(mean))
#summary(ModisNDVI.wetland)
# get rid of any nin-polygons smaller than the Modis grid cell size (62500 sqm)
dim(ModisNDVI.wetland)
ModisNDVI.wetland <- ModisNDVI.wetland %>% filter(as.numeric(area_meters) >= 62500)
dim(ModisNDVI.wetland)
# split date into year, month & day
ModisNDVI.wetland <- ModisNDVI.wetland %>%
  dplyr::mutate(year = lubridate::year(date), 
                month = lubridate::month(date), 
                day = lubridate::day(date))
# we are using max NDVI per year in every NiN polygon
ModisNDVI.wetland <- ModisNDVI.wetland %>%
  group_by(id, year) %>%
  filter(mean == max(mean, na.rm=TRUE))
#summary(ModisNDVI.wetland)

## Landsat
# join nin.wetland & Landsat NDVI data
LandsatNDVI.wetland <- full_join(nin.wetland, df.l, by="id")

#summary(LandsatNDVI.wetland)
LandsatNDVI.wetland <- LandsatNDVI.wetland %>%
  mutate(hovedoekosystem = as.factor(hovedoekosystem),
         hovedtype = as.factor(hovedtype),
         ninkartleggingsenheter = as.factor(ninkartleggingsenheter), 
         lokalitetskvalitet = as.factor(lokalitetskvalitet),
         tilstand = as.factor(tilstand),
         area_meters = st_area(LandsatNDVI.wetland))
#summary(LandsatNDVI.wetland)
# get rid of NAs (i.e. NDVI cells that were not in wetland polygons)
LandsatNDVI.wetland <- LandsatNDVI.wetland %>% filter(!is.na(hovedtype))
LandsatNDVI.wetland <- LandsatNDVI.wetland %>% filter(!is.na(mean))
#summary(LandsatNDVI.wetland)
# get rid of any nin-polygons smaller than the Landsat grid cell size (900 sqm)
dim(LandsatNDVI.wetland)
LandsatNDVI.wetland <- LandsatNDVI.wetland %>% filter(as.numeric(area_meters) >= 900)
dim(LandsatNDVI.wetland)
# split date into year, month & day
LandsatNDVI.wetland <- LandsatNDVI.wetland %>%
  dplyr::mutate(year = lubridate::year(date), 
                month = lubridate::month(date), 
                day = lubridate::day(date))
# we are using max NDVI per year in every NiN polygon
LandsatNDVI.wetland <- LandsatNDVI.wetland %>%
  group_by(id, year) %>%
  filter(mean == max(mean, na.rm=TRUE))

#summary(LandsatNDVI.wetland)
```

### Exploratory analyses, Sentinel-2

```{r}
## how are polygon sizes distributed?
summary(SentinelNDVI.wetland$area_meters)
hist(SentinelNDVI.wetland$area_meters,xlim=c(0,20000),breaks=50000)
abline(v=100,lty=2)
# wetland polygons stretch a large number of sizes from just above 100 sqm to mor ethan 4 million sqm. 75% of all polygons are below 14264 sqm.

## how does NDVI vary over the years (all years)
SentinelNDVI.wetland %>%
  ggplot( aes(x=year, y=mean )) + 
  geom_point() +
  facet_grid( tilstand~hovedtype)
# 2022 stands out with the highest NDVI values missing
# the pattern is strange, suggest to omit 2022 for this analysis
SentinelNDVI.wetland <- SentinelNDVI.wetland %>% filter(year != '2022')

## looking at major types for polygons in good condition only (and only for NDVI data matching NiN-mapping)
SentinelNDVI.wetland %>%
  filter(tilstand == 'God') %>%
  filter(year == kartleggingsaar) %>%
  ggplot(aes(x=hovedtype, y=mean )) + 
  geom_violin() +
  geom_point(size=0.7, shape=16, color="grey")
# NDVI varies between hovedtyper: highest in seminat wet-meadows (V10), lowest in ombrotrophic bogs (V3). The overlaps are considerable.

## looking at subtypes for polygons in good condition only (and only for NDVI data matching NiN-mapping)
SentinelNDVI.wetland %>%
  filter(tilstand == 'God') %>%
  filter(year == kartleggingsaar) %>%
  ggplot(aes(x=subtype, y=mean )) + 
  geom_violin() +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap( ~hovedtype)
# NDVI appears to increase with increasing limestone level (C1-C4, C5-C8, and E1-E3 in V1, C1-C3 in the other major types), but the overlap is large.

## NDVI across regions for polygons in good condition only (for major types and NDVI years data matching NiN-mapping years)
SentinelNDVI.wetland %>%
  filter(tilstand == 'God') %>%
  filter(year == kartleggingsaar) %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x=region, y=mean )) + 
  geom_violin() +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap( ~hovedtype) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust=0.2))
# no obvious trends in geography

## NDVI across condition classes (for major types and NDVI years data matching NiN-mapping years)
SentinelNDVI.wetland %>%
  filter(year == kartleggingsaar) %>%
  ggplot( aes(x=tilstand, y=mean )) + 
  geom_violin() +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap( ~hovedtype)
# in seminat types (V9,V10), kaldkilde (V4), and bogs (V3) ndvi seems to increase as condition deteriorates, in fens (V1) not much of a pattern. Overlaps between condition classes are generally very large.
```

Which ecological relevant variable(s) is the condition measure in the NiN data related to?
```{r}
nin.wetland2 %>%
  filter(hovedtype=='V9') %>%
  filter(NiN_variable_code %in% c("1AG-B","7GR-GI","1AG-A-G","1AR-C-L","7TK","PRSL")) %>%
  mutate(NiN_variable_des=recode(NiN_variable_code,
                                 "1AG-B"="Shrub cover",
                                 "7GR-GI"="Abundance of ditches",
                                 "1AG-A-G"="Tree cover",
                                 "1AR-C-L"="Woody plants in field layer",
                                 "7TK"="Heavy vehicles",
                                 "PRSL"="Wear/tear")) %>%
  ggplot(aes(x=NiN_variable_value, y=tilstand_num )) +
  geom_point(size=2) +  
  geom_jitter(width = 0.25, height=0.25) +
  facet_wrap(~NiN_variable_des,nrow=2, ncol=3) +
  theme(legend.position="none")

nin.wetland2 %>%
  filter(hovedtype=='V10') %>%
  filter(NiN_variable_code %in% c("7JB-BA","7RA-SJ","7FA","7JB-GJ")) %>%
  mutate(NiN_variable_des=recode(NiN_variable_code,
                                 "7JB-BA"="Land use intensity",
                                 "7RA-SJ"="Succession",
                                 "7FA"="Alien species",
                                 "7JB-GJ"="Fertilization")) %>%
  ggplot(aes(x=NiN_variable_value, y=tilstand_num )) +
  geom_point(size=2) +  
  geom_jitter(width = 0.25, height=0.25) +
  facet_wrap(~NiN_variable_des,nrow=2, ncol=2) +
  theme(legend.position="none")

nin.wetland2 %>%
  filter(hovedtype=='V1') %>%
  filter(NiN_variable_code %in% c("7GR-GI","PRTK","7FA","7TK","7SE","PRSL")) %>%
  mutate(NiN_variable_des=recode(NiN_variable_code,
                                 "7GR-GI"="Abundance of ditches",
                                 "PRTK"="Heavy vehicles",
                                 "7FA"="Alien species",
                                 "7TK"="Heavy vehicles2",
                                 "7SE"="Wear/tear",
                                 "PRSL"="Wear/tear2")) %>%
  ggplot(aes(x=NiN_variable_value, y=tilstand_num )) +
  geom_point(size=2) +  
  geom_jitter(width = 0.25, height=0.25) +
  facet_wrap(~NiN_variable_des,nrow=2, ncol=3) +
  theme(legend.position="none")

```
In these plots, condition=1 is best and 4 is worst (degraded). The panels for V1, V9 and V10 differ because Mdir-instruks uses different variables in these major ecosystem types to define condition.

In V9, semi-natural mires, reduced condition is mainly related to increasing cover of shrubs and trees. In V10, semi-natural wet-meadows, reduced condition is related to insufficient land-use intensity and succession. In both cases, we'd expect an increase in the greenness signal. Thus, NDVI can, in principle, be indicative of ecological condition in semi-natural wetlands due to a direct connection to the main pressures on the ecosystems.

In V1 (minerotrophic mires), reduced condition is mainly related to the abundance of ditches, but also increased use of heavy vehicles and high alien species abundance are related to very reduce condition. Under these drivers, greenness would not directly be affected. although drainage due to ditches may lead to succession and woody plant encroachment, and thus greening, in the longer run. Currently, NDVI seems poorly linked to ecological condition in and the major pressures on minerotrophic mires.

### Regression analyses Sentinel-2, NDVI as a function of condition
We can investigate further how NDVI and condition are connected using regression analyses.
```{r}
# NDVI data are bound between -1 and 1, and thus require modelling with an appropriate method that can handle bounded data
# we can transform the variable to be bounded between 0 and 1 and use beta-regression models
SentinelNDVI.wetland$mean_beta <- (SentinelNDVI.wetland$mean + 1) / 2
# NDVI data from the year of NiN-mapping (and thus with condition assessment) to train the condition models
# we drop V4 for the analysis as it lacks data for most combinations of condition and region, and thus would cause convergence issues

SentinelNDVI.wetland.train <- SentinelNDVI.wetland %>%
  filter(year == kartleggingsaar) %>%
  filter(hovedtype != 'V4')

# make an extra numeric condition variable
unique(SentinelNDVI.wetland.train$tilstand)
SentinelNDVI.wetland.train <- SentinelNDVI.wetland.train %>% mutate(tilstand_num = recode(tilstand, 
                                            "God" = '0',
                                            "Moderat" = '1',
                                            "Redusert" = '2',
                                            "Svært redusert" = '3'))
SentinelNDVI.wetland.train$tilstand_num <- as.numeric(SentinelNDVI.wetland.train$tilstand_num)

# We run a stepwise-function on the full model including condition, ecosystem type, and region to find the most parsimonious model
model.wetland.cond.Sent <- betareg(mean_beta~tilstand_num*region*hovedtype, data=SentinelNDVI.wetland.train)
model.wetland.cond.Sent <- StepBeta(model.wetland.cond.Sent)
summary(model.wetland.cond.Sent)
```
NDVI values do vary between:
- Regions
- Ecosystem types
- Condition classes

It is evident that there are differences between regions. But since we are not interested in the statistics for region differences, we run separate models for every region for ease of interpretation.
```{r, include = F}
rm(model.wetland.cond.Sent)
```

```{r}
model.wetland.cond.Sent.N <- betareg(mean_beta~tilstand_num*hovedtype, data=SentinelNDVI.wetland.train[SentinelNDVI.wetland.train$region=="Northern Norway",])
model.wetland.cond.Sent.C <- betareg(mean_beta~tilstand_num*hovedtype, data=SentinelNDVI.wetland.train[SentinelNDVI.wetland.train$region=="Central Norway",])
model.wetland.cond.Sent.W <- betareg(mean_beta~tilstand_num*hovedtype, data=SentinelNDVI.wetland.train[SentinelNDVI.wetland.train$region=="Western Norway",])
model.wetland.cond.Sent.E <- betareg(mean_beta~tilstand_num*hovedtype, data=SentinelNDVI.wetland.train[SentinelNDVI.wetland.train$region=="Eastern Norway",])
model.wetland.cond.Sent.S <- betareg(mean_beta~tilstand_num*hovedtype, data=SentinelNDVI.wetland.train[SentinelNDVI.wetland.train$region=="Southern Norway",])

```

# Northern Norway
```{r}
SentinelNDVI.wetland.train %>%
  filter(region=="Northern Norway") %>%

  ggplot( aes(x=tilstand, y=mean )) + 
  geom_violin() +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~hovedtype)

summary(model.wetland.cond.Sent.N)$coefficients
```
NDVI in good condition: V10 > V1 > V9 > V3

V1: NDVI decreases as condition deteriorates
V3: NDVI does not change as condition deteriorates
V9 & V10: NDVI increases as condition deteriorates

# Central Norway
```{r}
SentinelNDVI.wetland.train %>%
  filter(region=="Central Norway") %>%

  ggplot( aes(x=tilstand, y=mean )) + 
  geom_violin() +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~hovedtype)

summary(model.wetland.cond.Sent.C)$coefficients
```
NDVI in good condition: V10 > V1 & V9 > V3

V1: NDVI decreases as condition deteriorates
V3, V9 & V10: NDVI increases as condition deteriorates

# Western Norway
```{r}
SentinelNDVI.wetland.train %>%
  filter(region=="Western Norway") %>%

  ggplot( aes(x=tilstand, y=mean )) + 
  geom_violin() +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~hovedtype)

summary(model.wetland.cond.Sent.W)$coefficients
```
NDVI in good condition: V10 > V9 & V3 > V1

V1 & V10: NDVI increases as condition deteriorates
V3: NDVI decreases as condition deteriorates (but really insufficient data)
V9: NDVI does not change as condition deteriorates

# Eastern Norway
```{r}
SentinelNDVI.wetland.train %>%
  filter(region=="Eastern Norway") %>%

  ggplot( aes(x=tilstand, y=mean )) + 
  geom_violin() +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~hovedtype)

summary(model.wetland.cond.Sent.E)$coefficients
```
NDVI in good condition: V10 > V1 > V3 & V9

V1 & V3: NDVI does not change as condition deteriorates
V9 & V10: NDVI increases as condition deteriorates

# Southern Norway
```{r}
SentinelNDVI.wetland.train %>%
  filter(region=="Southern Norway") %>%

  ggplot( aes(x=tilstand, y=mean )) + 
  geom_violin() +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~hovedtype)

summary(model.wetland.cond.Sent.S)$coefficients
```
NDVI in good condition: V10 > V9 > V1 & V3

V1, V9 & V10: NDVI increases as condition deteriorates
V3: NDVI does not change as condition deteriorates (but also insuffcient data)

Overall take home messages:

In semi-natural wetland ecosystems (V9, V10) NDVI increases as ecosystem condition deteriorates, and this is consistent for the entire country. This supports the conclusion from the exploratory analyses that NDVI is indicative of ecological condition in semi-natural wetland ecosystems.
In natural fens (V1) NDVI increases in the South but decreases in the North as ecosystem condition deteriorates. A factor with potential for affecting vegetation composition and productivity, and thus NDVI, is Nitrogen deposition. Nitrogen deposition is higher in Southern parts of Norway than further north, and could thus be responsible for increased NDVI with reduced condition in the South, but it is not clear how Nitrogen deposition and the variables underlying the condition assessment in V1 (Ditches, Alien species, Heavy vehicles, Wear/tear) are connected. In bogs (V3) there is no consistent pattern.

### NDVI across time - Sentinel, MODIS & LandSat
As a last step, we can investigate how NDVI has changed over time. For this we include data from the MODIS and LandSat satellites in addition to Sentinel.

First, there's again some data handling to do. We merge the Sentinel, Modis, and Landsat data to show the full picture across time. Then we model the time series for each Satellite separately.
```{r}
## data handling for time series analysis
# Sentinel time series checked in exploratory analysis script
SentinelNDVI.wetland <- SentinelNDVI.wetland %>%
  group_by(id, year) %>%
  filter(mean == max(mean, na.rm=TRUE))
SentinelNDVI.wetland <- SentinelNDVI.wetland %>% filter(year != '2022')

# checking time series for MODIS
ModisNDVI.wetland %>%
  group_by(id, year) %>%
  filter(mean == max(mean, na.rm=TRUE)) %>%
  ggplot( aes(x=year, y=mean )) + 
  geom_point() +
  facet_wrap(~hovedtype)
# 2022 does not stand out as in the Sentinel data, so we keep it
ModisNDVI.wetland <- ModisNDVI.wetland %>%
  group_by(id, year) %>%
  filter(mean == max(mean, na.rm=TRUE))

# checking time series for Landsat
LandsatNDVI.wetland %>%
  group_by(id, year) %>%
  filter(mean == max(mean, na.rm=TRUE)) %>%
  ggplot( aes(x=year, y=mean )) + 
  geom_point() +
  facet_wrap(~hovedtype)
# nothing worrying to see here either
LandsatNDVI.wetland <- LandsatNDVI.wetland %>%
  group_by(id, year) %>%
  filter(mean == max(mean, na.rm=TRUE))

# transformation of NDVI scale to beta scale
#SentinelNDVI.wetland$mean_beta <- (SentinelNDVI.wetland$mean + 1) / 2
ModisNDVI.wetland$mean_beta <- (ModisNDVI.wetland$mean + 1) / 2
LandsatNDVI.wetland$mean_beta <- (LandsatNDVI.wetland$mean + 1) / 2

# check if there's any 0s or 1s (which beta cannot handle)
summary(SentinelNDVI.wetland$mean_beta)
summary(ModisNDVI.wetland$mean_beta)
summary(LandsatNDVI.wetland$mean_beta)
# replace 1s in Landsat data with 0.9999
LandsatNDVI.wetland <- LandsatNDVI.wetland %>%
  mutate(mean_beta = replace(mean_beta, mean_beta == 1, 0.9999))

# check if the three Satelite objects have the same structure (for concatenating them)
names(SentinelNDVI.wetland)
names(ModisNDVI.wetland)
names(LandsatNDVI.wetland)
# Sentinel and Landsat have each an extra column -> omit them when concatenating further below
# one column is named slightly differently in the Sentinel data -> rename it
SentinelNDVI.wetland <- SentinelNDVI.wetland %>%
  rename('system.index' = 'system:index')

# check if they have the same geometry
st_crs(SentinelNDVI.wetland)
st_crs(ModisNDVI.wetland)
st_crs(LandsatNDVI.wetland)
# all good

# add an increment to the year variable to avoid overlapping data being hidden in figures
SentinelNDVI.wetland$year_jit <- SentinelNDVI.wetland$year + 0.3
ModisNDVI.wetland$year_jit <- ModisNDVI.wetland$year - 0.3
LandsatNDVI.wetland$year_jit <- LandsatNDVI.wetland$year

# concatenate the three Satellite objects
allSatNDVI.wetland <- rbind(
SentinelNDVI.wetland[,!names(SentinelNDVI.wetland) %in% "subtype"],
ModisNDVI.wetland,
LandsatNDVI.wetland[,!names(LandsatNDVI.wetland) %in% "column_label"]
)
# add variable for Satellite identity
allSatNDVI.wetland$Sat <- c(
  rep("Sentinel",nrow(SentinelNDVI.wetland)),
  rep("Modis",nrow(ModisNDVI.wetland)),
  rep("Landsat",nrow(LandsatNDVI.wetland))
)
allSatNDVI.wetland$Sat <- factor(allSatNDVI.wetland$Sat,levels=c("Sentinel","Modis","Landsat"))
levels(allSatNDVI.wetland$Sat)
```
Now we can plot the time series for each main ecosystem type, showing each satellite separately  
```{r}
# plot 
allSatNDVI.wetland %>%
  ggplot( aes(x=year_jit, y=mean, color=Sat )) + 
  geom_point() +
  facet_wrap(~hovedtype, ncol = 1)
```
It is quite obvious that the NDVI values from the three satellites are not quantitatively comparable. They vary both in their placement along the y-axis and in their variance. Modis data are absent from the V4-ecosystem type because of a too large pixel size compared to the NiN-polygon sizes. In V10, Modis is poorly represented for the same reason.

We can test if there is consistent tempral changes in NDVI by running regressions including year as a covariate.

We first investigate a model testing NDVI as a function of year and main ecosystem type for polygons in good condition, and we do so separately for each satellite.
```{r}
model.wetland.time.GodTilst.Sent <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=SentinelNDVI.wetland[SentinelNDVI.wetland$tilstand=="God",])
model.wetland.time.GodTilst.Modi <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=ModisNDVI.wetland[ModisNDVI.wetland$tilstand=="God",])
model.wetland.time.GodTilst.Land <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=LandsatNDVI.wetland[LandsatNDVI.wetland$tilstand=="God",])

summary(model.wetland.time.GodTilst.Sent)$coefficients$cond
summary(model.wetland.time.GodTilst.Modi)$coefficients$cond
summary(model.wetland.time.GodTilst.Land)$coefficients$cond
```

```{r}
model.wetland.time.Sent <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=SentinelNDVI.wetland)
model.wetland.time.Modi <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=ModisNDVI.wetland)
model.wetland.time.Land <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=LandsatNDVI.wetland)

summary(model.wetland.time.Sent)$coefficients$cond
summary(model.wetland.time.Modi)$coefficients$cond
summary(model.wetland.time.Land)$coefficients$cond
```

### Eksport file (final product)
<!-- Export final file. Ideally a georeferenced shape or raster wit indicators values (raw and normalised), reference values and errors. -->