### 2023-10-18 indipri.Rmd
add chapter to embed indipri

# IndiPri

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, out.width='100%', echo=FALSE}
knitr::include_app("https://anderskolstad.shinyapps.io/indiPri/",
                   height = "800px")
```