### 2023-10-19 DRAFT_birdIndicators.Rmd
About the data

---
title: "Breeding bird indicators"
output:
  html_document
---

# Breeding bird indicators

<br />

_Author and date:_
Chloé R. Nater, Diego Pavón-Jordán

<br />

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F}
Ecosystem <- "Skog, fjell, våtmark, seminaturlig mark" 
Egenskap  <- "Biologisk mangfold" 
ECT       <- "Compositional state characteristic" 
Contact   <- "Chloé R. Nater" 
```

```{r, echo=F}
metaData <- data.frame(Ecosystem,
                       "Økologisk egenskap" = Egenskap,
                       "ECT class" = ECT)
knitr::kable(metaData)
```

<br />
<br />
<hr />

## Introduction
The Norwegian common breeding bird monitoring scheme (hekkefuglovervåking in Norwegian; formerly called TOV-E) was established between 2005 and 2010. The main goal of the programme is to gather data on the abundance of the most common terrestrial bird species to assess population trends and changes across Norway (available at the website), and help proposing efficient management and conservation measures for these species and their habitats in Norway. These data are the basis to update the Red List of Birds in Norway following the IUCN protocol, and are also used to generate indicators of mountain, farmland and forest habitats.

Furthermore, these data are shared with the Pan-European Common Bird Monitoring Scheme (PECBMS; https://pecbms.info/), which, by combining data from different long-term and large-scale bird monitoring programmes in all European countries, aims at producing indicators of the general state of nature at continental scale.

Data from hekkefuglovervåking are also use in scientific research both at a national and supra-national scale.

The hekkefuglovervåking, coordinated by NINA and BirdLife Norway and supported financially by the Norwegian Environment Agency (Miljødirektoratet), replaced the previous scheme, which ran between 1995 and 2008.

## About the underlying data
These data are gathered mostly by volunteer experienced bird-watchers following a standardized sampling protocol (https://hekkefuglovervakingen.nina.no/Fugl/public/papirskjema/MethodologyEng.pdf). Observers will visit each route once a year between 20.05 and 10.06 (depending on latitude) and carry out the census in good weather conditions. The observer will record any bird seen or heard during 5 minutes at each of the pre-selected counting points in the route. A total of 229 species are typically recorded during this census. After 5 minutes, the observer walks to the next point in a straight line (when possible). During this 'transect' the observer is requested to record also a subset of species that are considered rare or difficult to capture during the 5-minute point counts (n = 121).

### Representativity in time and space
Nation-wide monitoring scheme consisting in 492 fix routes across all terrestrial habitats present in Norway (MAP HERE?? I can provide this). Each route has between 12 and 20 counting points (ca. 9000 across Norway), which typically are 300 m apart, describing a square with side 1.5 km long (the observer walks on a straight line between two consecutive points). Deviations of this typical route shape are normally due to the landscape configuration (e.g. rough terrain, presence of a lake or river, sea, glacier...).  Due to lack on resources, only a random sample of 370 routes are surveyed every year. The aim is to survey at least 80% of the routes yearly.

This monitoring scheme is designed to gather representative sample in all (common) habitats in both the latitudinal and altitudinal gradients.

Although this monitoring scheme started in 2005, full coverage at national scale was achieved in 2010.

### Original units
The observers will record 'number of individuals' seen or heard during the census. Then, those values are converted into 'number of pairs' by multiplying the number of individuals by a conversion factor, which is 2 or 5 depending on the species (for national reporting only).
Information on bird abundance is then processed in the software TRIM to estimate abundance indices relative to a base year (in Norway this is typically 2008). Such abundance indices are used for reporting population trends across Norway, generate multi-species indicators in Norway and shared with PECBMS for pan-European assessment of bird populations and their habitats.

### Temporal coverage
The Norwegian common breeding bird monitoring scheme was established in 2005 in central Norway and but reached full development (nation-wide) in 2010. These scheme provides high quality data for many species already since 2008. This monitoring programme is still ongoing as of 2023.
A previous monitoring scheme existed, which was established in 1995 and was discontinued in 2008. 
Although these two monitoring programmes are not the same, tools available (e.g. RSWAN) can combine the abundance indices calculated with TRIM from the two different schemes and then combine them for a single trend estimate. This is possible because both monitoring schemes overlapped for one year (2008).

### Aditional comments about the dataset
From all species recorded within the Norwegian common breeding bird monitoring scheme (n = 229), only 71 species are requested by PECBMS (most widely distribute with continental-wide relevance). Longer (combined) time series can be calculated for 55 species of which 39 species are used for the multi-species indicators (farmland = 7 species; forest = 24 species; mountain = 8 species).

## Ecosystem characteristic

### Norwegain standard

Breeding bird indicators are regarded as indicators of biodiversity ("biologisk mangfold").
It is important to note though that the indicators represent species abundance (trends) in this context and are **not** indicators of species richness or species distribution.

### SEEA EA (UN standard)
Breeding bird indicators are tagged as "Compositional state characteristic".
We develop the indicators here to be as consistent as possible with the internationally reported and well-established farmland bird index (https://agridata.ec.europa.eu/Qlik_Downloads/InfoSheetEnvironmental/infoC35.html) and common bird indices (https://www.eea.europa.eu/ims/abundance-and-distribution-of-selected).

## Collinearities with other indicators
Several species that are part of the breeding bird indicators are also part of the Norwegian Nature Index (NI). For some ecosystems, e.g. wetlands, NI is dominated by bird indicators, potentially leading to substantial collinearity with the corresponding breeding bird index. In the chapter on NI products, we suggest to remove bird indicators from NI products when using them in ecosystem condition assessments that also make use of composite breeding bird indicators.

## Reference condition and values
### Reference condition
In general, the reference condition for breeding bird indicators should be defined along the same lines as for the other indicators in ecological condition assessments.

The internationally established composite bird indicators (farmland bird index, common bird index, etc.), however, are trend indicators can be presented with a few different baseline years. It is the availability of monitoring data of sufficient quality that determines which years may be used as baselines, and the "Hekkefuglovervåkingen", the monitoring programme that supplies the data for the Norwegian breeding bird indicators, has been running since 2007/2008. For some species, a predecessor monitoring scheme has been providing data since 1996. The possible baselines for the breeding bird indicators can therefore be no earlier than 1996, and using baselines consistent with the other nationally (https://miljostatus.miljodirektoratet.no/tema/arter/fugler/; https://hekkefuglovervakingen.nina.no/Fugl/Default.aspx) and internationally reported indices (https://pecbms.info/trends-and-indicators/) seems sensible.

We note that composite breeding bird indicators could also be built as Nature Index (NI) product (see separate chapters). For the species included in NI, reference values corresponding to "ecosystems with minimal human impacts" were determined previously, and an NI product for a selection of breeding birds would therefore be an indicator taking values between 0 and 1, where 1 = reference condition.
We refer to the chapter on NI products for more information (and implementation code) for this alternative approach.

### Reference values, thresholds for defining _good ecological condition_, minimum and/or maximum values
The breeding bird indicators represent average trends relative to a reference year and are therefore scaled with respect to population size in the reference year. Therefore, the indicator value in the reference year is 1. In theory, the indicators can take values between 0 (representing complete extinction of all species) and infinity (infinite population increase). 
The reference value 1 represents unchanged abundance relative to the reference year. Determining a threshold for what level of decrease in bird abundance is acceptable under "good ecological condition", however, is a challenging task. One approach to setting a threshold may be to consider thresholds of population decline used in other frameworks. The IUCN Red List of threatened species, for example, considers single species as "endangered" when a population decrease of 30% or more has been observed for the most recent decade [@iucn2001]. While it is unclear to what degree that can (or should) be extended to assemblages of multiple species, one may use this as a guideline to stipulate that >70% of the original population (indicator value 0.7) should be present (i.e. the species assemblage should not be endangered) when¨the ecosystem is in good condition. Another approach to determining a threshold would be to assume that bird populations should not be declining at all if an ecosystem is in good condition, i.e. that 1 is the threshold for good ecological condition instead of the reference value. In that case, however, one may need to come up with a reference value that is greater than 1, and determining what level of population increase would be ideal is a nearly impossible task.
We want to highlight here that these are very preliminary thoughts, and that the reference and threshold values have not been discussed in any more details yet. These discussions will nonetheless be needed, and possibly augmented with simulations, before putting the breeding bird indicators to use for ecosystem condition assessments.

## Uncertainties
<!-- Text here -->

## References

::: {#refs}
:::

## Analyses
### Data sets
<!-- Text here -->

#### Monitoring data from "Hekkefuglovervåkingen"
<!-- Text here -->

#### PECBMS Trim results from predecessor monitoring programme
<!-- Text here -->

#### Regions

As per now, the breeding bird indicators are calculated at the national scale only and the data is therefore not linked to any geofiles here.

We note that some species may be sufficiently common and have enough data to allow for calculating indicator values for two major regions, Northern Norway and Southern Norway, but this requires careful evaluation that needs to be planned and organized as part of a future project.

### PECBMS Trim analyses

### Combining data series using SWAN

### Scaled indicator values

### Uncertainty

## Prepare export

### Eksport file (final product)