### 2023-10-20 DRAFT_birdIndicators.Rmd
Edited / added to Diego's text

---
title: "Breeding bird indicators"
output:
  html_document
---

# Breeding bird indicators

<br />

_Author and date:_
Chloé R. Nater, Diego Pavón-Jordán

<br />

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F}
Ecosystem <- "Skog, fjell, våtmark, seminaturlig mark" 
Egenskap  <- "Biologisk mangfold" 
ECT       <- "Compositional state characteristic" 
Contact   <- "Chloé R. Nater" 
```

```{r, echo=F}
metaData <- data.frame(Ecosystem,
                       "Økologisk egenskap" = Egenskap,
                       "ECT class" = ECT)
knitr::kable(metaData)
```

<br />
<br />
<hr />

## Introduction
The Norwegian common breeding bird monitoring scheme ("hekkefuglovervåking" in Norwegian; formerly called TOV-E) was established between 2005 and 2010. The main goal of the programme is to gather data on the abundance of the most common terrestrial bird species to assess population trends and changes across Norway (available at the website), and help proposing efficient management and conservation measures for these species and their habitats in Norway. The "hekkefuglovervåking" is coordinated by NINA and BirdLife Norway and supported financially by the Norwegian Environment Agency (Miljødirektoratet), and has replaced predecessor monitoring schemes that ran from 1996 to 2008.

These data collected through "hekkefuglovervåkingen" form the basis for a variety of national and international assessments including nationally reported species population trends (https://hekkefuglovervakingen.nina.no/hekkefugl), updates to the Red List of Birds in Norway following the IUCN protocol (https://artsdatabanken.no/lister/rodlisteforarter), indicators for the Norwegian Nature Index (https://www.naturindeks.no/), and single- and multi-species indicators that are reported nationally (https://miljostatus.miljodirektoratet.no/tema/arter/fugler/) and internationally to e.g. the Pan European Bird Monitoring Scheme (PECBMS, https://pecbms.info/) and the OECD.

Workflows generating outputs and numbers for several of these reporting routes are already integrated thanks to coordination and standardization efforts by PECMBS, and the calculation of breeding bird indicators for use in ecosystem condition assessments should be incorporated into the same workflow. We therefore suggest to use the same multi-species indicators in the framework for ecosystem condition assessment as are reported nationally and internationally: the established "Farmland bird index", "Mountain bird index", and "Forest bird index", as well as a corresponding index for wetland birds.

## About the underlying data
The data from "hekkefuglovervåkingen" are gathered mostly by experienced volunteer bird-watchers following a standardized sampling protocol (https://hekkefuglovervakingen.nina.no/Fugl/public/papirskjema/MethodologyEng.pdf). Observers will visit each route once a year between 20. May and 10. June (depending on latitude) and carry out the census in good weather conditions. The observers record any bird seen or heard during 5 minutes at each of the pre-selected counting points in the route. A total of 229 species are typically recorded during this census. After 5 minutes, the observer walks to the next point in a straight line (when possible). During this 'transect' the observer is requested to record also a subset of species that are considered rare or difficult to capture during the 5-minute point counts (n = 121).

The data are subsequently entered into a database (https://hekkefuglovervakingen.nina.no/Fugl/Default.aspx) and quality-checked by NINA. Subsequently, the raw data are processed using a combination of country-specific wrangling routines and internationally standardized scripts distributed by PECBMS, which together make up the now integrated workflow below:

<img src="Hekkefugl_Workflow.png" width="600"/>

### Representativity in time and space
"Hekkefuglovervåking" is a nation-wide monitoring scheme consisting of 492 fixed routes across all terrestrial habitats present in Norway (MAP HERE?? I can provide this). Each route has between 12 and 20 counting points (ca. 9000 across Norway), which typically are 300 m apart, describing a square with side 1.5 km long (the observer walks on a straight line between two consecutive points). Deviations of this typical route shape are normally due to the landscape configuration (e.g. rough terrain, presence of a lake or river, sea, glacier, etc.). Due to limited resources, only a random sample of 370 routes are surveyed every year with an aim to survey at least 80% of the routes yearly.

This monitoring scheme is therefore designed to gather representative samples in all (common) habitats along both latitudinal and altitudinal gradients.

Even though implementation of the monitoring scheme started in 2005, full coverage at the national scale was not achieved until 2010.

### Original units
The observers in the field record "number of individuals" seen or heard during the census. Those values are then converted into "number of pairs" by multiplying the number of individuals by a conversion factor, which is 2 or 5 depending on the species (for national reporting only).
In the first step of the workflow, the raw counts are converted into species-specific abundance indices, defined as % of the population in a given baseline year (in Norway this is typically 2008). The abundance indices then form the basis of further analyses and aggregation routines.

### Temporal coverage
The "hekkefuglovervåkingen" was established in 2005 in central Norway and expanded from there to reach full national coverage by 2010. For many species, data is already of good coverage and high quality starting in 2008. The monitoring programme is still ongoing as of 2023.

Predecessor monitoring schemes existed prior to the establishment of "hekkefuglovervåkingen" and led to the collection of decent data in the period from 1996 up to and including 2008. 
Although the predecessor monitoring programmes and "hekkefuglovervåkingen" are not the same, the overlap for the years 2007 and 2008 makes it possible to combine abundance indices from across time using tools like "RSWAN", which is provided by PECBMS. For a subset of 55 species, longer time-series of abundance indices (1996-2022, as of now) are therefore available.

### Aditional comments about the dataset
From all species recorded within the Norwegian common breeding bird monitoring scheme (n = 229), 85 have sufficient data for abundance trend analyses. Abundance indices of 71 of those species, representing those with wide distributions and high continental-scale relevance, are reported annually to PECBMS. Longer (combined) time series can be calculated for 55 species, and 54 of those serve as indicators in the Norwegian Nature Index. The established breeding bird indicators contain varying numbers of species: the farmland bird index has 7 species, the mountain bird index has 8 species, and the forest bird index has 24 species. For the new index for wetland birds, we used 7 out of the 8 species suggested for this purpose by @husby2011. The one species we dropped was the Common Crane _(Grus grus)_ as it has seen strong recent population increases that are linked to anthropogenic impacts unrelated to the status of their wetland breeding habitats (Bård Stokke, pers. communication).

## Ecosystem characteristic

### Norwegian standard

Breeding bird indicators are regarded as indicators of biodiversity ("biologisk mangfold").
It is important to note though that the indicators represent species abundance (trends) in this context and are **not** indicators of species richness or species distribution.

### SEEA EA (UN standard)
Breeding bird indicators are tagged as "Compositional state characteristic".
We develop the indicators here to be as consistent as possible with the internationally reported and well-established farmland bird index (https://agridata.ec.europa.eu/Qlik_Downloads/InfoSheetEnvironmental/infoC35.html) and common bird indices (https://www.eea.europa.eu/ims/abundance-and-distribution-of-selected).

## Collinearities with other indicators
Several species that are part of the breeding bird indicators are also part of the Norwegian Nature Index (NI). For some ecosystems, e.g. wetlands, NI is dominated by bird indicators, potentially leading to substantial collinearity with the corresponding breeding bird index. In the chapter on NI products, we suggest to remove bird indicators from NI products when using them in ecosystem condition assessments that also make use of composite breeding bird indicators.

## Reference condition and values
### Reference condition
In general, the reference condition for breeding bird indicators should be defined along the same lines as for the other indicators in ecological condition assessments.

The internationally established composite bird indicators (farmland bird index, common bird index, etc.), however, are trend indicators that can be presented with a few different baseline years. It is the availability of monitoring data of sufficient quality that determines which years may be used as baselines, and the "hekkefuglovervåkingen", the monitoring programme that supplies the data for the Norwegian breeding bird indicators, has been running since 2007/2008. For some species, predecessor monitoring schemes have been providing data since 1996. The possible baselines for the breeding bird indicators can therefore be no earlier than 1996 or 2008 depending on the species (assemblage), and using baselines consistent with the other nationally (https://miljostatus.miljodirektoratet.no/tema/arter/fugler/; https://hekkefuglovervakingen.nina.no/Fugl/Default.aspx) and internationally reported indices (https://pecbms.info/trends-and-indicators/) seems sensible.

We note that composite breeding bird indicators could also be built as Nature Index (NI) product (see separate chapters). For the species included in NI, reference values corresponding to "ecosystems with minimal human impacts" were determined previously, and an NI product for a selection of breeding birds would therefore be an indicator taking values between 0 and 1, where 1 = reference condition.
We refer to the chapter on NI products for more information (and implementation code) for this alternative approach.

### Reference values, thresholds for defining _good ecological condition_, minimum and/or maximum values
The breeding bird indicators represent average trends relative to a reference year and are therefore scaled with respect to population size in the reference year. Therefore, the indicator value in the reference year is 100 (percent). In theory, the indicators can take values between 0 (representing complete extinction of all species) and infinity (infinite population increase). 
The reference value 100 represents unchanged abundance relative to the reference year. Determining a threshold for what level of decrease in bird abundance is acceptable under "good ecological condition", however, is a challenging task. 
One approach to setting a threshold may be to consider thresholds of population decline used in other frameworks. The IUCN Red List of threatened species, for example, considers single species as "endangered" when a population decrease of 30% or more has been observed for the most recent decade or over the most recent 3 generations [@iucn2001]. While it is unclear to what degree that can (or should) be extended to assemblages of multiple species, one may use this as a guideline to stipulate that >70% of the original population (indicator value 70 when considering a recent trend over 10 years or 3 average generations) should be present (i.e. the species assemblage should not be endangered) when the ecosystem is in good condition. Another approach to determining a threshold would be to assume that bird populations should not be declining at all if an ecosystem is in good condition, i.e. that 100 is the threshold for good ecological condition instead of the reference value. In that case, however, one may need to come up with a reference value that is greater than 100, and determining what level of population increase would be ideal is a nearly impossible task.
We want to highlight here that these are very preliminary thoughts, and that the reference and threshold values have not been discussed in any more details yet. These discussions will nonetheless be needed, and possibly augmented with simulation studies, before putting the breeding bird indicators to use for ecosystem condition assessments.

## Uncertainties
There is a substantial amount of uncertainty related to which year should be considered as a baseline and how the reference and threshold values should be set (see preceeding section).

Another source of uncertainty comes from the assumption that multi-species abundance indices use a representative selection of species sampled in a representative set of habitats. However, due to resource- and logistic constraints, that is often not the case and non-random sampling along, for example, rural-urban gradients may bias breeding bird indices derived from data from "hekkefuglovervåkingen" [@husby2021].

## References

::: {#refs}
:::

## Analyses
### Data sets
All data used are gathered under the Norwegian common breeding bird monitoring scheme (Hekkeguglovervåking in Norwegian; formely named TOV-E). For the calculation of multi-species indicators (MSI) of farmland and forest habitats, however, a longer dataset is used by combininng data from the 'predecessor monitoring programme', whic ran from 1995 to 2008.

#### Monitoring data from "Hekkefuglovervåkingen"
The route-based data are compiled after each breeding season and stored at NINA's database. The route-based data are then used to calculate the long-term trends (since 2008) in abundance for XX species, which have sufficient data to properly estimate the yearly abundance indices at a National scale.

#### PECBMS Trim results from predecessor monitoring programme
The abundance indices generated by the PECBMS RTRIM workflow using data from the predecessor scheme are stored in NINAs local file repositories. Unlike the Trim analyses for the data from "hekkefuglovervåkingen", which are re-run every year after new data has been added, the old abundance indices remain unchanged as no new data are being added.

#### Regions

As per now, the breeding bird indicators are calculated at the national scale only and the data is therefore not linked to any geofiles here.

We note that some species may be sufficiently common and have enough data to allow for calculating indicator values for two major regions, Northern Norway and Southern Norway, but this requires careful evaluation that needs to be planned and organized as part of a future project.

### PECBMS RTRIM analyses

### Combining data series using RSWAN

### Aggregation to multi-species indices (MSIs)

### Uncertainty

## Prepare export

### Eksport file (final product)